### 2017-10-06 Proposal/eval.tex
Init Proposal.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Evaluation}
\label{sec:evaluation}
Maecenas aliquet sagittis nunc, ac sagittis justo scelerisque in. Maecenas at libero leo. Integer maximus gravida ante, accumsan tempor dui. Quisque accumsan semper erat, at mattis ex vestibulum at. Proin eu nulla vitae ex molestie facilisis eu id dolor. Phasellus lobortis efficitur purus, sit amet laoreet lacus sodales vulputate. Maecenas venenatis est a justo blandit consequat. Donec vitae pharetra orci. Duis tortor diam, interdum sed ex non, ultricies suscipit lorem. Pellentesque eros arcu, dignissim ut leo sit amet, sodales sodales tortor. Ut convallis, elit nec auctor mollis, erat neque feugiat ex, bibendum pharetra nulla risus sed neque. Aliquam feugiat elit metus, non convallis nisl malesuada pharetra. Donec ligula neque, blandit eget ligula placerat, lacinia finibus odio. Quisque sit amet suscipit diam.

Nunc consectetur turpis non sollicitudin pellentesque. Nullam nec eleifend lectus. Nunc sem enim, bibendum quis elit nec, laoreet posuere dui. Vivamus sodales nisi sed facilisis consectetur. Donec venenatis justo sed consectetur ullamcorper. Nulla dignissim magna eget arcu consectetur aliquet. Praesent consequat, elit quis auctor egestas, dolor mi pretium erat, eu ullamcorper risus velit sed metus. Praesent lectus urna, tempor sed imperdiet id, condimentum id nunc. Aliquam id mi sollicitudin, venenatis justo id, congue diam. Sed quis dictum odio. Donec auctor enim quis ultricies consequat. Cras maximus a velit luctus feugiat.

### 2017-10-06 Proposal/implementation.tex
Init Proposal.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implementation}
\label{sec:implementation}
Nunc ornare tortor at finibus faucibus. Sed bibendum efficitur nibh ac ultricies. Donec maximus purus lectus, id auctor diam molestie vel. Pellentesque ex odio, ultricies vel magna et, laoreet placerat augue. Suspendisse nunc ipsum, semper rhoncus sem at, dignissim mollis mauris. Cras viverra, eros eget faucibus porttitor, quam ex aliquet mi, eget sodales elit dolor in mi. Aenean congue augue id nunc blandit, ac molestie augue sagittis. Duis rhoncus odio eu arcu posuere, eget luctus diam lobortis. Aliquam sed ullamcorper velit. Suspendisse sed ligula sed risus rhoncus placerat ut fermentum nulla. Nam facilisis dui metus, non malesuada lacus feugiat quis. Vivamus volutpat non nibh eu gravida. Suspendisse efficitur leo maximus tortor interdum maximus. Nulla feugiat elit nec tincidunt tincidunt. In fermentum lacus tellus, sit amet tempor leo pulvinar quis.

Duis tortor mauris, dignissim vehicula ligula non, lobortis finibus odio. Curabitur eu lobortis erat. Donec arcu ex, cursus in finibus sit amet, commodo ac mi. Integer mattis iaculis ante placerat tempus. Phasellus eu diam blandit, ultricies mauris ut, viverra odio. Praesent accumsan, lectus ut accumsan semper, sem elit pharetra enim, at vehicula felis ante a tortor. Duis at tellus id enim mollis posuere et vitae nisi. Nam convallis vulputate commodo. Vestibulum eget libero sed eros dictum facilisis. Praesent posuere vulputate ipsum non sagittis.

### 2017-10-06 Proposal/intro.tex
Init Proposal.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
Vivamus et venenatis justo, sed porttitor diam. Sed sed lectus nunc. Duis cursus mi quis ornare fringilla. Fusce pellentesque nulla tellus, eget imperdiet enim varius vel. Nam viverra at augue non commodo. Maecenas porta lobortis erat in malesuada. Suspendisse pulvinar ut metus et venenatis. Vivamus quis tristique ligula. Nam vitae ex eget arcu gravida sollicitudin. Duis velit magna, pretium nec elit ut, placerat accumsan velit. Suspendisse orci nisl, imperdiet eget bibendum commodo, fermentum at libero. Etiam sit amet lectus hendrerit, convallis tortor bibendum, condimentum libero. Suspendisse sit amet elit metus. Nunc varius ac libero ultrices facilisis. Donec sit amet purus non ante efficitur ullamcorper vehicula at enim. Pellentesque at vulputate est, blandit porttitor felis.

Curabitur convallis, odio sed dapibus interdum, quam eros interdum leo, sed pretium urna tortor eu sem. Pellentesque id interdum nulla. Sed dapibus vel massa quis accumsan. Nulla vitae vehicula odio. Nam hendrerit ultricies elit sed convallis. Fusce efficitur pellentesque massa, at ultrices nulla semper et. Suspendisse in enim at justo ultrices ultrices. Quisque congue euismod odio, non vehicula elit condimentum vel. Vivamus vestibulum tellus ac libero facilisis molestie. Aliquam accumsan sit amet augue eget interdum. Aliquam feugiat tincidunt velit, sed tincidunt purus auctor a. Vestibulum nunc enim, dignissim vitae urna varius, sagittis lobortis purus. Cras eget quam non ipsum tincidunt pretium at eu est.

### 2017-10-06 Proposal/paper.tex
Init Proposal.

\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage[normalem]{ulem}
% Copyright
\setcopyright{none}

\input{annotations}

\begin{document}

\title{TCPTokens: Introducing Currency into Data Center Congestion Control}

\author{Anand Jayarajan}
\affiliation{%
  \institution{University of British Columbia}
}
\email{anandj@cs.ubc.ca}

\author{Michael Przystupa}
\affiliation{%
  \institution{University of British Columbia}
}
\email{michael.przystupa@gmail.com}

\author{Robert Reiss}
\affiliation{%
  \institution{University of British Columbia}
}
\email{rreiss@cs.ubc.ca}

\author{Fabian Ruffy}
\affiliation{%
  \institution{University of British Columbia}
}
\email{fruffy@cs.ubc.ca}

\author{??? ???}
\affiliation{%
  \institution{University of British Columbia}
}
\email{pzarei1993@gmail.com}

\begin{abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum lacinia arcu ac diam mattis pharetra. Curabitur luctus posuere dictum. In quam tortor, cursus luctus tortor ut, porta fermentum lacus. Nunc eget augue dui. Proin ante lectus, malesuada a sem ut, lobortis eleifend turpis. Morbi sit amet odio interdum, sollicitudin sem eget, mollis orci. Curabitur rhoncus arcu magna, eget rhoncus nunc porttitor at. Nulla quis posuere urna. Nam enim nunc, hendrerit et gravida et, mollis a urna. Donec aliquam massa sollicitudin tortor cursus placerat. Ut non nisi purus. Etiam laoreet ut libero et tempus. 
\end{abstract}

\keywords{TCP, Congestion Control, SDN, Data center}

\maketitle

\input{intro}
% \input{background}
\input{relatedworks}
\input{motivation}
\input{implementation}
\input{eval}

\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}

\end{document}

### 2017-10-06 Proposal/relatedworks.tex
Init Proposal.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
\label{sec:related}
Nulla facilisi. Donec congue, quam sed faucibus varius, libero dolor mollis ipsum, vel placerat felis urna quis nibh. Morbi tempor ipsum nec urna porttitor interdum. Phasellus cursus nisi nec tempor blandit. Curabitur at sem convallis, sagittis purus ornare, sagittis urna. Nullam eget eleifend orci. Suspendisse vel justo at elit mollis porta. Nulla congue sapien nibh, tincidunt pulvinar orci elementum in. Aliquam mollis tempus velit non pretium. Aliquam hendrerit at orci id bibendum. Praesent gravida sapien eu eleifend faucibus.

Sed at tristique augue. Proin non venenatis lacus. Sed sodales ornare urna, id vulputate dui dictum in. Duis blandit, ex in ornare vulputate, est nibh ornare felis, vitae egestas libero leo nec enim. Fusce non arcu id eros suscipit feugiat. Nullam dapibus nibh a vestibulum maximus. Fusce convallis tortor odio, vitae egestas quam lacinia vitae. Aliquam laoreet convallis eros vel ultricies. Morbi non orci nunc. Proin accumsan posuere felis vitae pharetra. Aenean suscipit congue felis, in sollicitudin metus aliquam et. Pellentesque at commodo mauris. Morbi aliquet venenatis nisl, a cursus ligula cursus et. Cras sit amet lorem ac turpis vestibulum ultrices. Praesent cursus nibh vel malesuada convallis.

### 2017-10-09 Proposal/intro.tex
Push first intro draft

@@
 \label{sec:intro}
-Vivamus et venenatis justo, sed porttitor diam.
-Sed sed lectus nunc.
-Duis cursus mi quis ornare fringilla.
-Fusce pellentesque nulla tellus, eget imperdiet enim varius vel.
-Nam viverra at augue non commodo.
-Maecenas porta lobortis erat in malesuada.
-Suspendisse pulvinar ut metus et venenatis.
-Vivamus quis tristique ligula.
-Nam vitae ex eget arcu gravida sollicitudin.
-Duis velit magna, pretium nec elit ut, placerat accumsan velit.
-Suspendisse orci nisl, imperdiet eget bibendum commodo, fermentum at libero.
-Etiam sit amet lectus hendrerit, convallis tortor bibendum, condimentum libero.
-Suspendisse sit amet elit metus.
-Nunc varius ac libero ultrices facilisis.
-Donec sit amet purus non ante efficitur ullamcorper vehicula at enim.
-Pellentesque at vulputate est, blandit porttitor felis.
+In the early years, research in network congestion control has been dominated by the traditional assumption of a decentralised, autonomous network.
+End-hosts only have control over the amount of traffic they send, and are unaware of the intentions or traffic rate of their peers.
+Similarly, switches and routers are unaware of global traffic patterns and only forward based on their local notion of optimality.
 
-Curabitur convallis, odio sed dapibus interdum, quam eros interdum leo, sed pretium urna tortor eu sem.
-Pellentesque id interdum nulla.
-Sed dapibus vel massa quis accumsan.
-Nulla vitae vehicula odio.
-Nam hendrerit ultricies elit sed convallis.
-Fusce efficitur pellentesque massa, at ultrices nulla semper et.
-Suspendisse in enim at justo ultrices ultrices.
-Quisque congue euismod odio, non vehicula elit condimentum vel.
-Vivamus vestibulum tellus ac libero facilisis molestie.
-Aliquam accumsan sit amet augue eget interdum.
-Aliquam feugiat tincidunt velit, sed tincidunt purus auctor a.
-Vestibulum nunc enim, dignissim vitae urna varius, sagittis lobortis purus.
-Cras eget quam non ipsum tincidunt pretium at eu est.
+In line of these assumptions, TCP has been designed to optimize traffic globally on a simple fairness principle.
+Nodes react to packet loss and assume that others will do the same.
+An equilibrium is reached when all end-hosts achieved a traffic rate that, in sum, conforms to the maximum available bandwidth of the congested link.
+
+TCP works excellently in scenarios where many different distrustful participants compete for limited bandwidth.
+Still, TCP is a reactive protocol.
+The fact that packet loss and latency increase occur in the network, already indicates a problem.
+Packet loss is largely caused by an overflow of queues ("bufferbloat") in forwarding elements, implying that traffic has not been optimally distributed.
+
+Ideally, a network should always be "zero-queue", i.e., latency will merely be caused by propagation, and not queuing delay.
+Queueing has generally not been a dominant issue in wide-area and enterprise networks, as traffic is sufficiently distributed and diverse, with only few "hot" target hosts.
+Optimal traffic optimization is a substantial challenge, if not impossible, as network operators have no control over the individual network elements and its participants.
+Under these conditions TCP and its extensions can be considered a best-effort solution.
+
+However, these assumptions have changed in the past decade due to new developments. Datacenters have emerged as a driving force in networking research, as they pose new design challenges.
+
+They are driven by minimization of costs and maximization of compute power and are intended to run at maximum utilization to achieve an optimal compute/cost ratio.
+In such a scenario, inefficient routing can lead to a quick deterioration and collapse of a high-load network, requiring more sophisticated approaches to solve congestion control.
+However, operators now also have the ability to freely control and adapt their network architecture, leading to highly customized systems and fine-grained control.
+
+Much in line with the trend of datacenters, Software-Defined Networking (SDN) emerged as a new networking paradigm.
+Moving away from the principle of distributed communication and routing, SDN introduced the notion of "centralized management".
+A single controller with global knowledge is able to automatically modify and adapt the forwarding tables of all switches in the network, as well as notify end hosts of changes in the network.
+These two new trends in systems allowed for impactful new innovation in the space of TCP congestion research.
+Traffic can now be managed centralised based on global knowledge of the entire topology and traffic patterns.
+In recent years, a new line of centralised schedulers has emerged, which make use of these advantages and achieve close to optimal bandwidth optimization.{CITATIONS}
+
+However, these schedulers are fundamentally still reactive.
+The central controller responds to changes in the network or requests by applications, which may cost valuable roundtrip latency.
+Often, short-term flows or bursts are unaccounted for which causes undesirable packet loss and back-propagating congestion.
+
+A much more desirable solution is a global, centralised arbiter which is able to predict and fairly distribute flows in the network before bursts or congestion occurs.
+By treating the network’s compute and forwarding power as a single finite resource, a controller could act like the OS scheduler distributing CPU time slices to processes.
+This design approach also follows SDN's aspiration of introducing operating systems abstractions to the networking domain space.
+
+In this project, we plan to explore the possibilities of a centralised, proactive flow scheduler. We ask ourselves the following research questions:
+\begin{enumerate}
+\item Is it possible to design a centralised token-based scheduling network?
+\item Is it possible to predict traffic and pre-emptively schedule flows and token distribution in a datacenter context?
+\item Using this approach, are we able to achieve better performance and utilization than existing solutions?
+\end{enumerate}
+
+In the scope of this course, we attempt to answer question 1 and design a simple token-based scheduler in Mininet.
+If we succeed, we will benchmark our results and evaluate the level of utilization compared to contemporary scheduling systems.

### 2017-10-09 Proposal/design.tex
Quick fixes

--- 2017-10-06 Proposal/relatedworks.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}
-\label{sec:related}
-Nulla facilisi.
-Donec congue, quam sed faucibus varius, libero dolor mollis ipsum, vel placerat felis urna quis nibh.
-Morbi tempor ipsum nec urna porttitor interdum.
-Phasellus cursus nisi nec tempor blandit.
-Curabitur at sem convallis, sagittis purus ornare, sagittis urna.
-Nullam eget eleifend orci.
-Suspendisse vel justo at elit mollis porta.
-Nulla congue sapien nibh, tincidunt pulvinar orci elementum in.
-Aliquam mollis tempus velit non pretium.
-Aliquam hendrerit at orci id bibendum.
-Praesent gravida sapien eu eleifend faucibus.
-
-Sed at tristique augue.
-Proin non venenatis lacus.
-Sed sodales ornare urna, id vulputate dui dictum in.
-Duis blandit, ex in ornare vulputate, est nibh ornare felis, vitae egestas libero leo nec enim.
-Fusce non arcu id eros suscipit feugiat.
-Nullam dapibus nibh a vestibulum maximus.
-Fusce convallis tortor odio, vitae egestas quam lacinia vitae.
-Aliquam laoreet convallis eros vel ultricies.
-Morbi non orci nunc.
-Proin accumsan posuere felis vitae pharetra.
-Aenean suscipit congue felis, in sollicitudin metus aliquam et.
-Pellentesque at commodo mauris.
-Morbi aliquet venenatis nisl, a cursus ligula cursus et.
-Cras sit amet lorem ac turpis vestibulum ultrices.
-Praesent cursus nibh vel malesuada convallis.
+\section{Design}
+\label{sec:design}

### 2017-10-09 Proposal/eval.tex
Quick fixes

@@
 \label{sec:evaluation}
-Maecenas aliquet sagittis nunc, ac sagittis justo scelerisque in.
-Maecenas at libero leo.
-Integer maximus gravida ante, accumsan tempor dui.
-Quisque accumsan semper erat, at mattis ex vestibulum at.
-Proin eu nulla vitae ex molestie facilisis eu id dolor.
-Phasellus lobortis efficitur purus, sit amet laoreet lacus sodales vulputate.
-Maecenas venenatis est a justo blandit consequat.
-Donec vitae pharetra orci.
-Duis tortor diam, interdum sed ex non, ultricies suscipit lorem.
-Pellentesque eros arcu, dignissim ut leo sit amet, sodales sodales tortor.
-Ut convallis, elit nec auctor mollis, erat neque feugiat ex, bibendum pharetra nulla risus sed neque.
-Aliquam feugiat elit metus, non convallis nisl malesuada pharetra.
-Donec ligula neque, blandit eget ligula placerat, lacinia finibus odio.
-Quisque sit amet suscipit diam.
-
-Nunc consectetur turpis non sollicitudin pellentesque.
-Nullam nec eleifend lectus.
-Nunc sem enim, bibendum quis elit nec, laoreet posuere dui.
-Vivamus sodales nisi sed facilisis consectetur.
-Donec venenatis justo sed consectetur ullamcorper.
-Nulla dignissim magna eget arcu consectetur aliquet.
-Praesent consequat, elit quis auctor egestas, dolor mi pretium erat, eu ullamcorper risus velit sed metus.
-Praesent lectus urna, tempor sed imperdiet id, condimentum id nunc.
-Aliquam id mi sollicitudin, venenatis justo id, congue diam.
-Sed quis dictum odio.
-Donec auctor enim quis ultricies consequat.
-Cras maximus a velit luctus feugiat.

### 2017-10-09 Proposal/implementation.tex
Quick fixes

@@
 \label{sec:implementation}
-Nunc ornare tortor at finibus faucibus.
-Sed bibendum efficitur nibh ac ultricies.
-Donec maximus purus lectus, id auctor diam molestie vel.
-Pellentesque ex odio, ultricies vel magna et, laoreet placerat augue.
-Suspendisse nunc ipsum, semper rhoncus sem at, dignissim mollis mauris.
-Cras viverra, eros eget faucibus porttitor, quam ex aliquet mi, eget sodales elit dolor in mi.
-Aenean congue augue id nunc blandit, ac molestie augue sagittis.
-Duis rhoncus odio eu arcu posuere, eget luctus diam lobortis.
-Aliquam sed ullamcorper velit.
-Suspendisse sed ligula sed risus rhoncus placerat ut fermentum nulla.
-Nam facilisis dui metus, non malesuada lacus feugiat quis.
-Vivamus volutpat non nibh eu gravida.
-Suspendisse efficitur leo maximus tortor interdum maximus.
-Nulla feugiat elit nec tincidunt tincidunt.
-In fermentum lacus tellus, sit amet tempor leo pulvinar quis.
-
-Duis tortor mauris, dignissim vehicula ligula non, lobortis finibus odio.
-Curabitur eu lobortis erat.
-Donec arcu ex, cursus in finibus sit amet, commodo ac mi.
-Integer mattis iaculis ante placerat tempus.
-Phasellus eu diam blandit, ultricies mauris ut, viverra odio.
-Praesent accumsan, lectus ut accumsan semper, sem elit pharetra enim, at vehicula felis ante a tortor.
-Duis at tellus id enim mollis posuere et vitae nisi.
-Nam convallis vulputate commodo.
-Vestibulum eget libero sed eros dictum facilisis.
-Praesent posuere vulputate ipsum non sagittis.

### 2017-10-09 Proposal/intro.tex
Quick fixes

@@
 \label{sec:intro}
-In the early years, research in network congestion control has been dominated by the traditional assumption of a decentralised, autonomous network.
+
+In the early years, research in network congestion has been dominated by the traditional assumption of a decentralised, autonomous network.
 End-hosts only have control over the amount of traffic they send, and are unaware of the intentions or traffic rate of their peers.
@@
 In line of these assumptions, TCP has been designed to optimize traffic globally on a simple fairness principle.
-Nodes react to packet loss and assume that others will do the same.
+Nodes react to packet loss and assume that others will behave accordingly.
 An equilibrium is reached when all end-hosts achieved a traffic rate that, in sum, conforms to the maximum available bandwidth of the congested link.
-
 TCP works excellently in scenarios where many different distrustful participants compete for limited bandwidth.
@@
 The fact that packet loss and latency increase occur in the network, already indicates a problem.
-Packet loss is largely caused by an overflow of queues ("bufferbloat") in forwarding elements, implying that traffic has not been optimally distributed.
+Packet loss is largely created by an overflow of queues in forwarding elements, implying that traffic has not been optimally distributed.
 
-Ideally, a network should always be "zero-queue", i.e., latency will merely be caused by propagation, and not queuing delay.
+Ideally, a network should always be "zero-queue", i.e., latency will merely be induced by propagation, and not queuing delay. 
 Queueing has generally not been a dominant issue in wide-area and enterprise networks, as traffic is sufficiently distributed and diverse, with only few "hot" target hosts.
@@
 
-However, these assumptions have changed in the past decade due to new developments. Datacenters have emerged as a driving force in networking research, as they pose new design challenges.
-
-They are driven by minimization of costs and maximization of compute power and are intended to run at maximum utilization to achieve an optimal compute/cost ratio.
-In such a scenario, inefficient routing can lead to a quick deterioration and collapse of a high-load network, requiring more sophisticated approaches to solve congestion control.
-However, operators now also have the ability to freely control and adapt their network architecture, leading to highly customized systems and fine-grained control.
+However, new developments in the past decade have changed the general networking environment.
+Datacenters have emerged as a guiding force of research, posing new design challenges as well as opportunities.
+Driven by minimization of costs and maximization of compute power they are intended to run at maximum utilization to achieve an optimal compute/cost ratio.
+In such a scenario, inefficient routing can quickly lead to bufferbloat and the eventual collapse of a high-load network, requiring more sophisticated approaches to solve congestion control.
+On the other hand, operators now have the ability to freely control and adapt their network architecture. leading highly customized systems and fine-grained control.
 
@@
 A single controller with global knowledge is able to automatically modify and adapt the forwarding tables of all switches in the network, as well as notify end hosts of changes in the network.
-These two new trends in systems allowed for impactful new innovation in the space of TCP congestion research.
+These two new trends in systems facilitated impactful new innovation opportunities in the space of TCP congestion research.
 Traffic can now be managed centralised based on global knowledge of the entire topology and traffic patterns.
  In recent years, a new line of centralised schedulers has emerged, which make use of these advantages and achieve close to optimal bandwidth optimization.{CITATIONS}
-
-However, these schedulers are fundamentally still reactive.
+However, these schedulers are still reactive  in their nature.
 The central controller responds to changes in the network or requests by applications, which may cost valuable roundtrip latency.
-Often, short-term flows or bursts are unaccounted for which causes undesirable packet loss and back-propagating congestion.
+Often, short-term flows or bursts are unaccounted for, which causes undesirable packet loss and backpropagating congestion.
 
@@
 By treating the network’s compute and forwarding power as a single finite resource, a controller could act like the OS scheduler distributing CPU time slices to processes.
-This design approach also follows SDN's aspiration of introducing operating systems abstractions to the networking domain space.
+This design approach also follows SDN’s aspiration of introducing operating systems abstractions to the networking domain space.
 
@@
 \item Is it possible to design a centralised token-based scheduling network?
-\item Is it possible to predict traffic and pre-emptively schedule flows and token distribution in a datacenter context?
+\item Is it possible to predict traffic and preemptively schedule flows and token distribution in a datacenter context?
 \item Using this approach, are we able to achieve better performance and utilization than existing solutions?

### 2017-10-09 Proposal/paper.tex
Push first intro draft
Quick fixes

@@
 
-\author{??? ???}
-\affiliation{%
-  \institution{University of British Columbia}
-}
-\email{pzarei1993@gmail.com}
+\begin{abstract}
 
-\begin{abstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum lacinia arcu ac diam mattis pharetra.
-Curabitur luctus posuere dictum.
-In quam tortor, cursus luctus tortor ut, porta fermentum lacus.
-Nunc eget augue dui.
-Proin ante lectus, malesuada a sem ut, lobortis eleifend turpis.
-Morbi sit amet odio interdum, sollicitudin sem eget, mollis orci.
-Curabitur rhoncus arcu magna, eget rhoncus nunc porttitor at.
-Nulla quis posuere urna.
-Nam enim nunc, hendrerit et gravida et, mollis a urna.
-Donec aliquam massa sollicitudin tortor cursus placerat.
-Ut non nisi purus.
-Etiam laoreet ut libero et tempus.
 \end{abstract}
@@
 \input{intro}
-% \input{background}
+
 \input{relatedworks}
-\input{motivation}
+\input{design}
+%\input{motivation}
 \input{implementation}

### 2017-10-09 Proposal/relatedworks.tex
Quick fixes

@@
 \label{sec:related}
-Nulla facilisi.
-Donec congue, quam sed faucibus varius, libero dolor mollis ipsum, vel placerat felis urna quis nibh.
-Morbi tempor ipsum nec urna porttitor interdum.
-Phasellus cursus nisi nec tempor blandit.
-Curabitur at sem convallis, sagittis purus ornare, sagittis urna.
-Nullam eget eleifend orci.
-Suspendisse vel justo at elit mollis porta.
-Nulla congue sapien nibh, tincidunt pulvinar orci elementum in.
-Aliquam mollis tempus velit non pretium.
-Aliquam hendrerit at orci id bibendum.
-Praesent gravida sapien eu eleifend faucibus.
-
-Sed at tristique augue.
-Proin non venenatis lacus.
-Sed sodales ornare urna, id vulputate dui dictum in.
-Duis blandit, ex in ornare vulputate, est nibh ornare felis, vitae egestas libero leo nec enim.
-Fusce non arcu id eros suscipit feugiat.
-Nullam dapibus nibh a vestibulum maximus.
-Fusce convallis tortor odio, vitae egestas quam lacinia vitae.
-Aliquam laoreet convallis eros vel ultricies.
-Morbi non orci nunc.
-Proin accumsan posuere felis vitae pharetra.
-Aenean suscipit congue felis, in sollicitudin metus aliquam et.
-Pellentesque at commodo mauris.
-Morbi aliquet venenatis nisl, a cursus ligula cursus et.
-Cras sit amet lorem ac turpis vestibulum ultrices.
-Praesent cursus nibh vel malesuada convallis.

### 2017-10-10 Proposal/design.tex
Add the library.

@@
 \label{sec:design}
+
+Simple System
+In our initial simple TCPToken design, a centralized controller regulates all node traffic by provisioning end-hosts with tokens.
+These tokens act as the “currency” of the system.
+The total amount is fixed to the aggregate bandwidth available in the network and the traffic window size of sending end hosts is calculated based on the current token availability.
+If a node or switch is overburdened, it may notify the scheduler, which will adjust the traffic window of responsible sending nodes remotely.
+Any management or control operations such as token distribution and updates are priority-queued to guarantee a fast and low-latency response.
+In the case of an end-host requiring more bandwidth it can also notify the controller, which may comply based on priority and availability.
+Initially, the controller will compute optimal route configuration based on the topology and link bandwidth using a simple heuristic bin-packing approach.
+End-hosts will be initialized with a fixed low-to-medium bandwidth guarantee, which will be adjusted over time.
+
+Advanced System
+The aforementioned system is intended only as a proof of concept.
+It is still a  synchronous and reactive approach, dependent on notifications and requests by overburdened switches and underprovisioned hosts.
+Ideally, all management in the TCPToken system should be asynchronous and preemptive. 
+1) Paying for traffic
+In a complete design, tokens will be used as a transport pass instead of simply acting as traffic shaping parameters.
+On a per-flow basis, nodes will be able to attach tokens to traffic as a form of payment.
+Enforcement is performed on the switch level, any flow that is not “paid” by a token will be dropped.
+Nodes may be handed different types of tokens to serve load-balancing and traffic shaping purposes.
+Applications operating on traditionally short-lived flows and bursts may sign their traffic with a different token type than long-lived permanent processes.
+Switches will forward traffic and balance flows accordingly with only minor interference from the central controller.
+
+2) Predicting traffic
+ML-based scheduling is an emerging research field, also known as Knowledge-Defined Networking. Datacenter traffic follows patterns, which can be predicted using statistical methods.
+Combined with using tokens as a traffic engineering enforcement, a preemptive scheduling strategy in datacenters may be feasible.
+
+We do not intend to implement the advanced system for this project, but are using it as a guideline to motivate our research.
+Based on the experiences and knowledge gained from implementing the basic system, we will reevaluate the feasibility and practicality of the second concept.
+
+Fault-Tolerance and Scalability
+To handle fault-tolerance we plan to integrate a fallback-mechanism to conventional TCP, if the node has not received a TCPToken in a prefixed amount of time.
+End-hosts will only be marginally affected by a controller or link failure, as they are guaranteed a fixed amount of bandwidth based on their token availability.
+To avoid overburdening the controller, we envision a distributed hive of schedulers each managing their own region in the final design.
+Add ONOS CITATION In case of failure, a shadow node or neighbouring device may take over computation temporarily, until the main node has recovered.
+CITATION

### 2017-10-10 Proposal/implementation.tex
Add the library.

@@
 \label{sec:implementation}
+We intend to emulate our system in Mininet to observe traffic patterns and infer a suitable token algorithm.
+Mininet has proven itself to be a viable tool to model new congestion control algorithms CITATION, and will help us prototype our concept efficiently.
+We will build a custom SDN controller that interacts with traditional OpenFlow software switches as well as end-hosts.
+End-hosts will run a custom real-world traffic generation script that adjust based on information packets send by the controller.
+If time permits, we may expand our implementation to MaxiNet, which can emulate large-scale network stress tests of thousands of nodes on multiple physical hosts.
+
+We initially considered a second implementation alternative in C/C++ based on the FastPass  CITATION source code.
+This would provide us with a fully deployable system which we could fork our implementation from.
+A major advantage of this approach is the ability to test scenarios and traffic algorithms using real software code.
+However, several concerns made us favour a Mininet emulation instead.
+First, FastPass relies on DPDK integration, which requires actual hardware interfaces.
+The central arbiter in the FastPass design would need to run on a dedicated machine, which increases prototyping and development complexity substantially.
+Secondly, the FastPass code is highly specialized and optimized research code with only little available documentation.
+Modifying and evolving the source code will require thorough understanding of kernel and networking development, a significant time-sink.
+For a class project, these may be major initial hurdles, taking away from the research aspect of the design concept.
+Consequently, we have decided to pursue an approach which allows us to quickly develop an understanding of the problem without being obstructed by engineering work.

### 2017-10-10 Proposal/intro.tex
Add the library.

@@
 
-In the early years, research in network congestion has been dominated by the traditional assumption of a decentralised, autonomous network.
-End-hosts only have control over the amount of traffic they send, and are unaware of the intentions or traffic rate of their peers.
+Initially, research in network congestion  was dominated by the  assuming a decentralised, autonomous network.
+End-hosts only have control over the amount of traffic they send, and are unaware of the intentions or traffic rates of their peers.
 Similarly, switches and routers are unaware of global traffic patterns and only forward based on their local notion of optimality.
 
-In line of these assumptions, TCP has been designed to optimize traffic globally on a simple fairness principle.
-Nodes react to packet loss and assume that others will behave accordingly.
-An equilibrium is reached when all end-hosts achieved a traffic rate that, in sum, conforms to the maximum available bandwidth of the congested link.
-TCP works excellently in scenarios where many different distrustful participants compete for limited bandwidth.
-Still, TCP is a reactive protocol.
-The fact that packet loss and latency increase occur in the network, already indicates a problem.
-Packet loss is largely created by an overflow of queues in forwarding elements, implying that traffic has not been optimally distributed.
+In line with these assumptions, TCP was designed to optimize traffic globally on a simple fairness principle; nodes react to packet loss and assume that others will behave accordingly.
+An equilibrium is reached when all end-hosts achieve a traffic rate that, in sum, conforms to the maximum available bandwidth of the most congested link.
+TCP works well  in scenarios where many different distrustful participants compete for limited bandwidth.
+Still, TCP is a \textit{reactive protocol}; the fact that packet loss and latency increases occur in the network already indicates a problem.
+Packet loss is primarily  due to overflowing  queues in forwarding elements, implying that traffic has not been optimally distributed.
+Ideally, a network should always be “zero-queue”, i.e., latency will merely be induced by propagation, and not queuing delay.
 
-Ideally, a network should always be "zero-queue", i.e., latency will merely be induced by propagation, and not queuing delay. 
-Queueing has generally not been a dominant issue in wide-area and enterprise networks, as traffic is sufficiently distributed and diverse, with only few "hot" target hosts.
-Optimal traffic optimization is a substantial challenge, if not impossible, as network operators have no control over the individual network elements and its participants.
-Under these conditions TCP and its extensions can be considered a best-effort solution.
+Queueing has generally not been a dominating issue in wide-area and enterprise networks, as traffic is sufficiently distributed and diverse, with only few “hot” target hosts.~\cite{hedera, microte} Traffic optimization is a substantial challenge; network operators have no control over the individual network elements nor its participants.
+Under these conditions, TCP and its extensions can be considered a best-effort solution.
 
-However, new developments in the past decade have changed the general networking environment.
-Datacenters have emerged as a guiding force of research, posing new design challenges as well as opportunities.
-Driven by minimization of costs and maximization of compute power they are intended to run at maximum utilization to achieve an optimal compute/cost ratio.
-In such a scenario, inefficient routing can quickly lead to bufferbloat and the eventual collapse of a high-load network, requiring more sophisticated approaches to solve congestion control.
-On the other hand, operators now have the ability to freely control and adapt their network architecture. leading highly customized systems and fine-grained control.
-
-Much in line with the trend of datacenters, Software-Defined Networking (SDN) emerged as a new networking paradigm.
-Moving away from the principle of distributed communication and routing, SDN introduced the notion of "centralized management".
+Developments in the past decade have changed the general networking environment. Datacenters have emerged as an exciting new research frontier, posing novel  design challenges and opportunities.
+Driven by minimization of costs and maximization of compute power;  data centers must  run at maximum utilization to achieve an optimal compute/cost ratio.
+Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load network, requiring more sophisticated approaches to solve congestion control.
+On the other hand, operators now have the ability to freely control and adapt their network architecture leading highly customized systems and fine-grained optimization.
+As a result  Software-Defined Networking (SDN) emerged as a new networking paradigm.
+Moving away from the principle of distributed communication and routing, SDN introduces the notion of “centralized management”.
 A single controller with global knowledge is able to automatically modify and adapt the forwarding tables of all switches in the network, as well as notify end hosts of changes in the network.
 These two new trends in systems facilitated impactful new innovation opportunities in the space of TCP congestion research.
-Traffic can now be managed centralised based on global knowledge of the entire topology and traffic patterns.
- In recent years, a new line of centralised schedulers has emerged, which make use of these advantages and achieve close to optimal bandwidth optimization.{CITATIONS}
-However, these schedulers are still reactive  in their nature.
+Traffic can now be managed in a  \textit{centralised} fashion based on \textit{global knowledge} of the entire topology and traffic patterns.
+A new line of centralised schedulers has emerged that  can achieve close to optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
+However, these schedulers are still  \textit{reactive}  in nature.
 The central controller responds to changes in the network or requests by applications, which may cost valuable roundtrip latency.

### 2017-10-10 Proposal/implementation.tex
Push near final draft for tonight

@@
 \label{sec:implementation}
-We intend to emulate our system in Mininet to observe traffic patterns and infer a suitable token algorithm.
-Mininet has proven itself to be a viable tool to model new congestion control algorithms CITATION, and will help us prototype our concept efficiently.
+We intend to emulate our system in Mininet~\cite{mininet} to observe traffic patterns and infer a suitable token algorithm.
+Mininet has proven itself to be a viable tool to model new congestion control algorithms~\cite{mininet_learning}, and will help us prototype our concept efficiently.
 We will build a custom SDN controller that interacts with traditional OpenFlow software switches as well as end-hosts.
-End-hosts will run a custom real-world traffic generation script that adjust based on information packets send by the controller.
+End-hosts will run a custom real-world traffic generation script which adjusts based on information packets sent by the controller.
 If time permits, we may expand our implementation to MaxiNet, which can emulate large-scale network stress tests of thousands of nodes on multiple physical hosts.
 
-We initially considered a second implementation alternative in C/C++ based on the FastPass  CITATION source code.
+We initially considered a second implementation alternative in C/C++ based on the FastPass~\cite{fastpass} source code.
 This would provide us with a fully deployable system which we could fork our implementation from.
@@
 However, several concerns made us favour a Mininet emulation instead.
-First, FastPass relies on DPDK integration, which requires actual hardware interfaces.
+Firstly, FastPass relies on DPDK integration, which requires actual hardware interfaces.
 The central arbiter in the FastPass design would need to run on a dedicated machine, which increases prototyping and development complexity substantially.

### 2017-10-10 Proposal/intro.tex
Push near final draft for tonight

@@
 \label{sec:intro}
-
-Initially, research in network congestion  was dominated by the  assuming a decentralised, autonomous network.
-End-hosts only have control over the amount of traffic they send, and are unaware of the intentions or traffic rates of their peers.
+Initially, research in network congestion was dominated by the assumption of a decentralised, autonomous network - end-hosts only have control over the amount of traffic they send, and are unaware of the intentions or traffic rates of their peers.
 Similarly, switches and routers are unaware of global traffic patterns and only forward based on their local notion of optimality.
 
-In line with these assumptions, TCP was designed to optimize traffic globally on a simple fairness principle; nodes react to packet loss and assume that others will behave accordingly.
+In line with these assumptions, TCP was designed to optimize traffic globally on a simple fairness principle: nodes react to packet loss and assume that others will behave accordingly.
 An equilibrium is reached when all end-hosts achieve a traffic rate that, in sum, conforms to the maximum available bandwidth of the most congested link.
@@
 Packet loss is primarily due to overflowing queues in forwarding elements, implying that traffic has not been optimally distributed.
-Ideally, a network should always be “zero-queue”, i.e., latency will merely be induced by propagation, and not queuing delay.
+Ideally, a network should always be "zero-queue", i.e., latency will merely be induced by propagation, and not queuing delay.
 
-Queueing has generally not been a dominating issue in wide-area and enterprise networks, as traffic is sufficiently distributed and diverse, with only few “hot” target hosts.~\cite{hedera, microte} Traffic optimization is a substantial challenge; network operators have no control over the individual network elements nor its participants.
+Queueing has generally not been a dominating issue in wide-area and enterprise networks, as traffic is sufficiently distributed and diverse, with only few "hot" target hosts.~\cite{hedera, microte} Traffic optimization is a substantial challenge; network operators have no control over the individual network elements nor its participants.
 Under these conditions, TCP and its extensions can be considered a best-effort solution.
@@
 Developments in the past decade have changed the general networking environment. Datacenters have emerged as an exciting new research frontier, posing novel  design challenges and opportunities.
-Driven by minimization of costs and maximization of compute power;  data centers must  run at maximum utilization to achieve an optimal compute/cost ratio.
+Driven by minimization of costs and maximization of compute power, data centers must  run at maximum utilization to achieve an optimal compute/cost ratio.
 Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load network, requiring more sophisticated approaches to solve congestion control.
-On the other hand, operators now have the ability to freely control and adapt their network architecture leading highly customized systems and fine-grained optimization.
-As a result  Software-Defined Networking (SDN) emerged as a new networking paradigm.
-Moving away from the principle of distributed communication and routing, SDN introduces the notion of “centralized management”.
-A single controller with global knowledge is able to automatically modify and adapt the forwarding tables of all switches in the network, as well as notify end hosts of changes in the network.
-These two new trends in systems facilitated impactful new innovation opportunities in the space of TCP congestion research.
+On the other hand, operators now have the ability to freely control and adapt their network architecture, leading to highly customized systems and fine-grained optimization.
+As a result , Software-Defined Networking (SDN) has emerged as a new networking paradigm.
+Moving away from the principle of distributed communication and routing, SDN introduces the notion of "centralized management".
+A single controller with global knowledge is able to automatically modify and adapt the forwarding tables of all switches in the network, while notifying end hosts of changes in the network.
+These two new trends in system design facilitated impactful new innovation opportunities in the space of TCP congestion research.
 Traffic can now be managed in a  \textit{centralised} fashion based on \textit{global knowledge} of the entire topology and traffic patterns.
@@
 A much more desirable solution is a global, centralised arbiter which is able to predict and fairly distribute flows in the network before bursts or congestion occurs.
-By treating the network’s compute and forwarding power as a single finite resource, a controller could act like the OS scheduler distributing CPU time slices to processes.
-This design approach also follows SDN’s aspiration of introducing operating systems abstractions to the networking domain space.
+By treating the network’s compute and forwarding power as a single finite resource, a controller acts like the OS scheduler distributing CPU time slices to processes.
+This design approach follows SDN’s aspiration of introducing operating systems abstractions to the networking domain space.
 
@@
 In the scope of this course, we attempt to answer question 1 and design a simple token-based scheduler in Mininet.
-If we succeed, we will benchmark our results and evaluate the level of utilization compared to contemporary scheduling systems.
+If successful, we will benchmark our results and evaluate the level of utilization compared to contemporary scheduling systems.

### 2017-10-10 Proposal/paper.tex
Push near final draft for tonight

@@
 \begin{abstract}
-
+Datacenter networks have become a hotbed for research in the recent past.
+Many works have been published on improving the bisection bandwidth, reducing cost, and improving congestion control as well as latency.
+As a core contribution of this research, centralized scheduling has entered the stage as dominant strategy to manage flows.
+Benefitting from a global view and full control over the network, centralized schedulers are able to enforce fine-grained traffic control, frequently achieving near-optimal bandwidth optimization.
+However, modern congestion control algorithms and schedulers are still fundamentally reactive.
+Many algorithms are designed to only respond to packet loss or significant increase in latency, not to actively prevent it.
+In this project, we plan to explore the opportunities of employing a preemptive and tightly controlling central network scheduler.
+Using tokens as a global tool to enforce traffic limiting and policies, the scheduler is able to proactively control the flow of traffic as well as respond to application requests.
+"Knowledge" and predictive analysis in networks is a growing trend in research which we intend to leverage in this system.
+We investigate the potential in such a new form of interactive congestion control and analyze it against to state-of-the-art solutions.
+Our tool of choice to model our system is Mininet, a rapid prototyping emulator for datacenter networks.
+We will compare our "TCPToken" system against established TCP-congestion systems such as Hedera and DCTCP.
+Based on the findings and measurement results, we will reassess the feasibility and prospect of a token-based, predictive congestion control algorithm.
 \end{abstract}
@@
 \input{eval}
+\input{timeline}
 

### 2017-10-10 Proposal/intro.tex
More changes

@@
 \label{sec:intro}
+
 Initially, research in network congestion was dominated by the assumption of a decentralised, autonomous network - end-hosts only have control over the amount of traffic they send, and are unaware of the intentions or traffic rates of their peers.
@@
 Packet loss is primarily due to overflowing queues in forwarding elements, implying that traffic has not been optimally distributed.
-Ideally, a network should always be "zero-queue", i.e., latency will merely be induced by propagation, and not queuing delay.
+Ideally, a network should always be “zero-queue”, i.e., latency will merely be induced by propagation, and not queuing delay.
 
-Queueing has generally not been a dominating issue in wide-area and enterprise networks, as traffic is sufficiently distributed and diverse, with only few "hot" target hosts.~\cite{hedera, microte} Traffic optimization is a substantial challenge; network operators have no control over the individual network elements nor its participants.
+Queueing has generally not been a dominating issue in wide-area and enterprise networks, as traffic is sufficiently distributed and diverse, with only few “hot” target hosts.~\cite{hedera, microte} Traffic optimization is a substantial challenge; network operators have no control over the individual network elements nor its participants.
 Under these conditions, TCP and its extensions can be considered a best-effort solution.
@@
 Developments in the past decade have changed the general networking environment. Datacenters have emerged as an exciting new research frontier, posing novel design challenges and opportunities.
+
 Driven by minimization of costs and maximization of compute power, data centers must  run at maximum utilization to achieve an optimal compute/cost ratio.
@@
 As a result , Software-Defined Networking (SDN) has emerged as a new networking paradigm.
-Moving away from the principle of distributed communication and routing, SDN introduces the notion of "centralized management".
+Moving away from the principle of distributed communication and routing, SDN introduces the notion of “centralized management”.
 A single controller with global knowledge is able to automatically modify and adapt the forwarding tables of all switches in the network, while notifying end hosts of changes in the network.
@@
 Traffic can now be managed in a  \textit{centralised} fashion based on \textit{global knowledge} of the entire topology and traffic patterns.
+
 A new line of centralised schedulers has emerged that  can achieve close to optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
@@
 
-In the scope of this course, we attempt to answer question 1 and design a simple token-based scheduler in Mininet.
+In the scope of this course, we attempt to answer question one and design a simple token-based scheduler in Mininet.
 If successful, we will benchmark our results and evaluate the level of utilization compared to contemporary scheduling systems.

### 2017-10-10 Proposal/paper.tex
More changes

@@
 Many algorithms are designed to only respond to packet loss or significant increase in latency, not to actively prevent it.
+
 In this project, we plan to explore the opportunities of employing a preemptive and tightly controlling central network scheduler.
@@
 "Knowledge" and predictive analysis in networks is a growing trend in research which we intend to leverage in this system.
+
 We investigate the potential in such a new form of interactive congestion control and analyze it against to state-of-the-art solutions.
@@
 
-\input{intro}
+%\input{intro}
+%
+%\input{relatedworks}
+%\input{design}
+% %\input{motivation}
+%\input{implementation}
+%\input{eval}
+%\input{timeline}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
+\label{sec:intro}
 
-\input{relatedworks}
-\input{design}
-%\input{motivation}
-\input{implementation}
-\input{eval}
-\input{timeline}
+Initially, research in network congestion was dominated by the assumption of a decentralised, autonomous network - end-hosts only have control over the amount of traffic they send, and are unaware of the intentions or traffic rates of their peers.
+Similarly, switches and routers are unaware of global traffic patterns and only forward based on their local notion of optimality.
+
+In line with these assumptions, TCP was designed to optimize traffic globally on a simple fairness principle: nodes react to packet loss and assume that others will behave accordingly.
+An equilibrium is reached when all end-hosts achieve a traffic rate that, in sum, conforms to the maximum available bandwidth of the most congested link.
+TCP works well  in scenarios where many different distrustful participants compete for limited bandwidth.
+Still, TCP is a \textit{reactive protocol}; the fact that packet loss and latency increases occur in the network already indicates a problem.
+Packet loss is primarily due to overflowing queues in forwarding elements, implying that traffic has not been optimally distributed.
+Ideally, a network should always be "zero-queue", i.e., latency will merely be induced by propagation, and not queuing delay.
+
+Queueing has generally not been a dominating issue in wide-area and enterprise networks, as traffic is sufficiently distributed and diverse, with only few "hot" target hosts.~\cite{hedera, microte} Traffic optimization is a substantial challenge; network operators have no control over the individual network elements nor its participants.
+Under these conditions, TCP and its extensions can be considered a best-effort solution.
+
+Developments in the past decade have changed the general networking environment. Datacenters have emerged as an exciting new research frontier, posing novel design challenges and opportunities.
+
+Driven by minimization of costs and maximization of compute power, data centers must  run at maximum utilization to achieve an optimal compute/cost ratio.
+Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load network, requiring more sophisticated approaches to solve congestion control.
+On the other hand, operators now have the ability to freely control and adapt their network architecture, leading to highly customized systems and fine-grained optimization.
+As a result , Software-Defined Networking (SDN) has emerged as a new networking paradigm.
+Moving away from the principle of distributed communication and routing, SDN introduces the notion of "centralized management".
+A single controller with global knowledge is able to automatically modify and adapt the forwarding tables of all switches in the network, while notifying end hosts of changes in the network.
+These two new trends in system design facilitated impactful new innovation opportunities in the space of TCP congestion research.
+Traffic can now be managed in a  \textit{centralised} fashion based on \textit{global knowledge} of the entire topology and traffic patterns.
+
+A new line of centralised schedulers has emerged that  can achieve close to optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
+However, these schedulers are still  \textit{reactive}  in nature.
+The central controller responds to changes in the network or requests by applications, which may cost valuable roundtrip latency.
+Often, short-term flows or bursts are unaccounted for, which causes undesirable packet loss and backpropagating congestion.
+
+A much more desirable solution is a global, centralised arbiter which is able to predict and fairly distribute flows in the network before bursts or congestion occurs.
+By treating the network’s compute and forwarding power as a single finite resource, a controller acts like the OS scheduler distributing CPU time slices to processes.
+This design approach follows SDN’s aspiration of introducing operating systems abstractions to the networking domain space.
+
+In this project, we plan to explore the possibilities of a centralised, proactive flow scheduler. We ask ourselves the following research questions:
+\begin{enumerate}
+\item Is it possible to design a centralised token-based scheduling network?
+\item Is it possible to predict traffic and preemptively schedule flows and token distribution in a datacenter context?
+\item Using this approach, are we able to achieve better performance and utilization than existing solutions?
+\end{enumerate}
+
+In the scope of this course, we attempt to answer question one and design a simple token-based scheduler in Mininet.
+If successful, we will benchmark our results and evaluate the level of utilization compared to contemporary scheduling systems.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Related Work}
+\label{sec:related}
+
+Most of the work in congestion control relies on a centralised scheduler which periodically gets updates about the flows in the network and reserves specific paths for long running flows.
+
+MicroTE~\cite{microte} does predictive OpenFlow routing and scheduling using a central SDN controller.
+Hedera~\cite{hedera} schedules large so called "elephant" flows by gathering switch statistics.
+However it pays poor attention to short term flows and network latency.
+FastPass~\cite{fastpass} on the other hand, proposes a zero-queue network and promises a very low latency network.
+The feasibility of this proposal remains questionable when considering the high cost of implementation.
+
+Recent proposals on decentralised scheduling like ExpressPass~\cite{expresspass} have heavily influenced our research.
+ExpressPass posits a credit based system similar to TCPTokens in order to control the congestion in the network.
+Effecting flow rerouting using ML techniques in KDN~\cite{kdn} and other works shows promising signs of further improvements in this area.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Design / Proposed Approach}
+\label{sec:design}
+
+\subsection{Simple System}
+In our initial simple TCPToken design, a centralized controller regulates all node traffic by provisioning end-hosts with tokens.
+These tokens act as the "currency" of the system.
+The total amount is fixed to some factor of the aggregate bandwidth available in the network, and the traffic window size of sending end hosts is calculated based on the host’s current token availability.
+If a node or switch is overburdened, it may notify the scheduler, which will adjust the traffic window of responsible sending nodes remotely.
+Management and control operations including token distribution and updates are priority-queued to guarantee a fast and low-latency response.
+If an end-host requires more bandwidth, it can also notify the controller, which may comply based on priority and availability.
+Initially, the controller will compute optimal route configuration based on the topology and link bandwidth using a simple heuristic bin-packing approach.
+End-hosts will be initialized with a fixed low-to-medium bandwidth guarantee that will be adjusted over time.
+
+\subsection{Advanced System}
+The aforementioned system is intended only as a proof of concept.
+It is still a  synchronous and reactive approach, dependent on notifications and requests by overburdened switches and underprovisioned hosts.
+Ideally, all management in the TCPToken system should be asynchronous and preemptive. 
+\subsubsection{ Paying for traffic}
+In a complete design, tokens will be used as a transport pass instead of simply acting as traffic shaping parameters.
+On a per-flow basis, nodes will be able to attach tokens to traffic as a form of payment.
+Enforcement is performed on the switch level, any flow that is not "paid" by a token will be dropped.
+Nodes may be handed different types of tokens to serve load-balancing and traffic shaping purposes.
+Applications operating on traditionally short-lived flows and bursts may sign their traffic with a different token type than long-lived permanent processes.
+Switches will forward traffic and balance flows accordingly with only minor interference from the central controller.
+
+\subsubsection{Predicting traffic}
+ML-based scheduling is an emerging research field, also known as Knowledge-Defined Networking. Datacenter traffic follows patterns, which can be predicted using statistical methods.
+Combined with using tokens as a traffic engineering enforcement, a preemptive scheduling strategy in datacenters may be feasible.
+
+We do not intend to implement the advanced system for this project, but are using it as a guideline to motivate our research.
+Based on the experiences and knowledge gained from implementing the basic system, we will reevaluate the feasibility and practicality of the second concept.
+
+\subsubsection{Fault-Tolerance and Scalability}
+To handle fault-tolerance we plan to integrate a fallback-mechanism to conventional TCP, if the node has not received a TCPToken in a prefixed amount of time.
+End-hosts will only be marginally affected by a controller or link failure, as they are guaranteed a fixed amount of bandwidth based on their token availability.
+
+To avoid overburdening the controller, we envision a distributed hive of schedulers (e.g., Onos) each managing their own region in the final design.~\cite{onos} In case of failure, a shadow node or neighbouring device may take over computation temporarily, until the main node has recovered.~\cite{disco}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Implementation}
+\label{sec:implementation}
+We intend to emulate our system in Mininet~\cite{mininet} to observe traffic patterns and infer a suitable token algorithm.
+Mininet has proven itself to be a viable tool to model new congestion control algorithms~\cite{mininet_learning}, and will help us prototype our concept efficiently.
+We will build a custom SDN controller that interacts with traditional OpenFlow software switches as well as end-hosts.
+End-hosts will run a custom real-world traffic generation script which adjusts based on information packets sent by the controller.
+If time permits, we may expand our implementation to MaxiNet, which can emulate large-scale network stress tests of thousands of nodes on multiple physical hosts.
+
+We initially considered a second implementation alternative in C/C++ based on the FastPass~\cite{fastpass} source code.
+This would provide us with a fully deployable system which we could fork our implementation from.
+A major advantage of this approach is the ability to test scenarios and traffic algorithms using real software code.
+However, several concerns made us favour a Mininet emulation instead.
+Firstly, FastPass relies on DPDK integration, which requires actual hardware interfaces.
+The central arbiter in the FastPass design would need to run on a dedicated machine, which increases prototyping and development complexity substantially.
+
+Secondly, the FastPass code is highly specialized and optimized research code with only little available documentation.
+Modifying and evolving the source code will require thorough understanding of kernel and networking development, a significant time-sink.
+For a class project, these may be major initial hurdles, taking away from the research aspect of the design concept.
+Consequently, we have decided to pursue an approach which allows us to quickly develop an understanding of the problem without being obstructed by engineering work.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Evaluation}
+\label{sec:evaluation}
+To evaluate the effectiveness of our system, we plan to measure against existing centralized as well as decentralized solutions.
+We will run datacenter traffic on a fat-tree topology of various depth and observe the effect each congestion algorithm will have on the overall bandwidth utilization and latency.
+The centralized design will be based on Hedera~\cite{hedera}, a common and influential datacenter scheduler.
+
+The decentralized congestion control mechanism will be DCTCP~\cite{dctcp}, a state-of-the-art TCP congestion algorithm.
+As baseline we will also compare to the default TCP congestion algorithm to estimate how our scheduler fares against the default case.
+As our initial simple design is very constrained in its ability to route and control traffic we expect to achieve less optimality against DCTCP and Hedera, but gain a substantial advantage over common TCP.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Timeline}
+\label{sec:timeline}
+\begin{itemize}
+\item \underline{October 15th:} Finalize draft and related work / background section.
+\item \underline{ October 22nd:} Iterate over literature and thoroughly analyze related work. Set up an initial prototyping environment in Mininet on which each team member can work on.
+\item \underline{October 30th:} Finalize the initial simple design and explore naive congestion modelling in Mininet. Integrate existing Mininet emulations of Hedera and DCTCP.
+\item \underline{November 12th:} Develop initial naive TCPTokens prototype involving controller and end-host design.
+\item \underline{November 19th:} Develop intelligent centralised congestion control heuristic.
+\item \underline{November 26th:} Run evaluations and microbenchmarks.
+\item \underline{December 3rd:} Iterate and improve the scheduler, explore topology and bandwidth constraints. Test at scale.
+\item \underline{December 10th:} Write report and prepare presentation.
+\end{itemize}
 

### 2017-10-10 Proposal/paper.tex
Final edit.

@@
 
-\title{TCPTokens: Introducing Currency into Data Center Congestion Control}
+\title{TCPTokens: Introducing Currency into Datacenter Congestion Control}
 
@@
 Datacenter networks have become a hotbed for research in the recent past.
-Many works have been published on improving the bisection bandwidth, reducing cost, and improving congestion control as well as latency.
+Many works have been published on improving the  bisection bandwidth, reducing cost, and latency.
 As a core contribution of this research, centralized scheduling has entered the stage as dominant strategy to manage flows.
@@
 
-We investigate the potential in such a new form of interactive congestion control and analyze it against to state-of-the-art solutions.
+We investigate the potential in such a new form of interactive congestion control and analyze it against state-of-the-art solutions.
 Our tool of choice to model our system is Mininet, a rapid prototyping emulator for datacenter networks.
 We will compare our "TCPToken" system against established TCP-congestion systems such as Hedera and DCTCP.
-Based on the findings and measurement results, we will reassess the feasibility and prospect of a token-based, predictive congestion control algorithm.
+Based on the findings and measurement results, we will reassess the feasibility and prospects of a token-based, predictive congestion control algorithm.
 \end{abstract}
 
-\keywords{TCP, Congestion Control, SDN, Data center}
+\keywords{TCP, Congestion Control, SDN, Datacenter}
 
@@
 %\input{timeline}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Developments in the past decade have changed the general networking environment. Datacenters have emerged as an exciting new research frontier, posing novel design challenges and opportunities.
-
-Driven by minimization of costs and maximization of compute power, data centers must  run at maximum utilization to achieve an optimal compute/cost ratio.
+Developments in the past decade have changed the general networking environment.
+Datacenters have emerged as an exciting new research frontier, posing novel design challenges and opportunities.Driven by minimization of costs and maximization of compute power, datacenters must  run at maximum utilization to achieve an optimal compute/cost ratio.
 Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load network, requiring more sophisticated approaches to solve congestion control.
 On the other hand, operators now have the ability to freely control and adapt their network architecture, leading to highly customized systems and fine-grained optimization.
-As a result , Software-Defined Networking (SDN) has emerged as a new networking paradigm.
+As a result, Software-Defined Networking (SDN) has emerged as a new networking paradigm.
+
 Moving away from the principle of distributed communication and routing, SDN introduces the notion of "centralized management".
@@
 
-A much more desirable solution is a global, centralised arbiter which is able to predict and fairly distribute flows in the network before bursts or congestion occurs.
-By treating the network’s compute and forwarding power as a single finite resource, a controller acts like the OS scheduler distributing CPU time slices to processes.
-This design approach follows SDN’s aspiration of introducing operating systems abstractions to the networking domain space.
+A much more desirable solution is a global, centralised arbiter which is able to predict and fairly distribute flows in the network before bursts or congestion occur.
+By treating the network's compute and forwarding power as a single finite resource, a controller acts like the OS scheduler distributing CPU time slices to processes.
+This design approach follows SDN's aspiration of introducing operating systems abstractions to the networking domain space.
 
@@
 These tokens act as the "currency" of the system.
-The total amount is fixed to some factor of the aggregate bandwidth available in the network, and the traffic window size of sending end hosts is calculated based on the host’s current token availability.
+The total amount is fixed to some function of the bisection and aggregate bandwidths available in the network, and the traffic window size of sending end hosts is calculated based on the host's current token availability.
 If a node or switch is overburdened, it may notify the scheduler, which will adjust the traffic window of responsible sending nodes remotely.

### 2017-10-17 Proposal/paper.tex
Final proposal.

--- 2017-10-10 Proposal/paper.tex
@@
 Datacenter networks have become a hotbed for research in the recent past.
-Many works have been published on improving the bisection bandwidth, reducing cost, and improving congestion control as well as latency.
-As a core contribution of this research, centralized scheduling has entered the stage as dominant strategy to manage flows.
+Many works have been published on improving bisection bandwidth, cost, and latency.
+As a core contribution of this research, centralized scheduling has entered the stage as a dominant strategy to manage flows.
 Benefitting from a global view and full control over the network, centralized schedulers are able to enforce fine-grained traffic control, frequently achieving near-optimal bandwidth optimization.
@@
 In this project, we plan to explore the opportunities of employing a preemptive and tightly controlling central network scheduler.
-Using tokens as a global tool to enforce traffic limiting and policies, the scheduler is able to proactively control the flow of traffic as well as respond to application requests.
-"Knowledge" and predictive analysis in networks is a growing trend in research which we intend to leverage in this system.
-We investigate the potential in such a new form of interactive congestion control and analyze it against to state-of-the-art solutions.
+Using tokens as a global tool to enforce traffic limits and admission control, the scheduler is able to proactively steer the flow of traffic.
+The notion of "knowledge" and predictive analysis in networks is a growing trend in research, which we intend to leverage in this system.
+
+We investigate the potential in such a new form of interactive congestion control and analyze it against state-of-the-art solutions.
 Our tool of choice to model our system is Mininet, a rapid prototyping emulator for datacenter networks.
-We will compare our "TCPToken" system against established TCP-congestion systems such as Hedera and DCTCP.
-Based on the findings and measurement results, we will reassess the feasibility and prospect of a token-based, predictive congestion control algorithm.
+We will compare our "Iroko" system against established TCP-congestion systems such as Hedera and DCTCP.
+Based on the findings and measurement results, we will reassess the feasibility and prospects of a predictive congestion control algorithm.
 \end{abstract}
 
-\keywords{TCP, Congestion Control, SDN, Data center}
+\keywords{TCP, Congestion Control, SDN, Datacenter}
 
@@
 
-\input{intro}
+%\input{intro}
+%
+%\input{relatedworks}
+%\input{design}
+% %\input{motivation}
+%\input{implementation}
+%\input{eval}
+%\input{timeline}
 
-\input{relatedworks}
-\input{design}
-%\input{motivation}
-\input{implementation}
-\input{eval}
-\input{timeline}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction / Background}
+\label{sec:intro}
+
+Initially, research in network congestion was dominated by the assumption of a decentralized, autonomous network - end-hosts only have control over the amount of traffic they send, and are unaware of the intentions or traffic rates of their peers.
+Similarly, switches and routers are unaware of global traffic patterns and only forward based on their local notion of optimality.
+
+In line with these assumptions, TCP was designed to optimize traffic globally on a simple fairness principle: nodes react to packet loss and assume that others will behave accordingly.
+An equilibrium is reached when all end-hosts achieve a traffic rate that, in sum, conforms to the maximum available bandwidth of the most congested link.
+TCP works well  in scenarios where many different distrustful participants compete for limited bandwidth.
+Still, TCP is a \textit{reactive protocol}; the fact that packet loss and latency increases occur in the network already indicates a problem.
+Packet loss is primarily due to overflowing queues in forwarding elements or mismatched hardware capabilities, implying that traffic has not been optimally distributed.
+Ideally, a network should always be "zero-queue", i.e., latency will merely be induced by propagation, and not queuing delay.
+
+Queueing has commonly not been a dominating issue in wide-area and enterprise networks, as traffic is sufficiently distributed and diverse, with only few "hot" target hosts.~\cite{hedera, microte} Traffic optimization is a substantial challenge; network operators have no control over the individual network elements nor its participants.
+Under these conditions, TCP and its extensions can be considered a best-effort solution to globally maximize utilization.
+
+Developments in the past decade have changed the general networking environment.
+Datacenters have emerged as an exciting new research frontier, posing novel design challenges and opportunities.
+Driven by minimization of costs and maximization of compute power, data centers must  run at maximum utilization to achieve an optimal compute/cost ratio.
+Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load network, requiring more sophisticated approaches to solve congestion control.
+
+On the other hand, operators now have the ability to freely control and adapt their network architecture, leading to highly customized systems and fine-grained optimization.
+In such an environment, Software-Defined Networking (SDN) has emerged as a new networking paradigm.~\cite{sdn_road}
+
+Moving away from the principle of distributed communication and routing, SDN introduces the notion of "centralized management".
+A single controller with global knowledge is able to automatically modify and adapt the forwarding tables of all switches in the network, while notifying end hosts of changes in the network.
+These two new trends in system design; full architectural control and centralized management, facilitated new opportunities in the space of TCP congestion research.
+Traffic can now be managed in a  \textit{centralized} fashion based on \textit{global knowledge} of the entire topology and traffic patterns.
+
+A new line of centralized schedulers has emerged that can achieve close to optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
+However, these schedulers are still  \textit{reactive}  in nature.
+The central controller responds to changes in the network or requests by applications, which may cost valuable round-trip latency.
+Often, short-term flows or bursts are unaccounted for, which causes undesirable packet loss and back-propagating congestion.
+
+The idea of admission control and service guarantees in networks is not new.~\cite{access_limit, access_limit2}.
+However, such designs traditionally aim to assure quality and bandwidth guarantees in a contentious, decentralized, and untrusted environments such as the internet.
+In a datacenter, these conditions do not apply.
+End-hosts are generally considered reliable and restricted in behavior, which allows for great simplification of enforcement and prioritization policies.
+
+A desirable solution is a global, centralized arbiter which is able to predict and fairly distribute flows in the network before bursts or congestion even occur.
+By treating the network's compute and forwarding power as a single finite resource, a controller acts akin to a OS scheduler distributing CPU time slices to processes.
+This design approach follows SDN's aspiration of introducing operating systems abstractions to the networking domain space.
+
+In this project, we plan to explore the possibilities of a centralized, proactive flow scheduler. We ask ourselves the following research questions:
+\begin{enumerate}
+    \item What are the requirements for such a centralized, predictive scheduler to succeed?
+    \item Is it possible to preemptively regulate a network by analyzing global traffic patterns?
+    \item What latency, packet loss and utilization are we able to achieve?
+\end{enumerate}
+
+In the scope of this course, we attempt to answer these questions and design a simple predictive scheduler in Mininet.
+If successful, we will benchmark our results and evaluate the level of utilization compared to contemporary scheduling systems.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Design / Proposed Approach}
+\label{sec:design}
+
+\subsection{Overview and Goals}
+In our initial simple system, "Iroko", a centralized controller regulates all node traffic by rate-limiting end-hosts.
+We have opted for a centralized manager in favor of a distributed protocol to leverage the advantages of a global network view.~\footnote{For now, we do not expect our design to scale up to thousands of switches.
+Iroko is intended for small to mid-tier size data centers, which may benefit from a simplistic, centralized scheduling model.}
+By polling each switch individually for port and utilization statistics we are able to infer a global traffic matrix, which we can amalgamate with static routing and topology information.
+
+Iroko guarantees minimal congestion and low average access latency.
+Reducing network-global packet loss and jitter is the priority and objective function of the arbiter, which will enforce these goals by restricting host bandwidth.
+In an ideal Iroko system, packet-loss will only rarely, if ever, occur.
+In respect to the free lunch theorem we aim to trade off maximum bandwidth utilization for optimal system stability and reliable latency.
+
+It is important to note that routing decisions are not in the current scope of Iroko.
+While it may certainly beneficial to include dynamic and adaptive routing decisions in a predictive scheduler, we do not include these features in the current system due to time and complexity limitations.
+Iroko operates on static flow routes defined and generated by ECMP.
+It is assumed that these routes will not vary substantially and that the computed ECMP forwarding hash is consistent.
+
+\subsection{Controlling traffic flow}
+Rate-limiting is performed on the granularity of the IP protocol, allowing us to restrict traffic to specific hot regions and to package flows into groups.
+End-hosts are guaranteed a limited amount of bandwidth which they may send to each destination address.
+The traffic window size for any flow is calculated based on the host's current traffic availability.
+Total bandwidth of a group of flows may never exceed the imposed bandwidth limitation, but nodes are free to distribute the allocated resources on singular subflows.
+
+The central arbiter enforces these bandwidth and access limitations by assigning "tokens" to nodes.
+These tokens are plain information packets specifying the affected destination address flow, its bandwidth guarantee, and the expiration time.
+Tokens act as the rate-limiter of the system and form a queue each end-host will cycle through.
+
+When a node opens a flow, it will look up the current bandwidth restriction for the destination IP in its token database and calculate the maximum congestion window possible for the particular TCP connection.
+In our simple design, flows are always assigned a fair fraction.
+
+Once a token has expired for a particular flow-group, the restrictions of the next token in the queue will be applied.
+This mechanism is the underlying basis for a predictive access control algorithm and does not require any application level modification.
+On end-hosts, only the transport layer services will have to be modified.
+
+\subsection{Adapting traffic flow}
+
+Initially, the controller will compute optimal route configuration based on the topology and link bandwidth using a simple heuristic bin-packing approach.
+End-hosts will be initialized with a fixed low-to-medium bandwidth guarantee that will be adjusted over time.
+This bandwidth guarantee will be below the host's proportional bi-section bandwidth that would be used to reach any other host.
+That is, if every end host transfers using the full bandwidth initially allocated to it,  there would be no dropped packets due to congestion.
+
+Of course, this would vastly underutilize network resources, thus the controller will dynamically adjust these allocations to ensure that hosts that require the bandwidth are allocated it, while host that are not currently utilizing all their allocated bandwidth have their allocated bandwidth reduced.
+To avoid starving hosts, some small amount of bandwidth must always be allocated to a given host even if it not utilizing network resources.
+Thus, the network, by design, will never reach 100\% utilization.
+
+Conversely however one should rarely see dropped packets due to network congestion.
+Furthermore, the controller must be able to react quickly to changing bandwidth needs for hosts.
+Ideally, using a predictive method.
+
+\begin{figure}
+\centering
+\includegraphics[width=1\linewidth]{Topology3}
+\caption{Simple sketch of the Iroko architecture.
+A central scheduler maintains a consistent view of the current network activity and computes the optimal end-host bandwidth distribution.
+Bandwidth is enforced at end-hosts on a per-IP basis.}
+\label{fig:Topology3}
+\end{figure}
+
+\subsection{Statistical System}
+One potential implementation for our arbiter is to use reinforcement learning.
+In reinforcement learning, an agent (in our case the arbiter) performs actions in the environment and receives a reward signal from the environment to adjust its decisions in the next epoch ~\cite{Sutton:1998:IRL:551283}.
+The reward is generally used to represent the goal the agent is trying to achieve.
+The agent can freely choose whatever actions are necessary to achieve this goal ~\cite{Sutton:1998:IRL:551283}.
+
+The environment in this case is defined as the statistics collected from each switch.
+This decision is based on the fact that this data is easy to collect and provides a current snapshot of the performance of the network.
+We wish to optimize over the packet loss, and define the reward to be good if packet loss is decreasing (+1 reward); bad if it is increasing ( -1 reward); and acceptable if packet loss remains the same within a threshold (0 reward).
+Using this representation, we take full advantage of all the information provided in our data center; This representation is replicable and can be applied beyond the scope of this project.
+
+Given the data center assumption, the number of end-hosts is known in the topology.
+We can define our arbiter's actions as a discretized representation among an n-dimensional vector where each cell is one of three values: increase decrease, or maintain the allocated bandwidth.
+This representation offers sufficient complexity.
+There are $3^n$ possible actions to perform and to achieve optimal results the agent must explore the environment extensively, precluding a significantly larger action space.
+
+A potential improvement to our representation is a hierarchical learning agent.
+In this case, our arbiter would manage agents for each host and these sub-agents would manage their respective host whether to increase, decrease, or maintain bandwidth for their particular host.
+This problem can be referred to as hierarchical reinforcement learning~\cite{Borga1993}, but may prove to be beyond the scope of our current proposed solution, although has the added benefit of solving a set of sub-problems to address the global problem of packet loss and simplifies the representation spaces in the sub-problems.
+
+Thus we frame our problem as one which can now be solved using techniques in reinforcement learning.
+A classic algorithm to consider is the use of SARSA~\cite{Sutton:1998:IRL:551283}, which is an online temporal difference algorithm used in classical reinforcement learning problems.
+One advantage of SARSA is its online nature~\cite{Sutton:1998:IRL:551283} which promotes more conservative decisions.
+Given full representation of our environment and actions, we can store information in an artificial neural network which is another popular choice in reinforcement learning literature.
+This model would store the reward signal information that is used to make future action choices, and is a theoretically proven function approximator which has made intractable environmental representations manageable~\cite{Sutton:1998:IRL:551283}.
+
+%Using this model, we hope to achieve near 0% packet loss and provide a solution that gets somebody fired at google for not having tried it first.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Implementation}
+\label{sec:implementation}
+We intend to emulate our system in Mininet~\cite{mininet} to observe traffic patterns and infer a suitable token algorithm.
+Mininet has proven itself to be a viable tool to model new congestion control algorithms~\cite{mininet_learning}, and will help us prototype our concept efficiently.
+We will build a custom SDN controller that interacts with traditional OpenFlow software switches as well as end-hosts.
+End-hosts will run a custom real-world traffic generation script which adjusts based on information packets sent by the controller.
+If time permits, we may expand our implementation to MaxiNet, which can emulate large-scale network stress tests on multiple physical hosts.
+
+We initially considered a second implementation alternative in C/C++ based on the FastPass~\cite{fastpass} source code.
+This would provide us with a fully deployable system which we could fork our implementation from.
+A major advantage of this approach is the ability to test scenarios and traffic algorithms using real software code.
+However, several concerns made us favour a Mininet emulation instead.
+Firstly, FastPass relies on DPDK integration, which requires actual hardware interfaces.
+The central arbiter in the FastPass design would need to run on a dedicated machine, which increases prototyping and development complexity substantially.
+
+Secondly, the FastPass code is highly specialized and optimized research code with only little available documentation.
+Modifying and evolving the source code will require thorough understanding of kernel and networking development, a significant time-sink.
+For a class project, these may be major initial hurdles, taking away from the research aspect of the design concept.
+Consequently, we have decided to pursue an approach which allows us to quickly develop an understanding of the problem without being obstructed by engineering work.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Evaluation}
+\label{sec:evaluation}
+
+In the absence of a commercial data center, the implementation of our network design is going to be done on top of mininet with simulated FatTree topologies of varying size.
+We stress test with iPerf and simulate data center traffic using tcpreplay and packet traces provided by~\cite{traffic} .
+
+To evaluate the general effectiveness of our system, we plan to measure against existing centralized as well as decentralized solutions.
+The centralized design will be based on Hedera~\cite{hedera}, a common and influential datacenter scheduler.
+The decentralized congestion control mechanism will be DCTCP~\cite{dctcp}, a state-of-the-art TCP congestion algorithm.
+
+Since we are primarily concerned about reducing the latency and packet drops while keeping utilisation at maximum, the measurements to get a good insight into how well the design can perform are as follows:
+\begin{enumerate}
+
+\item Latency: We are aiming for a low latency network which means that 99th percentile latency in the network across all flows should as low as possible.
+Latency should be measured when all hosts are sending packets at the maximum limit and also with random traffic patterns.
+There should be low latency even during sudden bursts as the transmission rate is limited by a base value.
+
+\item Packet drop rate: Ideally this metric will approach zero, as the objective of Iroko is to minimize packet loss.
+
+\item Fairness: Fairness can be tested by introducing a new host to a completely saturated network and increasing the transmission rate to see if all the hosts gets fair share of the total bandwidth.
+We are assuming all flows should be at equal priority.
+Differential priority is out of scope.
+
+\item Responsiveness: This metric depends on the predictive power and efficiency of Iroko.
+It may be interesting to analyze the response time of the central scheduler compared to decentralized DCTCP and TCP.
+In addition, it is valuable to identify at what data center and flow size the central computing node of the network may become a bottleneck.
+
+\item Network utilization: We will measure the used bandwidth and the total load on the hosts.
+These metrics can be measured by varying the load from each host.
+Load to utilization ratio should be 1 until the saturation point and after that utilization should stay stable at the maximum bisection bandwidth.
+It is also important to check that there is no starvation happening in any host.
+This can be measured by simulating random traffic patterns and plotting residual bandwidth and excess load.
+We expect overall utilization to be lower compared to Hedera and DCTCP.
+Although Iroko will maximize goodput, we do not measure it on grounds of simplicity.
+
+\end{enumerate}
+
+As baseline we will also compare to the default TCP congestion algorithm to estimate how our scheduler fares against the default case.
+As our initial simple design is very constrained in its ability to route and control traffic we expect to achieve less overall optimality against DCTCP and Hedera, but still gain a substantial advantage over common TCP.
+
+\section{Timeline}
+\label{sec:timeline}
+\begin{itemize}
+    \item \underline{October 15th:} Finalize draft.
+    \item \underline{ October 22nd:} Iterate over literature and thoroughly analyze related work. Set up an initial prototyping environment in Mininet on which each team member can work on.
+    \item \underline{October 30th:} Finalize the initial simple design and explore naive congestion modelling in Mininet. Integrate existing Mininet emulations of Hedera and DCTCP.
+    \item \underline{November 12th:} Develop initial naive Iroko prototype involving controller and end-host design.
+    \item \underline{November 19th:} Develop intelligent centralized congestion control heuristic.
+    \item \underline{November 26th:} Run evaluations and microbenchmarks.
+    \item \underline{December 3rd:} Iterate and improve the scheduler, explore topology and bandwidth constraints. Test at scale.
+    \item \underline{December 10th:} Write report and prepare presentation.
+\end{itemize}
 

### 2017-10-20 Hedera_Alt/controllers/hederaController.py
Bootstrap Hedera

@@ -1,0 +1,367 @@
+"""
+HederaController
+To do: poll edge switches for flow byte counts
+When a flow exceeds threshold (>10 percent of host link capacity)
+do a demand estimation
+find a link which will accommodate that demand and reserve
+"""
+
+import logging
+import random
+from struct import pack
+from zlib import crc32
+
+from pox.core import core
+from pox.lib.util import dpidToStr
+import pox.openflow.libopenflow_01 as of
+from pox.lib.revent import EventMixin
+from pox.lib.addresses import EthAddr
+from pox.lib.packet.ipv4 import ipv4
+from pox.lib.packet.udp import udp
+from pox.lib.packet.tcp import tcp
+
+from ripllib.mn import topos
+
+from util import buildTopo, getRouting
+
+log = core.getLogger()
+log.setLevel(logging.WARNING)
+
+# Number of bytes to send for packet_ins
+MISS_SEND_LEN = 2000
+
+
+IDLE_TIMEOUT = 10
+CAPACITY = 1
+
+
+# Borrowed from pox/forwarding/l2_multi
+class Switch (object):
+  def __init__ (self):
+    self.connection = None
+    self.ports = None
+    self.dpid = None
+    self._listeners = None
+
+  def __repr__ (self):
+    return dpidToStr(self.dpid)
+
+  def disconnect (self):
+    if self.connection is not None:
+      log.debug("Disconnect %s" % (self.connection,))
+      self.connection.removeListeners(self._listeners)
+      self.connection = None
+      self._listeners = None
+
+  def connect (self, connection):
+    if self.dpid is None:
+      self.dpid = connection.dpid
+    assert self.dpid == connection.dpid
+    if self.ports is None:
+      self.ports = connection.features.ports
+    self.disconnect()
+    log.debug("Connect %s" % (connection,))
+    self.connection = connection
+    self._listeners = connection.addListeners(self)
+
+  def send_packet_data(self, outport, data = None):
+    msg = of.ofp_packet_out(in_port=of.OFPP_NONE, data = data)
+    msg.actions.append(of.ofp_action_output(port = outport))
+    print(self.connection)
+    self.connection.send(msg)
+
+  def send_packet_bufid(self, outport, buffer_id = None):
+    msg = of.ofp_packet_out(in_port=of.OFPP_NONE)
+    msg.actions.append(of.ofp_action_output(port = outport))
[…]

### 2017-10-20 Hedera_Alt/ripllib/dctopo.py

@@ -1,0 +1,421 @@
+#!/usr/bin/env python
+'''@package dctopo
+
+Data center network topology creation and drawing.
+
+@author Brandon Heller (brandonh@stanford.edu)
+
+This package includes code to create and draw networks with a regular,
+repeated structure.  The main class is StructuredTopo, which augments the
+standard Mininet Topo object with layer metadata plus convenience functions to
+enumerate up, down, and layer edges.
+'''
+
+from mininet.topo import Topo
+
+
+PORT_BASE = 1  # starting index for OpenFlow switch ports
+
+
+class NodeID(object):
+    '''Topo node identifier.'''
+
+    def __init__(self, dpid = None):
+        '''Init.
+
+        @param dpid dpid
+        '''
+        # DPID-compatible hashable identifier: opaque 64-bit unsigned int
+        self.dpid = dpid
+
+    def __str__(self):
+        '''String conversion.
+
+        @return str dpid as string
+        '''
+        return str(self.dpid)
+
+    def name_str(self):
+        '''Name conversion.
+
+        @return name name as string
+        '''
+        return str(self.dpid)
+
+    def ip_str(self):
+        '''Name conversion.
+
+        @return ip ip as string
+        '''
+        hi = (self.dpid & 0xff0000) >> 16
+        mid = (self.dpid & 0xff00) >> 8
+        lo = self.dpid & 0xff
+        return "10.%i.%i.%i" % (hi, mid, lo)
+
+
+class StructuredNodeSpec(object):
+    '''Layer-specific vertex metadata for a StructuredTopo graph.'''
+
+    def __init__(self, up_total, down_total, up_speed, down_speed,
+                 type_str = None):
+        '''Init.
+
+        @param up_total number of up links
+        @param down_total number of down links
+        @param up_speed speed in Gbps of up links
+        @param down_speed speed in Gbps of down links
+        @param type_str string; model of switch or server
+        '''
+        self.up_total = up_total
+        self.down_total = down_total
+        self.up_speed = up_speed
+        self.down_speed = down_speed
+        self.type_str = type_str
+
+
+class StructuredEdgeSpec(object):
+    '''Static edge metadata for a StructuredTopo graph.'''
+
+    def __init__(self, speed = 1.0):
+        '''Init.
+
+        @param speed bandwidth in Gbps
+        '''
+        self.speed = speed
+
+
+class StructuredTopo(Topo):
+    '''Data center network representation for structured multi-trees.'''
+
[…]

### 2017-10-20 Hedera_Alt/run.sh

@@ -1,0 +1,18 @@
+#!/bin/bash
+for file in traffic/*
+do
+	killall -9 python2.7
+	./pox/pox.py controllers.riplpox --topo=ft,4 --routing=hashed --mode=reactive &
+	sleep 2
+	sudo python hedera.py ecmp $file
+	sleep 6
+done
+
+for file in traffic/*
+do
+ 	killall -9 python2.7
+	./pox/pox.py controllers.hederaController --topo=ft,4 &
+	sleep 2
+	sudo python hedera.py gff $file
+	sleep 6
+done

### 2017-10-20 Hedera/hedera.py
Add functional Hedera reproduction.

@@ -1,0 +1,253 @@
+'''
+@author: Milad Sharif (msharif@stanfor.edu)
+'''
+
+from mininet.topo import Topo
+from mininet.node import Controller, RemoteController, OVSKernelSwitch, CPULimitedHost
+from mininet.net import Mininet
+from mininet.link import TCLink
+from mininet.cli import CLI
+from mininet.util import custom
+from mininet.log import setLogLevel, info, warn, error, debug
+
+from DCTopo import FatTreeTopo, NonBlockingTopo
+from DCRouting import Routing
+
+from subprocess import Popen, PIPE
+from argparse import ArgumentParser
+import multiprocessing
+from time import sleep
+from monitor.monitor import monitor_devs_ng
+import os
+import sys
+
+
+# Number of pods in Fat-Tree 
+K = 4
+
+# Queue Size
+QUEUE_SIZE = 100
+
+# Link capacity (Mbps)
+BW = 10 
+
+parser = ArgumentParser(description="ECMP routing")
+
+parser.add_argument('-d', '--dir', dest='output_dir', default='log',
+        help='Output directory')
+
+parser.add_argument('-i', '--input', dest='input_file',
+        default='inputs/all_to_all_data',
+        help='Traffic generator input file')
+
+parser.add_argument('-t', '--time', dest='time', type=int, default=30,
+        help='Duration (sec) to run the experiment')
+
+parser.add_argument('-p', '--cpu', dest='cpu', type=float, default=-1,
+        help='cpu fraction to allocate to each host')
+
+parser.add_argument('-n', '--nonblocking', dest='nonblocking', default=False,
+        action='store_true', help='Run the experiment on the noneblocking topo')
+
+parser.add_argument('--iperf', dest='iperf', default=False, action='store_true',
+        help='Use iperf to generate traffics')
+
+parser.add_argument('--hedera',dest='hedera', default=False,
+        action='store_true', help='Run the experiment with hedera GFF scheduler')
+
+parser.add_argument('--ecmp',dest='ECMP',default=False,
+        action='store_true',help='Run the experiment with ECMP routing')
+
+args = parser.parse_args()
+
+def NonBlockingNet(k=4, bw=10, cpu=-1, queue=100):
+    ''' Create a NonBlocking Net '''
+
+    topo = NonBlockingTopo(k)
[…]

### 2017-10-20 Hedera/DCController.py
Fix Hedera and generate first eval graph

@@ -8,7 +8,7 @@
 import logging
 
 import sys
-sys.path.append('/home/ubuntu/hedera/')
+sys.path.append('./')
 
 from struct import pack
 from zlib import crc32

### 2017-10-20 Hedera/HController.py

@@ -6,7 +6,7 @@
 import logging
 
 import sys
-sys.path.append('/home/ubuntu/hedera/')
+sys.path.append('./')
 
 from struct import pack
 from zlib import crc32

### 2017-10-20 Hedera/pox/pox/datapaths/pcap_switch.py

@@ -95,7 +95,7 @@ def _do_ctl2 (event):
     return "Error: " + str(e)
 
 
-def launch (address = '127.0.0.1', port = 6633, max_retry_delay = 16,
+def launch (address = '127.0.0.1', port = 6653, max_retry_delay = 16,
     dpid = None, ports = '', extra = None, ctl_port = None,
     __INSTANCE__ = None):
   """

### 2017-10-20 Hedera/pox/pox/openflow/libopenflow_01.py

@@ -4430,9 +4430,9 @@ OFP_MAX_TABLE_NAME_LEN = 32
 OFP_DL_TYPE_NOT_ETH_TYPE = 0x05ff
 OFP_DEFAULT_MISS_SEND_LEN = 128
 OFP_MAX_PORT_NAME_LEN = 16
-OFP_SSL_PORT = 6633
+OFP_SSL_PORT = 6653
 OFPFW_ICMP_TYPE = OFPFW_TP_SRC
-OFP_TCP_PORT = 6633
+OFP_TCP_PORT = 6653
 SERIAL_NUM_LEN = 32
 OFP_DEFAULT_PRIORITY = 0x8000
 OFP_VLAN_NONE = 0xffff

### 2017-10-20 Hedera/pox/pox/openflow/of_01.py

@@ -817,7 +817,7 @@ def wrap_socket (new_sock):
   pcapfile = file(fname, "w")
   try:
     new_sock = OFCaptureSocket(new_sock, pcapfile,
-                               local_addrs=(None,None,6633))
+                               local_addrs=(None,None,6653))
   except Exception:
     import traceback
     traceback.print_exc()
@@ -831,7 +831,7 @@ class OpenFlow_01_Task (Task):
   """
   The main recoco thread for listening to openflow messages
   """
-  def __init__ (self, port = 6633, address = '0.0.0.0'):
+  def __init__ (self, port = 6653, address = '0.0.0.0'):
     Task.__init__(self)
     self.port = int(port)
     self.address = address
@@ -952,7 +952,7 @@ _set_handlers()
 # Used by the Connection class
 deferredSender = None
 
-def launch (port = 6633, address = "0.0.0.0"):
+def launch (port = 6653, address = "0.0.0.0"):
   if core.hasComponent('of_01'):
     return None
 

### 2017-11-20 Hedera/DCController.py
Setup eval infrastructure.

@@ -6,6 +6,7 @@
 '''
 
 import sys
+sys.path.append('./')
 
 from struct import pack
 from zlib import crc32
@@ -21,7 +22,6 @@ from pox.lib.packet.tcp import tcp
 
 from util import buildTopo, getRouting
 
-sys.path.append('./')
 
 log = core.getLogger()
 

### 2017-11-20 Hedera/HController.py

@@ -4,8 +4,11 @@
 '''
 
 import logging
-
+import os
+os.path.dirname(os.path.abspath(__file__))
 import sys
+print(os.path.dirname(os.path.abspath(__file__))+'/../../')
+sys.path.append(os.path.dirname(os.path.abspath(__file__))+'/../../')
 
 
 from struct import pack
@@ -23,7 +26,6 @@ from pox.lib.packet.tcp import tcp
 from util import buildTopo, getRouting
 from DemandEstimation import demand_estimation
 from threading import Timer, Lock
-sys.path.append('./')
 log = core.getLogger()
 
 # Number of bytes to send for packet_ins
@@ -207,7 +209,7 @@ class HController(EventMixin):
             self.hostsList = []
             for sw_name in self.t.layer_nodes(self.t.LAYER_EDGE):
                 sw_dpid = self.t.node_gen(name=sw_name).dpid
-                # print 'sw_dpid',sw_dpid ,'sw_name',sw_name
+                print 'sw_dpid',sw_dpid ,'sw_name',sw_name
                 for port in range(1, self.t.k + 1):
                     if not self.t.isPortUp(port):
                         msg = of.ofp_stats_request()

### 2017-11-20 Hedera/hedera.py

@@ -20,6 +20,7 @@ from time import sleep
 from monitor.monitor import monitor_devs_ng
 import os
 import sys
+sys.path.append('./')
 
 
 # Number of pods in Fat-Tree
@@ -196,10 +197,10 @@ def FatTreeTest(args, controller):
 
     hosts = net.hosts
 
-    if args.iperf:
-        iperfTrafficGen(args, hosts, net)
-    else:
-        trafficGen(args, hosts, net)
+    # if args.iperf:
+    #     iperfTrafficGen(args, hosts, net)
+    # else:
+    trafficGen(args, hosts, net)
 
     net.stop()
 

### 2017-11-20 Hedera/run_experiment.sh

@@ -3,27 +3,27 @@ OUTPUT_DIR=results
 INPUT_FILES='stag_prob_0_2_3_data stag_prob_1_2_3_data stag_prob_2_2_3_data stag_prob_0_5_3_data stag_prob_1_5_3_data stag_prob_2_5_3_data stride1_data stride2_data stride4_data stride8_data random0_data random1_data random2_data random0_bij_data random1_bij_data random2_bij_data random_2_flows_data random_3_flows_data random_4_flows_data hotspot_one_to_one_data'
 DURATION=50
 
-for f in $INPUT_FILES;
-do
-        input_file=$INPUT_DIR/$f
-        pref="nonblocking"
-        out_dir=$OUTPUT_DIR/$pref/$f
-        sudo python hedera.py -i $input_file -d $out_dir -p 0.03 -t $DURATION -n --iperf
-done
+# for f in $INPUT_FILES;
+# do
+#         input_file=$INPUT_DIR/$f
+#         pref="nonblocking"
+#         out_dir=$OUTPUT_DIR/$pref/$f
+#         sudo python hedera.py -i $input_file -d $out_dir -p 0.03 -t $DURATION -n
+# done
 
-for f in $INPUT_FILES;
-do
-        input_file=$INPUT_DIR/$f
-        pref="fattree-ecmp"
-        out_dir=$OUTPUT_DIR/$pref/$f
-        sudo python hedera.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --ecmp --iperf
-done
+# for f in $INPUT_FILES;
+# do
+#         input_file=$INPUT_DIR/$f
+#         pref="fattree-ecmp"
+#         out_dir=$OUTPUT_DIR/$pref/$f
+#         sudo python hedera.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --ecmp
+# done
 
 for f in $INPUT_FILES;
 do
         input_file=$INPUT_DIR/$f
         pref="fattree-hedera"
         out_dir=$OUTPUT_DIR/$pref/$f
-        sudo python hedera.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --hedera --iperf
+        sudo python hedera.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --hedera
 done
 

### 2017-11-20 Hedera/util.py

@@ -1,11 +1,16 @@
 # utility functions
-
+import sys
+import os
+print(os.path.dirname(os.path.abspath(__file__))+'/../Iroko')
+sys.path.append(os.path.dirname(os.path.abspath(__file__))+'/../Iroko')
 from DCTopo import FatTreeTopo
 from mininet.util import makeNumeric
 from DCRouting import HashedRouting
+from topo_ecmp import Fattree
 
-TOPOS = {'ft': FatTreeTopo}
+TOPOS = {'ft': FatTreeTopo, 'ft2': Fattree }
 ROUTING = {'ECMP' : HashedRouting}
+
 
 
 def buildTopo(topo):

### 2017-11-20 Iroko/iroko.py

@@ -30,11 +30,29 @@ parser.add_argument('-d', '--dir', dest='output_dir', default='log',
                     help='Output directory')
 
 parser.add_argument('-i', '--input', dest='input_file',
-                    default='../inputs/all_to_all_data',
+                    default='../inputs/stag_prob_0_2_3_data',
                     help='Traffic generator input file')
 
 parser.add_argument('-t', '--time', dest='time', type=int, default=60,
                     help='Duration (sec) to run the experiment')
+
+parser.add_argument('-p', '--cpu', dest='cpu', type=float, default=-1,
+                    help='cpu fraction to allocate to each host')
+
+parser.add_argument('-n', '--nonblocking', dest='nonblocking', default=False,
+                    action='store_true', help='Run the experiment on the noneblocking topo')
+
+parser.add_argument('--iperf', dest='iperf', default=False, action='store_true',
+                    help='Use iperf to generate traffics')
+
+parser.add_argument('--hedera', dest='hedera', default=False,
+                    action='store_true', help='Run the experiment with hedera GFF scheduler')
+
+parser.add_argument('--ecmp', dest='ECMP', default=False,
+                    action='store_true', help='Run the experiment with ECMP routing')
+
+parser.add_argument('--iroko', dest='iroko', default=False,
+                    action='store_true', help='Run the experiment with Iroko rate limiting')
 args = parser.parse_args()
 
 
@@ -49,59 +67,12 @@ def stop_tcpprobe():
     os.system("killall -9 cat")
 
 
-def iperfTrafficGen(args, hosts, net):
-    ''' Generate traffic pattern using iperf and monitor all of thr interfaces
-
-    input format:
-    src_ip dst_ip dst_port type seed start_time stop_time flow_size r/e
-    repetitions time_between_flows r/e (rpc_delay r/e)
-
-    '''
-    host_list = {}
-    for h in hosts:
-        print(h.IP())
-        host_list[h.IP()] = h
-
-    port = 5001
-
-    data = open(args.input_file)
-
-    start_tcpprobe()
[…]

### 2017-11-20 Iroko/plot_rate.py

@@ -1,0 +1,127 @@
+'''
+@author: Milad Sharif(msharif@stanford.edu)
+'''
+
+from monitor.helper import *
+from math import fsum
+import numpy as np
+
+parser = argparse.ArgumentParser()
+parser.add_argument('--input', '-f', dest='files', required=True, help='Input rates')
+parser.add_argument('--out', '-o', dest='out', default=None, 
+        help="Output png file for the plot.")
+
+args= parser.parse_args()
+
+
+''' Output of bwm-ng has the following format:
+    unix_timestamp;iface_name;bytes_out;bytes_in;bytes_total;packets_out;packets_in;packets_total;errors_out;errors_in
+    '''
+
+# traffics=['stag_prob_0_2_3_data', 'stag_prob_1_2_3_data', 'stag_prob_2_2_3_data',
+#         'stag_prob_0_5_3_data','stag_prob_1_5_3_data','stag_prob_2_5_3_data','stride1_data', 
+#         'stride2_data', 'stride4_data', 'stride8_data', 'random0_data', 'random1_data', 'random2_data', 
+#         'random0_bij_data', 'random1_bij_data','random2_bij_data', 'random_2_flows_data', 
+#         'random_3_flows_data', 'random_4_flows_data','hotspot_one_to_one_data']
+
+# labels=['stag0(0.2,0.3)', 'stag1(0.2,0.3)', 'stag2(0.2,0.3)', 'stag0(0.5,0.3)',
+#         'stag1(0.5,0.3)', 'stag2(0.5,0.3)', 'stride1', 'stride2',
+#         'stride4','stride8','rand0', 'rand1', 'rand2','randbij0',
+#         'randbij1','randbij2','randx2','randx3','randx4','hotspot']
+traffics=['stag_prob_0_2_3_data']
+
+labels=['stag0(0.2,0.3)']
+
+def get_bisection_bw(input_file, pat_iface):
+    pat_iface = re.compile(pat_iface)
+    
+    data = read_list(input_file)
+
+    rate = {} 
+    column = 2
+        
+    for row in data:
+        try:
+            ifname = row[1]
+        except:
+            break
+
+        if ifname not in ['eth0', 'lo']:
+            if not rate.has_key(ifname):
+                rate[ifname] = []
+            
+            try:
+                rate[ifname].append(float(row[column]) * 8.0 / (1 << 20))
[…]

### 2017-11-20 Iroko/run_experiment.sh

@@ -1,0 +1,39 @@
+INPUT_DIR=inputs
+OUTPUT_DIR=results
+INPUT_FILES='stag_prob_0_2_3_data stag_prob_1_2_3_data stag_prob_2_2_3_data stag_prob_0_5_3_data stag_prob_1_5_3_data stag_prob_2_5_3_data stride1_data stride2_data stride4_data stride8_data random0_data random1_data random2_data random0_bij_data random1_bij_data random2_bij_data random_2_flows_data random_3_flows_data random_4_flows_data hotspot_one_to_one_data'
+#INPUT_FILES='stag_prob_0_2_3_data'
+DURATION=50
+
+sudo mn -c
+for f in $INPUT_FILES;
+do
+        input_file=$INPUT_DIR/$f
+        pref="nonblocking"
+        out_dir=$OUTPUT_DIR/$pref/$f
+        sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION -n
+done
+
+for f in $INPUT_FILES;
+do
+        input_file=$INPUT_DIR/$f
+        pref="fattree-ecmp"
+        out_dir=$OUTPUT_DIR/$pref/$f
+        sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --ecmp
+done
+
+for f in $INPUT_FILES;
+do
+        input_file=$INPUT_DIR/$f
+        pref="fattree-iroko"
+        out_dir=$OUTPUT_DIR/$pref/$f
+        sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --iroko
+done
+
+# for f in $INPUT_FILES;
+# do
+#         input_file=$INPUT_DIR/$f
+#         pref="fattree-hedera"
+#         out_dir=$OUTPUT_DIR/$pref/$f
+#         sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --hedera
+# done
+

### 2017-11-20 Iroko/topo_ecmp.py

@@ -32,7 +32,37 @@ from mininet.util import custom
 import os
 
 MAX_QUEUE = 1
+DENSITY = 2
 
+class FatNode(object):
+
+    def __init__(self, pod=0, sw=0, host=0, name=None, dpid=None):
+        ''' Create FatTreeNode '''
+        if dpid:
+            self.pod = (dpid & 0xff0000) >> 16
+            self.sw = (dpid & 0xff00) >> 8
+            self.host = (dpid & 0xff)
+            self.dpid = dpid
+        else:
+            if name:
+                pod, sw, host = [int(s) for s in name.split('h')]
+
+            self.pod = pod
+            self.sw = sw
+            self.host = host
+            self.dpid = (pod << 16) + (sw << 8) + host
+
+    def name_str(self):
+        ''' Return name '''
+        return "%ih%ih%i" % (self.pod, self.sw, self.host)
+
+    def ip_str(self):
+        ''' Return IP address '''
+        return "10.%i.%i.%i" % (self.pod, self.sw, self.host)
+
+    def mac_str(self):
+        ''' Return MAC address '''
+        return "00:00:00:%02x:%02x:%02x" % (self.pod, self.sw, self.host)
 
 class Fattree(Topo):
     """
@@ -43,13 +73,14 @@ class Fattree(Topo):
     EdgeSwitchList = []
     HostList = []
 
-    def __init__(self, k, density):
+    def __init__(self, k):
         self.pod = k
-        self.density = density
+        self.density = DENSITY
         self.iCoreLayerSwitch = (k / 2)**2
         self.iAggLayerSwitch = k * k / 2
         self.iEdgeLayerSwitch = k * k / 2
-        self.iHost = self.iEdgeLayerSwitch * density
+        self.iHost = self.iEdgeLayerSwitch * DENSITY
+        self.node_gen = FatNode
 
         # Init Topo
         Topo.__init__(self)
@@ -123,17 +154,17 @@ class Fattree(Topo):
                     self.HostList[self.density * x + i],
                     bw=bw_e2h, max_queue_size=MAX_QUEUE)   # use_htb=False
 
-    def set_ovs_protocol_13(self,):
+    def set_ovs_protocol(self, ovs_v):
         """
             Set the OpenFlow version for switches.
[…]

### 2017-11-20 Iroko/topo_non_block.py

@@ -102,10 +102,16 @@ def set_host_ip(net, topo):
     for k in range(len(topo.HostList)):
         hostlist.append(net.get(topo.HostList[k]))
     i = 1
+    j = 1
+    # for host in hostlist:
+    #     host.setIP("10.0.0.%d" % i)
+    #     i += 1
     for host in hostlist:
-        host.setIP("10.0.0.%d" % i)
-        i += 1
-
+        host.setIP("10.%d.0.%d" % (i, j))
+        j += 1
+        if j == 3:
+            j = 1
+            i += 1
 
 def install_proactive(net, topo):
     """
@@ -122,6 +128,14 @@ def install_proactive(net, topo):
                 nw_dst=10.0.0.%d,actions=output:%d'" % (sw, i, i)
             os.system(cmd)
 
+def configureTopo(net, topo):
+    # Set OVS's protocol as OF13.
+    topo.set_ovs_protocol_13()
+    # Set hosts IP addresses.
+    set_host_ip(net, topo)
+    # Install proactive flow entries
+    install_proactive(net, topo)
+
 def createNonBlockTopo(pod, ip="172.18.232.60", port=6653, bw_h2c=10):
     """
             Firstly, start up Mininet;
@@ -137,14 +151,6 @@ def createNonBlockTopo(pod, ip="172.18.232.60", port=6653, bw_h2c=10):
     CONTROLLER_PORT = port
     net = Mininet(topo=topo, link=TCLink, controller=RemoteController, autoSetMacs=True)
     #net.addController('controller', controller=RemoteController, ip=CONTROLLER_IP, port=CONTROLLER_PORT)
-    net.start()
+    #net.start()
 
-    # Set the OpenFlow version for switches as 1.3.0.
-    topo.set_ovs_protocol_13()
-    # Set the IP addresses for hosts.
-    set_host_ip(net, topo)
-    # Install proactive flow entries.
-    install_proactive(net, topo)
-
-    time.sleep(5)
     return net, topo

### 2017-11-21 Documents/Proposal/eval.tex
Push complete eval code

--- 2017-10-10 Proposal/paper.tex
@@
-\documentclass[sigconf]{acmart}
-
-\usepackage{booktabs} % For formal tables
-\usepackage[normalem]{ulem}
-% Copyright
-\setcopyright{none}
-
-\input{annotations}
-
-\begin{document}
-
-\title{TCPTokens: Introducing Currency into Datacenter Congestion Control}
-
-\author{Anand Jayarajan}
-\affiliation{%
-  \institution{University of British Columbia}
-}
-\email{anandj@cs.ubc.ca}
-
-\author{Michael Przystupa}
-\affiliation{%
-  \institution{University of British Columbia}
-}
-\email{michael.przystupa@gmail.com}
-
-\author{Robert Reiss}
-\affiliation{%
-  \institution{University of British Columbia}
-}
-\email{rreiss@cs.ubc.ca}
-
-\author{Fabian Ruffy}
-\affiliation{%
-  \institution{University of British Columbia}
-}
-\email{fruffy@cs.ubc.ca}
-
-\begin{abstract}
-Datacenter networks have become a hotbed for research in the recent past.
-Many works have been published on improving the  bisection bandwidth, reducing cost, and latency.
-As a core contribution of this research, centralized scheduling has entered the stage as dominant strategy to manage flows.
-Benefitting from a global view and full control over the network, centralized schedulers are able to enforce fine-grained traffic control, frequently achieving near-optimal bandwidth optimization.
-However, modern congestion control algorithms and schedulers are still fundamentally reactive.
-Many algorithms are designed to only respond to packet loss or significant increase in latency, not to actively prevent it.
-
-In this project, we plan to explore the opportunities of employing a preemptive and tightly controlling central network scheduler.
-Using tokens as a global tool to enforce traffic limiting and policies, the scheduler is able to proactively control the flow of traffic as well as respond to application requests.
-"Knowledge" and predictive analysis in networks is a growing trend in research which we intend to leverage in this system.
-
-We investigate the potential in such a new form of interactive congestion control and analyze it against state-of-the-art solutions.
-Our tool of choice to model our system is Mininet, a rapid prototyping emulator for datacenter networks.
-We will compare our "TCPToken" system against established TCP-congestion systems such as Hedera and DCTCP.
-Based on the findings and measurement results, we will reassess the feasibility and prospects of a token-based, predictive congestion control algorithm.
-\end{abstract}
-
-\keywords{TCP, Congestion Control, SDN, Datacenter}
-
-\maketitle
-
-%\input{intro}
-%
-%\input{relatedworks}
-%\input{design}
-% %\input{motivation}
-%\input{implementation}
-%\input{eval}
-%\input{timeline}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-\label{sec:intro}
-
-Initially, research in network congestion was dominated by the assumption of a decentralised, autonomous network - end-hosts only have control over the amount of traffic they send, and are unaware of the intentions or traffic rates of their peers.
-Similarly, switches and routers are unaware of global traffic patterns and only forward based on their local notion of optimality.
-
-In line with these assumptions, TCP was designed to optimize traffic globally on a simple fairness principle: nodes react to packet loss and assume that others will behave accordingly.
-An equilibrium is reached when all end-hosts achieve a traffic rate that, in sum, conforms to the maximum available bandwidth of the most congested link.
-TCP works well  in scenarios where many different distrustful participants compete for limited bandwidth.
-Still, TCP is a \textit{reactive protocol}; the fact that packet loss and latency increases occur in the network already indicates a problem.
-Packet loss is primarily due to overflowing queues in forwarding elements, implying that traffic has not been optimally distributed.
-Ideally, a network should always be "zero-queue", i.e., latency will merely be induced by propagation, and not queuing delay.
-
-Queueing has generally not been a dominating issue in wide-area and enterprise networks, as traffic is sufficiently distributed and diverse, with only few "hot" target hosts.~\cite{hedera, microte} Traffic optimization is a substantial challenge; network operators have no control over the individual network elements nor its participants.
-Under these conditions, TCP and its extensions can be considered a best-effort solution.
-
-Developments in the past decade have changed the general networking environment.
-Datacenters have emerged as an exciting new research frontier, posing novel design challenges and opportunities.Driven by minimization of costs and maximization of compute power, datacenters must  run at maximum utilization to achieve an optimal compute/cost ratio.
-Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load network, requiring more sophisticated approaches to solve congestion control.
-On the other hand, operators now have the ability to freely control and adapt their network architecture, leading to highly customized systems and fine-grained optimization.
-As a result, Software-Defined Networking (SDN) has emerged as a new networking paradigm.
-
-Moving away from the principle of distributed communication and routing, SDN introduces the notion of "centralized management".
-A single controller with global knowledge is able to automatically modify and adapt the forwarding tables of all switches in the network, while notifying end hosts of changes in the network.
-These two new trends in system design facilitated impactful new innovation opportunities in the space of TCP congestion research.
-Traffic can now be managed in a  \textit{centralised} fashion based on \textit{global knowledge} of the entire topology and traffic patterns.
-
-A new line of centralised schedulers has emerged that  can achieve close to optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
-However, these schedulers are still  \textit{reactive}  in nature.
-The central controller responds to changes in the network or requests by applications, which may cost valuable roundtrip latency.
-Often, short-term flows or bursts are unaccounted for, which causes undesirable packet loss and backpropagating congestion.
-
-A much more desirable solution is a global, centralised arbiter which is able to predict and fairly distribute flows in the network before bursts or congestion occur.
-By treating the network's compute and forwarding power as a single finite resource, a controller acts like the OS scheduler distributing CPU time slices to processes.
-This design approach follows SDN's aspiration of introducing operating systems abstractions to the networking domain space.
-
-In this project, we plan to explore the possibilities of a centralised, proactive flow scheduler. We ask ourselves the following research questions:
-\begin{enumerate}
-	\item Is it possible to design a centralised token-based scheduling network?
-	\item Is it possible to predict traffic and preemptively schedule flows and token distribution in a datacenter context?
-	\item Using this approach, are we able to achieve better performance and utilization than existing solutions?
-\end{enumerate}
-
-In the scope of this course, we attempt to answer question one and design a simple token-based scheduler in Mininet.
-If successful, we will benchmark our results and evaluate the level of utilization compared to contemporary scheduling systems.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}
-\label{sec:related}
-
-Most of the work in congestion control relies on a centralised scheduler which periodically gets updates about the flows in the network and reserves specific paths for long running flows.
-
-MicroTE~\cite{microte} does predictive OpenFlow routing and scheduling using a central SDN controller.
-Hedera~\cite{hedera} schedules large so called "elephant" flows by gathering switch statistics.
-However it pays poor attention to short term flows and network latency.
-FastPass~\cite{fastpass} on the other hand, proposes a zero-queue network and promises a very low latency network.
-The feasibility of this proposal remains questionable when considering the high cost of implementation.
-
-Recent proposals on decentralised scheduling like ExpressPass~\cite{expresspass} have heavily influenced our research.
-ExpressPass posits a credit based system similar to TCPTokens in order to control the congestion in the network.
-Effecting flow rerouting using ML techniques in KDN~\cite{kdn} and other works shows promising signs of further improvements in this area.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Design / Proposed Approach}
-\label{sec:design}
-
-\subsection{Simple System}
-In our initial simple TCPToken design, a centralized controller regulates all node traffic by provisioning end-hosts with tokens.
-These tokens act as the "currency" of the system.
-The total amount is fixed to some function of the bisection and aggregate bandwidths available in the network, and the traffic window size of sending end hosts is calculated based on the host's current token availability.
-If a node or switch is overburdened, it may notify the scheduler, which will adjust the traffic window of responsible sending nodes remotely.
-Management and control operations including token distribution and updates are priority-queued to guarantee a fast and low-latency response.
-If an end-host requires more bandwidth, it can also notify the controller, which may comply based on priority and availability.
-Initially, the controller will compute optimal route configuration based on the topology and link bandwidth using a simple heuristic bin-packing approach.
-End-hosts will be initialized with a fixed low-to-medium bandwidth guarantee that will be adjusted over time.
-
-\subsection{Advanced System}
-The aforementioned system is intended only as a proof of concept.
-It is still a  synchronous and reactive approach, dependent on notifications and requests by overburdened switches and underprovisioned hosts.
-Ideally, all management in the TCPToken system should be asynchronous and preemptive. 
-\subsubsection{ Paying for traffic}
-In a complete design, tokens will be used as a transport pass instead of simply acting as traffic shaping parameters.
-On a per-flow basis, nodes will be able to attach tokens to traffic as a form of payment.
-Enforcement is performed on the switch level, any flow that is not "paid" by a token will be dropped.
-Nodes may be handed different types of tokens to serve load-balancing and traffic shaping purposes.
-Applications operating on traditionally short-lived flows and bursts may sign their traffic with a different token type than long-lived permanent processes.
-Switches will forward traffic and balance flows accordingly with only minor interference from the central controller.
-
-\subsubsection{Predicting traffic}
-ML-based scheduling is an emerging research field, also known as Knowledge-Defined Networking. Datacenter traffic follows patterns, which can be predicted using statistical methods.
-Combined with using tokens as a traffic engineering enforcement, a preemptive scheduling strategy in datacenters may be feasible.
-
-We do not intend to implement the advanced system for this project, but are using it as a guideline to motivate our research.
-Based on the experiences and knowledge gained from implementing the basic system, we will reevaluate the feasibility and practicality of the second concept.
-
-\subsubsection{Fault-Tolerance and Scalability}
-To handle fault-tolerance we plan to integrate a fallback-mechanism to conventional TCP, if the node has not received a TCPToken in a prefixed amount of time.
-End-hosts will only be marginally affected by a controller or link failure, as they are guaranteed a fixed amount of bandwidth based on their token availability.
-
-To avoid overburdening the controller, we envision a distributed hive of schedulers (e.g., Onos) each managing their own region in the final design.~\cite{onos} In case of failure, a shadow node or neighbouring device may take over computation temporarily, until the main node has recovered.~\cite{disco}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Implementation}
-\label{sec:implementation}
-We intend to emulate our system in Mininet~\cite{mininet} to observe traffic patterns and infer a suitable token algorithm.
-Mininet has proven itself to be a viable tool to model new congestion control algorithms~\cite{mininet_learning}, and will help us prototype our concept efficiently.
-We will build a custom SDN controller that interacts with traditional OpenFlow software switches as well as end-hosts.
-End-hosts will run a custom real-world traffic generation script which adjusts based on information packets sent by the controller.
-If time permits, we may expand our implementation to MaxiNet, which can emulate large-scale network stress tests of thousands of nodes on multiple physical hosts.
-
-We initially considered a second implementation alternative in C/C++ based on the FastPass~\cite{fastpass} source code.
-This would provide us with a fully deployable system which we could fork our implementation from.
-A major advantage of this approach is the ability to test scenarios and traffic algorithms using real software code.
-However, several concerns made us favour a Mininet emulation instead.
-Firstly, FastPass relies on DPDK integration, which requires actual hardware interfaces.
-The central arbiter in the FastPass design would need to run on a dedicated machine, which increases prototyping and development complexity substantially.
-
-Secondly, the FastPass code is highly specialized and optimized research code with only little available documentation.
-Modifying and evolving the source code will require thorough understanding of kernel and networking development, a significant time-sink.
-For a class project, these may be major initial hurdles, taking away from the research aspect of the design concept.
-Consequently, we have decided to pursue an approach which allows us to quickly develop an understanding of the problem without being obstructed by engineering work.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The centralized design will be based on Hedera~\cite{hedera}, a common and influential datacenter scheduler.
-
 The decentralized congestion control mechanism will be DCTCP~\cite{dctcp}, a state-of-the-art TCP congestion algorithm.
@@
 As our initial simple design is very constrained in its ability to route and control traffic we expect to achieve less optimality against DCTCP and Hedera, but gain a substantial advantage over common TCP.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Timeline}
-\label{sec:timeline}
-\begin{itemize}
-	\item \underline{October 15th:} Finalize draft and related work / background section.
-	\item \underline{ October 22nd:} Iterate over literature and thoroughly analyze related work. Set up an initial prototyping environment in Mininet on which each team member can work on.
-	\item \underline{October 30th:} Finalize the initial simple design and explore naive congestion modelling in Mininet. Integrate existing Mininet emulations of Hedera and DCTCP.
-	\item \underline{November 12th:} Develop initial naive TCPTokens prototype involving controller and end-host design.
-	\item \underline{November 19th:} Develop intelligent centralised congestion control heuristic.
-	\item \underline{November 26th:} Run evaluations and microbenchmarks.
-	\item \underline{December 3rd:} Iterate and improve the scheduler, explore topology and bandwidth constraints. Test at scale.
-	\item \underline{December 10th:} Write report and prepare presentation.
-\end{itemize}
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{paper}
-
-\end{document}

### 2017-11-21 Documents/Proposal/timeline.tex
Push complete eval code

--- 2017-10-10 Proposal/paper.tex
@@
-\documentclass[sigconf]{acmart}
-
-\usepackage{booktabs} % For formal tables
-\usepackage[normalem]{ulem}
-% Copyright
-\setcopyright{none}
-
-\input{annotations}
-
-\begin{document}
-
-\title{TCPTokens: Introducing Currency into Datacenter Congestion Control}
-
-\author{Anand Jayarajan}
-\affiliation{%
-  \institution{University of British Columbia}
-}
-\email{anandj@cs.ubc.ca}
-
-\author{Michael Przystupa}
-\affiliation{%
-  \institution{University of British Columbia}
-}
-\email{michael.przystupa@gmail.com}
-
-\author{Robert Reiss}
-\affiliation{%
-  \institution{University of British Columbia}
-}
-\email{rreiss@cs.ubc.ca}
-
-\author{Fabian Ruffy}
-\affiliation{%
-  \institution{University of British Columbia}
-}
-\email{fruffy@cs.ubc.ca}
-
-\begin{abstract}
-Datacenter networks have become a hotbed for research in the recent past.
-Many works have been published on improving the  bisection bandwidth, reducing cost, and latency.
-As a core contribution of this research, centralized scheduling has entered the stage as dominant strategy to manage flows.
-Benefitting from a global view and full control over the network, centralized schedulers are able to enforce fine-grained traffic control, frequently achieving near-optimal bandwidth optimization.
-However, modern congestion control algorithms and schedulers are still fundamentally reactive.
-Many algorithms are designed to only respond to packet loss or significant increase in latency, not to actively prevent it.
-
-In this project, we plan to explore the opportunities of employing a preemptive and tightly controlling central network scheduler.
-Using tokens as a global tool to enforce traffic limiting and policies, the scheduler is able to proactively control the flow of traffic as well as respond to application requests.
-"Knowledge" and predictive analysis in networks is a growing trend in research which we intend to leverage in this system.
-
-We investigate the potential in such a new form of interactive congestion control and analyze it against state-of-the-art solutions.
-Our tool of choice to model our system is Mininet, a rapid prototyping emulator for datacenter networks.
-We will compare our "TCPToken" system against established TCP-congestion systems such as Hedera and DCTCP.
-Based on the findings and measurement results, we will reassess the feasibility and prospects of a token-based, predictive congestion control algorithm.
-\end{abstract}
-
-\keywords{TCP, Congestion Control, SDN, Datacenter}
-
-\maketitle
-
-%\input{intro}
-%
-%\input{relatedworks}
-%\input{design}
-% %\input{motivation}
-%\input{implementation}
-%\input{eval}
-%\input{timeline}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-\label{sec:intro}
-
-Initially, research in network congestion was dominated by the assumption of a decentralised, autonomous network - end-hosts only have control over the amount of traffic they send, and are unaware of the intentions or traffic rates of their peers.
-Similarly, switches and routers are unaware of global traffic patterns and only forward based on their local notion of optimality.
-
-In line with these assumptions, TCP was designed to optimize traffic globally on a simple fairness principle: nodes react to packet loss and assume that others will behave accordingly.
-An equilibrium is reached when all end-hosts achieve a traffic rate that, in sum, conforms to the maximum available bandwidth of the most congested link.
-TCP works well  in scenarios where many different distrustful participants compete for limited bandwidth.
-Still, TCP is a \textit{reactive protocol}; the fact that packet loss and latency increases occur in the network already indicates a problem.
-Packet loss is primarily due to overflowing queues in forwarding elements, implying that traffic has not been optimally distributed.
-Ideally, a network should always be "zero-queue", i.e., latency will merely be induced by propagation, and not queuing delay.
-
-Queueing has generally not been a dominating issue in wide-area and enterprise networks, as traffic is sufficiently distributed and diverse, with only few "hot" target hosts.~\cite{hedera, microte} Traffic optimization is a substantial challenge; network operators have no control over the individual network elements nor its participants.
-Under these conditions, TCP and its extensions can be considered a best-effort solution.
-
-Developments in the past decade have changed the general networking environment.
-Datacenters have emerged as an exciting new research frontier, posing novel design challenges and opportunities.Driven by minimization of costs and maximization of compute power, datacenters must  run at maximum utilization to achieve an optimal compute/cost ratio.
-Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load network, requiring more sophisticated approaches to solve congestion control.
-On the other hand, operators now have the ability to freely control and adapt their network architecture, leading to highly customized systems and fine-grained optimization.
-As a result, Software-Defined Networking (SDN) has emerged as a new networking paradigm.
-
-Moving away from the principle of distributed communication and routing, SDN introduces the notion of "centralized management".
-A single controller with global knowledge is able to automatically modify and adapt the forwarding tables of all switches in the network, while notifying end hosts of changes in the network.
-These two new trends in system design facilitated impactful new innovation opportunities in the space of TCP congestion research.
-Traffic can now be managed in a  \textit{centralised} fashion based on \textit{global knowledge} of the entire topology and traffic patterns.
-
-A new line of centralised schedulers has emerged that  can achieve close to optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
-However, these schedulers are still  \textit{reactive}  in nature.
-The central controller responds to changes in the network or requests by applications, which may cost valuable roundtrip latency.
-Often, short-term flows or bursts are unaccounted for, which causes undesirable packet loss and backpropagating congestion.
-
-A much more desirable solution is a global, centralised arbiter which is able to predict and fairly distribute flows in the network before bursts or congestion occur.
-By treating the network's compute and forwarding power as a single finite resource, a controller acts like the OS scheduler distributing CPU time slices to processes.
-This design approach follows SDN's aspiration of introducing operating systems abstractions to the networking domain space.
-
-In this project, we plan to explore the possibilities of a centralised, proactive flow scheduler. We ask ourselves the following research questions:
-\begin{enumerate}
-	\item Is it possible to design a centralised token-based scheduling network?
-	\item Is it possible to predict traffic and preemptively schedule flows and token distribution in a datacenter context?
-	\item Using this approach, are we able to achieve better performance and utilization than existing solutions?
-\end{enumerate}
-
-In the scope of this course, we attempt to answer question one and design a simple token-based scheduler in Mininet.
-If successful, we will benchmark our results and evaluate the level of utilization compared to contemporary scheduling systems.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}
-\label{sec:related}
-
-Most of the work in congestion control relies on a centralised scheduler which periodically gets updates about the flows in the network and reserves specific paths for long running flows.
-
-MicroTE~\cite{microte} does predictive OpenFlow routing and scheduling using a central SDN controller.
-Hedera~\cite{hedera} schedules large so called "elephant" flows by gathering switch statistics.
-However it pays poor attention to short term flows and network latency.
-FastPass~\cite{fastpass} on the other hand, proposes a zero-queue network and promises a very low latency network.
-The feasibility of this proposal remains questionable when considering the high cost of implementation.
-
-Recent proposals on decentralised scheduling like ExpressPass~\cite{expresspass} have heavily influenced our research.
-ExpressPass posits a credit based system similar to TCPTokens in order to control the congestion in the network.
-Effecting flow rerouting using ML techniques in KDN~\cite{kdn} and other works shows promising signs of further improvements in this area.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Design / Proposed Approach}
-\label{sec:design}
-
-\subsection{Simple System}
-In our initial simple TCPToken design, a centralized controller regulates all node traffic by provisioning end-hosts with tokens.
-These tokens act as the "currency" of the system.
-The total amount is fixed to some function of the bisection and aggregate bandwidths available in the network, and the traffic window size of sending end hosts is calculated based on the host's current token availability.
-If a node or switch is overburdened, it may notify the scheduler, which will adjust the traffic window of responsible sending nodes remotely.
-Management and control operations including token distribution and updates are priority-queued to guarantee a fast and low-latency response.
-If an end-host requires more bandwidth, it can also notify the controller, which may comply based on priority and availability.
-Initially, the controller will compute optimal route configuration based on the topology and link bandwidth using a simple heuristic bin-packing approach.
-End-hosts will be initialized with a fixed low-to-medium bandwidth guarantee that will be adjusted over time.
-
-\subsection{Advanced System}
-The aforementioned system is intended only as a proof of concept.
-It is still a  synchronous and reactive approach, dependent on notifications and requests by overburdened switches and underprovisioned hosts.
-Ideally, all management in the TCPToken system should be asynchronous and preemptive. 
-\subsubsection{ Paying for traffic}
-In a complete design, tokens will be used as a transport pass instead of simply acting as traffic shaping parameters.
-On a per-flow basis, nodes will be able to attach tokens to traffic as a form of payment.
-Enforcement is performed on the switch level, any flow that is not "paid" by a token will be dropped.
-Nodes may be handed different types of tokens to serve load-balancing and traffic shaping purposes.
-Applications operating on traditionally short-lived flows and bursts may sign their traffic with a different token type than long-lived permanent processes.
-Switches will forward traffic and balance flows accordingly with only minor interference from the central controller.
-
-\subsubsection{Predicting traffic}
-ML-based scheduling is an emerging research field, also known as Knowledge-Defined Networking. Datacenter traffic follows patterns, which can be predicted using statistical methods.
-Combined with using tokens as a traffic engineering enforcement, a preemptive scheduling strategy in datacenters may be feasible.
-
-We do not intend to implement the advanced system for this project, but are using it as a guideline to motivate our research.
-Based on the experiences and knowledge gained from implementing the basic system, we will reevaluate the feasibility and practicality of the second concept.
-
-\subsubsection{Fault-Tolerance and Scalability}
-To handle fault-tolerance we plan to integrate a fallback-mechanism to conventional TCP, if the node has not received a TCPToken in a prefixed amount of time.
-End-hosts will only be marginally affected by a controller or link failure, as they are guaranteed a fixed amount of bandwidth based on their token availability.
-
-To avoid overburdening the controller, we envision a distributed hive of schedulers (e.g., Onos) each managing their own region in the final design.~\cite{onos} In case of failure, a shadow node or neighbouring device may take over computation temporarily, until the main node has recovered.~\cite{disco}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Implementation}
-\label{sec:implementation}
-We intend to emulate our system in Mininet~\cite{mininet} to observe traffic patterns and infer a suitable token algorithm.
-Mininet has proven itself to be a viable tool to model new congestion control algorithms~\cite{mininet_learning}, and will help us prototype our concept efficiently.
-We will build a custom SDN controller that interacts with traditional OpenFlow software switches as well as end-hosts.
-End-hosts will run a custom real-world traffic generation script which adjusts based on information packets sent by the controller.
-If time permits, we may expand our implementation to MaxiNet, which can emulate large-scale network stress tests of thousands of nodes on multiple physical hosts.
-
-We initially considered a second implementation alternative in C/C++ based on the FastPass~\cite{fastpass} source code.
-This would provide us with a fully deployable system which we could fork our implementation from.
-A major advantage of this approach is the ability to test scenarios and traffic algorithms using real software code.
-However, several concerns made us favour a Mininet emulation instead.
-Firstly, FastPass relies on DPDK integration, which requires actual hardware interfaces.
-The central arbiter in the FastPass design would need to run on a dedicated machine, which increases prototyping and development complexity substantially.
-
-Secondly, the FastPass code is highly specialized and optimized research code with only little available documentation.
-Modifying and evolving the source code will require thorough understanding of kernel and networking development, a significant time-sink.
-For a class project, these may be major initial hurdles, taking away from the research aspect of the design concept.
-Consequently, we have decided to pursue an approach which allows us to quickly develop an understanding of the problem without being obstructed by engineering work.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Evaluation}
-\label{sec:evaluation}
-To evaluate the effectiveness of our system, we plan to measure against existing centralized as well as decentralized solutions.
-We will run datacenter traffic on a fat-tree topology of various depth and observe the effect each congestion algorithm will have on the overall bandwidth utilization and latency.
-The centralized design will be based on Hedera~\cite{hedera}, a common and influential datacenter scheduler.
-
-The decentralized congestion control mechanism will be DCTCP~\cite{dctcp}, a state-of-the-art TCP congestion algorithm.
-As baseline we will also compare to the default TCP congestion algorithm to estimate how our scheduler fares against the default case.
-As our initial simple design is very constrained in its ability to route and control traffic we expect to achieve less optimality against DCTCP and Hedera, but gain a substantial advantage over common TCP.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{itemize}
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{paper}
-
-\end{document}

### 2017-11-21 Deprecated/Hedera/util.py

@@ -1,0 +1,16 @@
+# utility functions
+from DCTopo import FatTreeTopo
+from mininet.util import makeNumeric
+from DCRouting import HashedRouting
+
+TOPOS = {'ft': FatTreeTopo}
+ROUTING = {'ECMP': HashedRouting}
+
+
+def buildTopo(topo):
+    topo_name, topo_param = topo.split(',')
+    return TOPOS[topo_name](makeNumeric(topo_param))
+
+
+def getRouting(routing, topo):
+    return ROUTING[routing](topo)

### 2017-11-21 Hedera/util.py

@@ -1,22 +1,0 @@
-# utility functions
-import sys
-import os
-print(os.path.dirname(os.path.abspath(__file__))+'/../Iroko')
-sys.path.append(os.path.dirname(os.path.abspath(__file__))+'/../Iroko')
-from DCTopo import FatTreeTopo
-from mininet.util import makeNumeric
-from DCRouting import HashedRouting
-from topo_ecmp import Fattree
-
-TOPOS = {'ft': FatTreeTopo, 'ft2': Fattree }
-ROUTING = {'ECMP' : HashedRouting}
-
-
-
-def buildTopo(topo):
-    topo_name, topo_param = topo.split( ',' )
-    return TOPOS[topo_name](makeNumeric(topo_param))
-
-
-def getRouting(routing, topo):
-    return ROUTING[routing](topo)

### 2017-11-21 Iroko/hedera/DemandEstimation.py

@@ -1,0 +1,122 @@
+'''the demand estimation algorithm from Hedera paper is implemented here'''
+import random
+import pdb
+
+
+def demand_estimation(flows, hostsList):
+    M = {}
+    for i in hostsList:
+        M[i] = {}
+        for j in hostsList:
+            M[i][j] = {'demand': 0, 'demandInit': 0, 'converged': False, 'FlowNmbr': 0}
+
+    for flow in flows:
+        M[flow['src']][flow['dst']]['FlowNmbr'] += 1
+
+    demandChange = True
+    while demandChange:
+        demandChange = False
+        for src in hostsList:
+            Est_Src(M, flows, src)
+
+        for dst in hostsList:
+            Est_Dst(M, flows, dst)
+
+        for i in hostsList:
+            for j in hostsList:
+                if M[i][j]['demandInit'] != M[i][j]['demand']:
+                    NoChange = True
+                    M[i][j]['demandInit'] = M[i][j]['demand']
+
+    print"********************estimated demands*********************\n", demandsPrinting(M, hostsList)
+    return (M, flows)
+
+
+def Est_Src(M, flows, src):
+    dF = 0
+    nU = 0
+    for flow in flows:
+        if flow['src'] == src:
+            if flow['converged']:
+                dF += flow['demand']
+            else:
+                nU += 1
+    if nU != 0:
+
+        eS = (1.0 - dF) / nU
+        for flow in flows:
+            if flow['src'] == src and not flow['converged']:
+                M[flow['src']][flow['dst']]['demand'] = eS
+                # pdb.set_trace()
+                flow['demand'] = eS
+
+
+def Est_Dst(M, flows, dst):
+    dT = 0
+    dS = 0
+    nR = 0
+    for flow in flows:
+        if flow['dst'] == dst:
+            flow['recLimited'] = True
+            dT += flow['demand']
+            nR += 1
+    if dT <= 1.0:
+        return
+    eS = 1.0 / nR
+
+    flagFlip = True
+    while flagFlip:
+        flagFlip = False
+        nR = 0
+        for flow in flows:
+            if flow['dst'] == dst and flow['recLimited']:
+                if flow['demand'] < eS:
+                    dS += flow['demand']
+                    flow['recLimited'] = False
+                    flagFlip = True
+                else:
+                    nR += 1
+
+        eS = (1.0 - dS) / nR
+
+    for flow in flows:
[…]

### 2017-11-21 Iroko/hedera/HController.py

@@ -1,0 +1,346 @@
+''' Hedera data center controller
+
+@author: Behnam Montazeri (behnamm@stanford.edu)
+'''
+
+import logging
+import os
+import sys
+os.path.dirname(os.path.abspath(__file__))
+print(os.path.dirname(os.path.abspath(__file__)) + '/../../')
+sys.path.append(os.path.dirname(os.path.abspath(__file__)) + '/../../')
+
+
+from struct import pack
+from zlib import crc32
+
+from pox.core import core
+import pox.openflow.libopenflow_01 as of
+from pox.lib.revent import EventMixin
+from pox.lib.util import dpidToStr
+
+from pox.lib.packet.ipv4 import ipv4
+from pox.lib.packet.udp import udp
+from pox.lib.packet.tcp import tcp
+
+from util import buildTopo, getRouting
+from DemandEstimation import demand_estimation
+from threading import Timer, Lock
+log = core.getLogger()
+
+# Number of bytes to send for packet_ins
+MISS_SEND_LEN = 2000
+
+
+class Switch(EventMixin):
+    def __init__(self):
+        self.connection = None
+        self.dpid = None
+        self.ports = None
+
+    def connect(self, connection):
+        if self.dpid is None:
+            self.dpid = connection.dpid
+        assert self.dpid == connection.dpid
+        self.connection = connection
+
+    def send_packet_data(self, outport, data=None):
+        msg = of.ofp_packet_out(in_port=of.OFPP_NONE, data=data)
+        msg.actions.append(of.ofp_action_output(port=outport))
+        self.connection.send(msg)
+
+    def send_packet_bufid(self, outport, buffer_id=-1):
+        msg = of.ofp_packet_out(in_port=of.OFPP_NONE)
+        msg.actions.append(of.ofp_action_output(port=outport))
+        msg.buffer_id = buffer_id
+        self.connection.send(msg)
+
+    def install(self, port, match, buf=-1, deleteFlow=False, idle_timeout=0):
+        msg = of.ofp_flow_mod()
+        msg.match = match
+        msg.idle_timeout = idle_timeout
+        msg.actions.append(of.ofp_action_output(port=port))
+        if deleteFlow:
+            msg.command = of.OFPFC_DELETE
+        #msg.buffer_id = buf
+        msg.flags = of.OFPFF_SEND_FLOW_REM
+
+        self.connection.send(msg)
+
+
+class HController(EventMixin):
+    def __init__(self, t, r, bw):
[…]

### 2017-11-21 Iroko/hedera/pox/pox/datapaths/pcap_switch.py

@@ -1,0 +1,271 @@
+# Copyright 2013 James McCauley
+#
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at:
+#
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""
+Software switch with PCap ports
+
+Example:
+./pox.py --no-openflow datapaths.pcap_switch --address=localhost
+"""
+
+from pox.core import core
+from pox.datapaths import do_launch
+from pox.datapaths.switch import SoftwareSwitchBase, OFConnection
+from pox.datapaths.switch import ExpireMixin
+import pox.lib.pxpcap as pxpcap
+from Queue import Queue
+from threading import Thread
+import pox.openflow.libopenflow_01 as of
+from pox.lib.packet import ethernet
+import logging
+
+log = core.getLogger()
+
+DEFAULT_CTL_PORT = 7791
+
+_switches = {}
+
+def _do_ctl (event):
+  r = _do_ctl2(event)
+  if r is None:
+    r = "Okay."
+  event.worker.send(r + "\n")
+
+def _do_ctl2 (event):
+  def errf (msg, *args):
+    raise RuntimeError(msg % args)
+
+  args = event.args
+
+  def ra (low, high = None):
+    if high is None: high = low
+    if len(args) < low or len(args) > high:
+      raise RuntimeError("Wrong number of arguments")
+    return False
+
+  try:
+    if event.first == "add-port":
+      ra(1,2)
+      if len(event.args) == 1 and len(_switches) == 1:
+        sw = _switches[_switches.keys()[0]]
+        p = args[0]
+      else:
+        ra(2)
+        if event.args[0] not in _switches:
+          raise RuntimeError("No such switch")
+        sw = _switches[event.args[0]]
+        p = args[1]
+      sw.add_interface(p, start=True, on_error=errf)
+    elif event.first == "del-port":
+      ra(1,2)
+      if len(event.args) == 1:
+        for sw in _switches.values():
+          for p in sw.ports:
+            if p.name == event.args[0]:
[…]

### 2017-11-21 Iroko/iroko.py

@@ -1,28 +1,27 @@
+import os
+import logging
+import multiprocessing
+
 from mininet.net import Mininet
-from mininet.node import Controller, RemoteController
+from mininet.node import RemoteController
 from mininet.cli import CLI
-from mininet.log import setLogLevel, info, warn, error, debug
-from mininet.link import Link, Intf, TCLink
-from mininet.topo import Topo
-from mininet.util import dumpNodeConnections
-from mininet.examples.consoles import ConsoleApp
 
 from time import sleep
-import multiprocessing
 from mininet.node import OVSKernelSwitch, CPULimitedHost
 from mininet.util import custom
+from mininet.log import setLogLevel, info, warn, error, debug
 
 from subprocess import Popen, PIPE
 from argparse import ArgumentParser
 from monitor.monitor import monitor_devs_ng
-from monitor.monitor import monitor_qlen
 from multiprocessing import Process
 from mininet.term import makeTerm
+from mininet.link import TCLink
+from hedera.DCTopo import FatTreeTopo
 
 import topo_ecmp
 import topo_non_block
-import os
-import logging
+
 
 parser = ArgumentParser(description="Iroko Parser")
 
@@ -95,8 +94,9 @@ def trafficGen(args, hosts, net):
     monitor.start()
 
     sleep(args.time)
-
+    info('*** Stopping monitor\n')
     monitor.terminate()
+    os.system("killall bwm-ng")
 
     info('*** Stopping load-generators\n')
     for h in hosts:
@@ -143,7 +143,7 @@ def traffic_generation(net, topo, duration):
 
     # 4. Shut down.
     monitor.terminate()
-    os.system('killall bwm-ng')
+    Popen("killall -9 top bwm-ng", shell=True).wait()
     os.system('killall iperf')
 
 
@@ -186,14 +186,14 @@ def FatTreeTest(args, controller):
 
     net, topo = topo_ecmp.createECMPTopo(4, 2)
     c0 = RemoteController('c0', ip='127.0.0.1', port=6653)
-    ovs_v = 13 # default value
-    is_ecmp = True # default value
+    ovs_v = 13  # default value
+    is_ecmp = True  # default value
     if controller == "Iroko":
         makeTerm(c0, cmd="./ryu/bin/ryu-manager --observe-links --ofp-tcp-listen-port 6653 network_monitor.py")
 
[…]

### 2017-11-21 Iroko/plot_rate.py

@@ -8,38 +8,39 @@ import numpy as np
 
 parser = argparse.ArgumentParser()
 parser.add_argument('--input', '-f', dest='files', required=True, help='Input rates')
-parser.add_argument('--out', '-o', dest='out', default=None, 
-        help="Output png file for the plot.")
+parser.add_argument('--out', '-o', dest='out', default=None,
+                    help="Output png file for the plot.")
 
-args= parser.parse_args()
+args = parser.parse_args()
 
 
 ''' Output of bwm-ng has the following format:
     unix_timestamp;iface_name;bytes_out;bytes_in;bytes_total;packets_out;packets_in;packets_total;errors_out;errors_in
     '''
 
-# traffics=['stag_prob_0_2_3_data', 'stag_prob_1_2_3_data', 'stag_prob_2_2_3_data',
-#         'stag_prob_0_5_3_data','stag_prob_1_5_3_data','stag_prob_2_5_3_data','stride1_data', 
-#         'stride2_data', 'stride4_data', 'stride8_data', 'random0_data', 'random1_data', 'random2_data', 
-#         'random0_bij_data', 'random1_bij_data','random2_bij_data', 'random_2_flows_data', 
-#         'random_3_flows_data', 'random_4_flows_data','hotspot_one_to_one_data']
+traffics = ['stag_prob_0_2_3_data', 'stag_prob_1_2_3_data', 'stag_prob_2_2_3_data',
+                'stag_prob_0_5_3_data', 'stag_prob_1_5_3_data', 'stag_prob_2_5_3_data', 'stride1_data',
+                'stride2_data', 'stride4_data', 'stride8_data', 'random0_data', 'random1_data', 'random2_data',
+                'random0_bij_data', 'random1_bij_data', 'random2_bij_data', 'random_2_flows_data',
+                'random_3_flows_data', 'random_4_flows_data', 'hotspot_one_to_one_data']
 
-# labels=['stag0(0.2,0.3)', 'stag1(0.2,0.3)', 'stag2(0.2,0.3)', 'stag0(0.5,0.3)',
-#         'stag1(0.5,0.3)', 'stag2(0.5,0.3)', 'stride1', 'stride2',
-#         'stride4','stride8','rand0', 'rand1', 'rand2','randbij0',
-#         'randbij1','randbij2','randx2','randx3','randx4','hotspot']
[…]

### 2017-11-21 Iroko/run_experiment.sh

@@ -23,17 +23,17 @@ done
 
 for f in $INPUT_FILES;
 do
-        input_file=$INPUT_DIR/$f
+        input_file=hedera/$INPUT_DIR/$f
         pref="fattree-iroko"
         out_dir=$OUTPUT_DIR/$pref/$f
         sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --iroko
 done
 
-# for f in $INPUT_FILES;
-# do
-#         input_file=$INPUT_DIR/$f
-#         pref="fattree-hedera"
-#         out_dir=$OUTPUT_DIR/$pref/$f
-#         sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --hedera
-# done
+for f in $INPUT_FILES;
+do
+        input_file=$INPUT_DIR/$f
+        pref="fattree-hedera"
+        out_dir=$OUTPUT_DIR/$pref/$f
+        sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --hedera
+done
 

### 2017-11-21 Iroko/topo_ecmp.py

@@ -34,35 +34,6 @@ import os
 MAX_QUEUE = 1
 DENSITY = 2
 
-class FatNode(object):
-
-    def __init__(self, pod=0, sw=0, host=0, name=None, dpid=None):
-        ''' Create FatTreeNode '''
-        if dpid:
-            self.pod = (dpid & 0xff0000) >> 16
-            self.sw = (dpid & 0xff00) >> 8
-            self.host = (dpid & 0xff)
-            self.dpid = dpid
-        else:
-            if name:
-                pod, sw, host = [int(s) for s in name.split('h')]
-
-            self.pod = pod
-            self.sw = sw
-            self.host = host
-            self.dpid = (pod << 16) + (sw << 8) + host
-
-    def name_str(self):
-        ''' Return name '''
-        return "%ih%ih%i" % (self.pod, self.sw, self.host)
-
-    def ip_str(self):
-        ''' Return IP address '''
-        return "10.%i.%i.%i" % (self.pod, self.sw, self.host)
-
-    def mac_str(self):
-        ''' Return MAC address '''
-        return "00:00:00:%02x:%02x:%02x" % (self.pod, self.sw, self.host)
 
 class Fattree(Topo):
     """
@@ -80,7 +51,6 @@ class Fattree(Topo):
         self.iAggLayerSwitch = k * k / 2
         self.iEdgeLayerSwitch = k * k / 2
         self.iHost = self.iEdgeLayerSwitch * DENSITY
-        self.node_gen = FatNode
 
         # Init Topo
         Topo.__init__(self)

### 2017-11-21 Iroko/topo_non_block.py

@@ -117,16 +117,25 @@ def install_proactive(net, topo):
     """
             Install proactive flow entries for the switch.
     """
+    hostlist = []
+    for k in range(len(topo.HostList)):
+        hostlist.append(net.get(topo.HostList[k]))
     for sw in topo.CoreSwitchList:
-        for i in range(1, topo.iHost + 1):
+        i = 1
+        j = 1
+        for k in range(1, topo.iHost + 1):
             cmd = "ovs-ofctl add-flow %s -O OpenFlow13 \
                 'table=0,idle_timeout=0,hard_timeout=0,priority=40,arp, \
-                nw_dst=10.0.0.%d,actions=output:%d'" % (sw, i, i)
+                nw_dst=10.%d.0.%d,actions=output:%d'" % (sw, i, j, k)
             os.system(cmd)
             cmd = "ovs-ofctl add-flow %s -O OpenFlow13 \
                 'table=0,idle_timeout=0,hard_timeout=0,priority=40,ip, \
-                nw_dst=10.0.0.%d,actions=output:%d'" % (sw, i, i)
+                nw_dst=10.%d.0.%d,actions=output:%d'" % (sw, i, j, k)
             os.system(cmd)
+            j += 1
+            if j == 3:
+                j = 1
+                i += 1
 
 def configureTopo(net, topo):
     # Set OVS's protocol as OF13.

### 2017-11-21 Deprecated/Hedera/run_experiment.sh
Push run script fixes

@@ -3,21 +3,21 @@ OUTPUT_DIR=results
 INPUT_FILES='stag_prob_0_2_3_data stag_prob_1_2_3_data stag_prob_2_2_3_data stag_prob_0_5_3_data stag_prob_1_5_3_data stag_prob_2_5_3_data stride1_data stride2_data stride4_data stride8_data random0_data random1_data random2_data random0_bij_data random1_bij_data random2_bij_data random_2_flows_data random_3_flows_data random_4_flows_data hotspot_one_to_one_data'
 DURATION=50
 
-# for f in $INPUT_FILES;
-# do
-#         input_file=$INPUT_DIR/$f
-#         pref="nonblocking"
-#         out_dir=$OUTPUT_DIR/$pref/$f
-#         sudo python hedera.py -i $input_file -d $out_dir -p 0.03 -t $DURATION -n
-# done
+for f in $INPUT_FILES;
+do
+        input_file=$INPUT_DIR/$f
+        pref="nonblocking"
+        out_dir=$OUTPUT_DIR/$pref/$f
+        sudo python hedera.py -i $input_file -d $out_dir -p 0.03 -t $DURATION -n
+done
 
-# for f in $INPUT_FILES;
-# do
-#         input_file=$INPUT_DIR/$f
-#         pref="fattree-ecmp"
-#         out_dir=$OUTPUT_DIR/$pref/$f
-#         sudo python hedera.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --ecmp
-# done
+for f in $INPUT_FILES;
+do
+        input_file=$INPUT_DIR/$f
+        pref="fattree-ecmp"
+        out_dir=$OUTPUT_DIR/$pref/$f
+        sudo python hedera.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --ecmp
+done
 
 for f in $INPUT_FILES;
 do

### 2017-11-21 Iroko/run_experiment.sh

@@ -23,7 +23,7 @@ done
 
 for f in $INPUT_FILES;
 do
-        input_file=hedera/$INPUT_DIR/$f
+        input_file=$INPUT_DIR/$f
         pref="fattree-iroko"
         out_dir=$OUTPUT_DIR/$pref/$f
         sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --iroko
@@ -31,7 +31,7 @@ done
 
 for f in $INPUT_FILES;
 do
-        input_file=$INPUT_DIR/$f
+        input_file=hedera/$INPUT_DIR/$f
         pref="fattree-hedera"
         out_dir=$OUTPUT_DIR/$pref/$f
         sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --hedera

### 2017-11-22 Iroko/iroko.py
More benchmark fixes...

@@ -233,8 +233,6 @@ def HederaTest(args):
                      controller=None)
     net.addController(c0)
     net.start()
-    CLI(net)
-    exit(1)
     # wait for the switches to connect to the controller
     info('** Waiting for switches to connect to the controller\n')
     sleep(5)

### 2017-11-22 Iroko/plot_rate.py

@@ -19,10 +19,10 @@ args = parser.parse_args()
     '''
 
 traffics = ['stag_prob_0_2_3_data', 'stag_prob_1_2_3_data', 'stag_prob_2_2_3_data',
-                'stag_prob_0_5_3_data', 'stag_prob_1_5_3_data', 'stag_prob_2_5_3_data', 'stride1_data',
-                'stride2_data', 'stride4_data', 'stride8_data', 'random0_data', 'random1_data', 'random2_data',
-                'random0_bij_data', 'random1_bij_data', 'random2_bij_data', 'random_2_flows_data',
-                'random_3_flows_data', 'random_4_flows_data', 'hotspot_one_to_one_data']
+            'stag_prob_0_5_3_data', 'stag_prob_1_5_3_data', 'stag_prob_2_5_3_data', 'stride1_data',
+            'stride2_data', 'stride4_data', 'stride8_data', 'random0_data', 'random1_data', 'random2_data',
+            'random0_bij_data', 'random1_bij_data', 'random2_bij_data', 'random_2_flows_data',
+            'random_3_flows_data', 'random_4_flows_data', 'hotspot_one_to_one_data']
 
 labels = ['stag0(0.2,0.3)', 'stag1(0.2,0.3)', 'stag2(0.2,0.3)', 'stag0(0.5,0.3)',
           'stag1(0.5,0.3)', 'stag2(0.5,0.3)', 'stride1', 'stride2',
@@ -74,31 +74,34 @@ def plot_results(args):
 
     bb = {'nonblocking': [], 'hedera': [], 'iroko': [], 'ecmp': []}
 
-    sw = '4h1h1'
+    # sw = '4h1h1'
+    sw = '1001'
     for t in traffics:
-        print "Nonblocking:", t
+        print("Nonblocking:", t)
         input_file = args.files + '/nonblocking/%s/rate.txt' % t
         vals = get_bisection_bw(input_file, sw)
         bb['nonblocking'].append(vals / fbb)
 
-    sw = '[0-3]h[0-1]h1'
+    # sw = '[0-3]h[0-1]h1'
+    sw = '[1-3]00[1-9]'
     for t in traffics:
-        print "ECMP:", t
+        print("ECMP:", t)
         input_file = args.files + '/fattree-ecmp/%s/rate.txt' % t
         vals = get_bisection_bw(input_file, sw)
         bb['ecmp'].append(vals / fbb / 2)
 
[…]

### 2017-11-23 Deprecated/Hedera/plot_rate.py
Even more benchmarking changes.

@@ -75,6 +75,7 @@ def plot_results(args):
         print "Nonblocking:", t
         input_file = args.files + '/nonblocking/%s/rate.txt' % t    
         vals = get_bisection_bw(input_file, sw)
+        print (vals)
         bb['nonblocking'].append(vals/fbb)
 
     sw = '[0-3]h[0-1]h1'
@@ -82,12 +83,14 @@ def plot_results(args):
         print "ECMP:", t
         input_file = args.files + '/fattree-ecmp/%s/rate.txt' % t
         vals = get_bisection_bw(input_file, sw)
+        print (vals)
         bb['ecmp'].append(vals/fbb/2)
    
     for t in traffics:
        print "Hedera:", t
        input_file = args.files + '/fattree-hedera/%s/rate.txt' % t
        vals = get_bisection_bw(input_file, sw)
+       print (vals)
        bb['hedera'].append(vals/fbb/2)
 
 

### 2017-11-23 Iroko/iroko.py

@@ -22,6 +22,7 @@ from hedera.DCTopo import FatTreeTopo
 import topo_ecmp
 import topo_non_block
 
+MAX_QUEUE = 50
 
 parser = ArgumentParser(description="Iroko Parser")
 
@@ -167,10 +168,14 @@ def pingTest(net):
 def clean():
     ''' Clean any the running instances of POX '''
 
-    p = Popen("ps aux | grep 'pox' | awk '{print $2}'",
+    p_pox = Popen("ps aux | grep 'pox' | awk '{print $2}'",
               stdout=PIPE, shell=True)
-    p.wait()
-    procs = (p.communicate()[0]).split('\n')
+    p_pox.wait()
+    procs = (p_pox.communicate()[0]).split('\n')
+    p_ryu = Popen("ps aux | grep 'ryu' | awk '{print $2}'",
+              stdout=PIPE, shell=True)
+    p_ryu.wait()
+    procs.extend((p_ryu.communicate()[0]).split('\n'))
     for pid in procs:
         try:
             pid = int(pid)
@@ -190,10 +195,12 @@ def FatTreeTest(args, controller):
     is_ecmp = True  # default value
     if controller == "Iroko":
         makeTerm(c0, cmd="./ryu/bin/ryu-manager --observe-links --ofp-tcp-listen-port 6653 network_monitor.py")
+        # Popen("./ryu/bin/ryu-manager --observe-links --ofp-tcp-listen-port 6653 network_monitor.py", shell=True)
 
-        #c0.cmdPrint('xterm  -T \"./ryu/bin/ryu-manager --observe-links network_monitor.py\" &')
     elif controller == "HController":
-        makeTerm(c0, cmd="hedera/pox/pox.py HController --topo=ft,4 --routing=ECMP")
+        # makeTerm(c0, cmd="hedera/pox/pox.py HController --topo=ft,4 --routing=ECMP")
+        Popen("hedera/pox/pox.py HController --topo=ft,4 --routing=ECMP", shell=True)
+
         ovs_v = 10
         is_ecmp = False
     net.addController(c0)
@@ -207,7 +214,7 @@ def FatTreeTest(args, controller):
     sleep(2)
     hosts = net.hosts
     trafficGen(args, hosts, net)
-
+    CLI(net)
     net.stop()
 
 
[…]

### 2017-11-23 Iroko/network_monitor.py

@@ -536,18 +536,18 @@ class NetworkMonitor(app_manager.RyuApp):
         print("Current Average Utilization: %f" % curr_avg_util)
         print("Current Average Utilization ALT: %f" % curr_avg_util_alt)
 
-        # if _type == 'port':
-        #     print('\ndatapath  port     '
-        #           ' rx-dropped '' tx-dropped '
-        #           ' port-bw(Kb/s)  port-speed(b/s)  port-freebw(Kb/s) ')
-        #     print('--------  --------  ' '----------  ---------- '' -------------  --------------- ' ' -----------------')
-        #     _format = '%8d  %8x  %9d   %9d   %10d  %15.1f  %17.1f'
-        #     for dpid in sorted(bodys.keys()):
-        #         for stat in sorted(bodys[dpid], key=attrgetter('port_no')):
-        #             # if stat.port_no == ofproto_v1_3.OFPP_LOCAL:
-        #             print(_format % (
-        #                 dpid, stat.port_no,
-        #                 stat.rx_dropped, stat.tx_dropped, 10000,
-        #                 abs(self.port_speed[(dpid, stat.port_no)][-1] * 8),
-        #                 self.free_bandwidth[dpid][stat.port_no]))
-        #     print
+        if _type == 'port':
+            print('\ndatapath  port     '
+                  ' rx-dropped '' tx-dropped '
+                  ' port-bw(Kb/s)  port-speed(b/s)  port-freebw(Kb/s) ')
+            print('--------  --------  ' '----------  ---------- '' -------------  --------------- ' ' -----------------')
+            _format = '%8d  %8x  %9d   %9d   %10d  %15.1f  %17.1f'
+            for dpid in sorted(bodys.keys()):
+                for stat in sorted(bodys[dpid], key=attrgetter('port_no')):
+                    # if stat.port_no == ofproto_v1_3.OFPP_LOCAL:
+                    print(_format % (
+                        dpid, stat.port_no,
+                        stat.rx_dropped, stat.tx_dropped, 10000,
+                        abs(self.port_speed[(dpid, stat.port_no)][-1] * 8),
+                        self.free_bandwidth[dpid][stat.port_no]))
+            print

### 2017-11-23 Iroko/plot_rate.py

@@ -80,6 +80,7 @@ def plot_results(args):
         print("Nonblocking:", t)
         input_file = args.files + '/nonblocking/%s/rate.txt' % t
         vals = get_bisection_bw(input_file, sw)
+        print (vals)
         bb['nonblocking'].append(vals / fbb)
 
     # sw = '[0-3]h[0-1]h1'
@@ -88,12 +89,14 @@ def plot_results(args):
         print("ECMP:", t)
         input_file = args.files + '/fattree-ecmp/%s/rate.txt' % t
         vals = get_bisection_bw(input_file, sw)
+        print (vals)
         bb['ecmp'].append(vals / fbb / 2)
 
     for t in traffics:
         print("Iroko:", t)
         input_file = args.files + '/fattree-iroko/%s/rate.txt' % t
         vals = get_bisection_bw(input_file, sw)
+        print (vals)
         bb['iroko'].append(vals / fbb / 2)
 
     sw = '[0-3]h[0-1]h1'
@@ -101,6 +104,7 @@ def plot_results(args):
         print("Hedera:", t)
         input_file = args.files + '/fattree-hedera/%s/rate.txt' % t
         vals = get_bisection_bw(input_file, sw)
+        print (vals)
         bb['hedera'].append(vals / fbb / 2)
 
     ind = np.arange(n_t)

### 2017-11-23 Iroko/run_experiment.sh

@@ -2,24 +2,24 @@ INPUT_DIR=inputs
 OUTPUT_DIR=results
 INPUT_FILES='stag_prob_0_2_3_data stag_prob_1_2_3_data stag_prob_2_2_3_data stag_prob_0_5_3_data stag_prob_1_5_3_data stag_prob_2_5_3_data stride1_data stride2_data stride4_data stride8_data random0_data random1_data random2_data random0_bij_data random1_bij_data random2_bij_data random_2_flows_data random_3_flows_data random_4_flows_data hotspot_one_to_one_data'
 #INPUT_FILES='stag_prob_0_2_3_data'
-DURATION=50
+DURATION=60
 
 sudo mn -c
-for f in $INPUT_FILES;
-do
-        input_file=$INPUT_DIR/$f
-        pref="nonblocking"
-        out_dir=$OUTPUT_DIR/$pref/$f
-        sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION -n
-done
+# for f in $INPUT_FILES;
+# do
+#         input_file=$INPUT_DIR/$f
+#         pref="nonblocking"
+#         out_dir=$OUTPUT_DIR/$pref/$f
+#         sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION -n
+# done
 
-for f in $INPUT_FILES;
-do
-        input_file=$INPUT_DIR/$f
-        pref="fattree-ecmp"
-        out_dir=$OUTPUT_DIR/$pref/$f
-        sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --ecmp
-done
+# for f in $INPUT_FILES;
+# do
+#         input_file=$INPUT_DIR/$f
+#         pref="fattree-ecmp"
+#         out_dir=$OUTPUT_DIR/$pref/$f
+#         sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --ecmp
+# done
 
 for f in $INPUT_FILES;
 do
@@ -29,11 +29,11 @@ do
         sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --iroko
 done
 
-for f in $INPUT_FILES;
-do
-        input_file=hedera/$INPUT_DIR/$f
-        pref="fattree-hedera"
-        out_dir=$OUTPUT_DIR/$pref/$f
[…]

### 2017-11-23 Iroko/topo_ecmp.py

@@ -31,7 +31,7 @@ from mininet.util import custom
 
 import os
 
-MAX_QUEUE = 1
+MAX_QUEUE = 50
 DENSITY = 2
 
 

### 2017-11-23 Iroko/topo_non_block.py

@@ -30,7 +30,7 @@ import sys
 parentdir = os.path.dirname(os.path.dirname(os.path.abspath(__file__)))
 sys.path.insert(0, parentdir)
 #import iperf_peers
-
+MAX_QUEUE = 50
 
 class NonBlocking(Topo):
     """
@@ -83,7 +83,7 @@ class NonBlocking(Topo):
         """
         for sw in self.CoreSwitchList:
             for host in self.HostList:
-                self.addLink(sw, host, bw=bw_h2c, max_queue_size=1000)   # use_htb=False
+                self.addLink(sw, host, bw=bw_h2c, max_queue_size=MAX_QUEUE)   # use_htb=False
 
     def set_ovs_protocol_13(self):
         """

### 2017-11-23 Iroko/iroko.py
Add interface to host mappings.

@@ -213,8 +213,8 @@ def FatTreeTest(args, controller):
     info('** Waiting for switches to connect to the controller\n')
     sleep(2)
     hosts = net.hosts
-    trafficGen(args, hosts, net)
     CLI(net)
+    trafficGen(args, hosts, net)
     net.stop()
 
 
@@ -290,7 +290,7 @@ if __name__ == '__main__':
     elif args.iroko:
         FatTreeTest(args, controller='Iroko')
     else:
-        info('**error** please specify either hedera, ecmp, or nonblocking\n')
+        info('**error** please specify either hedera, iroko, ecmp, or nonblocking\n')
     clean()
 
 

### 2017-11-23 Iroko/network_monitor.py

@@ -43,6 +43,11 @@ MAX_CAPACITY = 10000   # Max capacity of link
 # CONF = cfg.CONF
 weight = 'bw'
 
+i_h_map = {'3001-eth3': "192.168.10.1", '3001-eth4': "192.168.10.2", '3002-eth3': "192.168.10.3", '3002-eth4': "192.168.10.4",
+           '3003-eth3': "192.168.10.5", '3003-eth4': "192.168.10.6", '3004-eth3': "192.168.10.7", '3004-eth4': "192.168.10.8",
+           '3005-eth3': "192.168.10.9", '3005-eth4': "192.168.10.10", '3006-eth3': "192.168.10.11", '3006-eth4': "192.168.10.12",
+           '3007-eth3': "192.168.10.13", '3007-eth4': "192.168.10.14", '3008-eth3': "192.168.10.15", '3008-eth4': "192.168.10.16", }
+
 
 class NetworkMonitor(app_manager.RyuApp):
     """
@@ -93,7 +98,7 @@ class NetworkMonitor(app_manager.RyuApp):
             if self.stats['flow'] or self.stats['port']:
                 # self.show_stat('flow')
                 self.show_stat('port')
-                #hub.sleep(0.01)
+                # hub.sleep(0.01)
 
     # def _save_bw_graph(self):
     #     """
@@ -518,6 +523,7 @@ class NetworkMonitor(app_manager.RyuApp):
                     #avg_util_new += float(output3)
                     i += 1
                     # print ("%s %d %d " % (dpid, stat.port_no, i))
+
         iface_list = self._get_active_ports()
         bandwidths = self._get_bandwidths(iface_list)
         avg_util_new = sum(bandwidths.itervalues())

### 2017-11-23 Iroko/run_experiment.sh

@@ -5,21 +5,21 @@ INPUT_FILES='stag_prob_0_2_3_data stag_prob_1_2_3_data stag_prob_2_2_3_data stag
 DURATION=60
 
 sudo mn -c
-# for f in $INPUT_FILES;
-# do
-#         input_file=$INPUT_DIR/$f
-#         pref="nonblocking"
-#         out_dir=$OUTPUT_DIR/$pref/$f
-#         sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION -n
-# done
+for f in $INPUT_FILES;
+do
+        input_file=$INPUT_DIR/$f
+        pref="nonblocking"
+        out_dir=$OUTPUT_DIR/$pref/$f
+        sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION -n
+done
 
-# for f in $INPUT_FILES;
-# do
-#         input_file=$INPUT_DIR/$f
-#         pref="fattree-ecmp"
-#         out_dir=$OUTPUT_DIR/$pref/$f
-#         sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --ecmp
-# done
+for f in $INPUT_FILES;
+do
+        input_file=$INPUT_DIR/$f
+        pref="fattree-ecmp"
+        out_dir=$OUTPUT_DIR/$pref/$f
+        sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --ecmp
+done
 
 for f in $INPUT_FILES;
 do
@@ -29,11 +29,11 @@ do
         sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --iroko
 done
 
-# for f in $INPUT_FILES;
-# do
-#         input_file=hedera/$INPUT_DIR/$f
-#         pref="fattree-hedera"
-#         out_dir=$OUTPUT_DIR/$pref/$f
-#         sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --hedera
-# done
+for f in $INPUT_FILES;
+do
+        input_file=hedera/$INPUT_DIR/$f
+        pref="fattree-hedera"
+        out_dir=$OUTPUT_DIR/$pref/$f
+        sudo python iroko.py -i $input_file -d $out_dir -p 0.03 -t $DURATION --hedera
+done
 

### 2017-11-23 Iroko/topo_non_block.py

@@ -30,7 +30,7 @@ import sys
 parentdir = os.path.dirname(os.path.dirname(os.path.abspath(__file__)))
 sys.path.insert(0, parentdir)
 #import iperf_peers
-MAX_QUEUE = 50
+MAX_QUEUE = 1000
 
 class NonBlocking(Topo):
     """

### 2017-11-23 LearningController/HederaTopologyProblem.py
Test code for a controller and neuralnetwork

@@ -1,0 +1,96 @@
+import numpy as np
+import torch.nn
+import torch
+from  torch.autograd import Variable
+import time
+
+samples = 6631171 - 20  #14973108 - 60
+indexes = np.arange(samples);
+np.random.shuffle(indexes)
+trainset = indexes[0:int(np.floor(samples * .90))]
+testset = indexes[int(np.ceil(samples*.90)):-1]
+
+data = np.load('data.npy')
+#data = np.random.rand(samples, 15)
+
+data = data[0:samples, :]
+
+trainset = data[trainset,:]
+testset = data[testset,:]
+
+## Set-up that sweet sweet neural network learning
+network = torch.nn.Sequential(
+    torch.nn.Linear(14, 8, True),
+    torch.nn.ReLU(),
+    torch.nn.Linear(8, 4,True) ,
+    torch.nn.ReLU(),
+    torch.nn.Linear(4,83),
+    torch.nn.LogSoftmax()) #1))
+
+optimizer = torch.optim.SGD(network.parameters(), lr = 1e-7)
+
+#criterion = torch.nn.MSELoss()
+criterion = torch.nn.NLLLoss()
+#for sample in trainset:
+for i in range(0, trainset.size / 15, 32):
+    sample = trainset[i:i+33,:] 
+    network.zero_grad()
+    #label = Variable(torch.from_numpy(np.array([sample[:,0]])).float())
+    #features = Variable(torch.from_numpy(sample[:,1:15]).float())
+    label = Variable(torch.from_numpy(np.array([sample[:,0]])).long())
+    features = Variable(torch.from_numpy(sample[:,1:15]).float())
+
+
+  #  print(features)
+    prediction = network(features)
+    loss = criterion(prediction, label.view(label.size()[1]))
+    totalLoss = torch.sum(loss.data, 0) / 33
+    #print(totalLoss)
+    #print('samples: ' + str(i) + ',' + str(i + 32) + ',' + 'average training loss: ' + str(totalLoss[0]) )
+    loss.backward()
+    optimizer.step()
+
+
+total_loss = 0
+
+label = Variable(torch.from_numpy(np.array([testset[:,0]])).long())
+features = Variable(torch.from_numpy(testset[:,1:15]).float())
+
+
+prediction = network(features)
+loss = criterion(prediction, label.view(label.size()[1]))
+total_loss = torch.sum(loss.data ) 
+
+"""
+for sample in testset:
+    #label = Variable(torch.from_numpy(np.array([sample[0]])).float())
[…]

### 2017-11-23 LearningController/LearningController.py

@@ -3,9 +3,9 @@ from  torch.autograd import Variable
 import torch
 import math
 import random
-
+import numpy as np
 class LearningController:
-    def __init__(self, inputs, actions, numNeuron1, numNeuron2, alpha, gamma, epsilon = .9, decay=.001):
+    def __init__(self, inputs, actions, numNeuron1, numNeuron2, alpha, gamma, epsilon = .9, decay=.001, filepath = None):
         self.model = torch.nn.Sequential(
                 torch.nn.Linear(inputs + actions, numNeuron1, True ), 
                 torch.nn.ReLU(),
@@ -18,26 +18,35 @@ class LearningController:
         self.actionlength = actions
         self.prevAction = torch.zeros(actions);
         self.prevState = torch.zeros(inputs);
-        self.optimizer = torch.optim.SGD(self.model.parameters(), lr = 0.01)
+        self.optimizer = torch.optim.SGD(self.model.parameters(), lr = 1e-6)
         self.epsilon = epsilon
         self.decay = decay
+        if (filepath != None):
+            self.model.load_state_dict(torch.load(filepath))
+
+
 
     def PerformAction(self):
         if(random.random() < self.epsilon):
-            self.epsilon = max(0, self.epsilon - self.decay)
-            return torch.rand( self.actionlength) >= .5 
+            # makes sure minimum is 0
+            self.epsilon = max(.05, self.epsilon - self.decay)
+            action = np.random.random_integers(0, self.actionlength - 1) 
+            ret = torch.zeros(self.actionlength)
+            ret[action] = 1.0
+            return ret 
         return self.ReturnBestAction() 
 
     def ReturnBestAction(self):
-        bAction = torch.rand(self.actionlength)
+        bAction = torch.zeros(self.actionlength)
         state = Variable(torch.cat((self.prevState, bAction),0))
         
         bVal = self.model(state)
-        for i in range(0, int(math.pow(2, self.actionlength) / 8)):
-            curVal = self.model(state); 
-
-            action = torch.rand(self.actionlength) >= .5
+        for i in range(0, self.actionlength):
+             
+            action = torch.zeros(self.actionlength) 
+            action[i] = 1
             state = Variable(torch.cat((self.prevState, action.float()), 0))  
+            curVal = self.model(state);
             if((curVal > bVal).data[0] ):
                 curVal = bVal
                 bAction = action
[…]

### 2017-11-23 LearningController/ParseFiles.py

@@ -1,0 +1,52 @@
+import os
+import csv
+import numpy as np
+currLabel = 0 
+#data = np.zeros((14973108, 15)) 
+#0 - 83 should be 84 labels
+numL = 83
+data = np.zeros((6631171,numL + 14))
+indx = 0
+invalid = 0
+labels = dict()
+nextLabel = 0
+for dirname in os.listdir('./results'):
+    if( dirname != 'fattree-hedera'):
+        continue
+    currLabel = currLabel + 1
+    print(dirname)
+    for dirname2 in os.listdir('./results/' + dirname):
+        for filename in os.listdir('./results/' + dirname +  '/' + dirname2):
+            with open('./results/' + dirname +  '/' + dirname2 +'/' + filename, 'rb') as csvfile:
+                rates = csv.reader(csvfile, delimiter=',')
+                # for this file
+                for row in rates:
+                    features = []
+                    exclude = 0
+                    label = np.zeros(numL)
+                    #for each element in the row
+                    for item in row:
+                        if(exclude == 0):
+                            exclude = exclude + 1
+                            continue
+                        elif(exclude == 1):
+                            exclude = exclude + 1
+                            if(not(item in labels)):
+                                    labels[item] = nextLabel
+                                    nextLabel += 1
+                            label[labels[item]] = 1.0
+                        else:
+                            features.append(item)
+                    features = np.concatenate((label, np.array(features)))
+                    if(features.size != (numL + 14)):
+                        print(row)
+                        invalid += 1 
+                    else:
+                        data[indx, :] = features
+                        indx += 1
+    print(indx)
+
+print(nextLabel)
+print(data.size)
+print(invalid)
+#np.save('data', data)

### 2017-11-23 LearningController/TestScript.py

@@ -1,23 +1,51 @@
 from LearningController import LearningController
 import torch.nn 
+import torch
+import gym
 
-Agent = LearningController(10, 10, 10, 15, .1, .9)
 
+#(self, inputs, actions, numNeuron1, numNeuron2, alpha, gamma, epsilon = .9, decay=.001):
+numSimulation = 100000000
 
-inputs = torch.Tensor([1, 2, 3, 4, 5, 6, 7, 8, 9, 0])
-action = torch.Tensor([0, 0, 0, 0, 0, 0, 0, 0, 0, 0])
-reward = 1;
-oddStates = 0
-numSimulation = 1000000
+simulationLengths = torch.zeros(numSimulation);
+simulationCount = 0
+maxSeen = 0
+#Agent = LearningController(4, 2,8 , 3, .001, .9, epsilon = .9, decay = 1e-8 )
+Agent = LearningController(4, 2,8 , 3, .001, .1, 1e-8, './modelconfig')
+for params in Agent.model.parameters():
+    print(params)
+
+env = gym.make('CartPole-v0')
+env.reset()
+env.render()
+numIterations = 0
+sinceFailure = 0
 for i in range(1, numSimulation):
     
-    Agent.UpdateValueFunction(inputs, action.float(), reward)
-
     action = Agent.PerformAction()
-    inputs = inputs + action.float()
-    if (int(torch.sum(inputs)) % 2 != 0):
-        reward = 1
-        oddStates = oddStates + 1
+    val, a = torch.max(action, 0)
+    observation, reward, done, info = env.step(a[0]) 
+    env.render()    
+    inputs = torch.from_numpy(observation)
+    numIterations += 1
+    if( done ):
+        simulationLengths[simulationCount] = numIterations
+        simulationCount += 1
+        reward = -100
+        if(maxSeen < numIterations):
+            maxSeen = numIterations
+            print(maxSeen)
+        numIterations = 0
+        env.reset()
+        sinceFailure = 0
     else:
-        reward = -1 
-    print(float(oddStates ) / i )
+        reward = 1
+
+    if(i % 1000 == 0):
+        Agent.saveNetwork()
+
+    Agent.UpdateValueFunction(inputs.float(), action.float(), reward)
+ 
+torch.save(simulationLengths[0:simulationCount], './simulationCounts')
+Agent.saveNetwork()
+

### 2017-11-23 LearningController/pytorchExample.py

@@ -1,0 +1,70 @@
+import torch
+from torch.autograd import Variable
+import torch.nn as nn
+import torch.nn.functional as F
+
+class Net(nn.Module):
+    def __init__(self):
+        super(Net, self).__init__()
+        self.conv1 = nn.Conv2d(1,6,5)
+        self.conv2 = nn.Conv2d(6, 16,5)
+        # an affine operation: y = Wx + b
+        self.fc1 = nn.Linear(16 * 5 * 5, 120)
+        self.fc2 = nn.Linear(120, 84)
+        self.fc3 = nn.Linear(84, 10)
+
+    def forward(self, x):
+        x = F.max_pool2d(F.relu(self.conv1(x)), (2 , 2))
+
+        x = F.max_pool2d(F.relu(self.conv2(x)), 2)
+        x = x.view(-1, self.num_flat_features(x))
+        x = F.relu(self.fc1(x))
+        x = F.relu(self.fc2(x))
+        x = self.fc3(x)
+        return x
+
+
+    def num_flat_features(self, x):
+        size = x.size()[1:]
+        num_features = 1
+        for s in size:
+            num_features *= s
+
+        return num_features
+
+
+net = Net()
+print(net)
+params = list(net.parameters())
+print(len(params))
+print(params[0].size())
+
+input = Variable(torch.randn(1, 1, 32, 32))
+out = net(input)
+print(out)
+
+net.zero_grad()
+out.backward(torch.randn(1,10))
+
+
+output = net(input)
+target = Variable(torch.arange(1,11))
+criterion = nn.MSELoss()
+
+loss = criterion(output, target)
+print(loss)
+
+print(loss.grad_fn)  # MSELoss
+print(loss.grad_fn.next_functions[0][0])  # Linear
+print(loss.grad_fn.next_functions[0][0].next_functions[0][0])  # ReLU
+
+
+net.zero_grad()     # zeroes the gradient buffers of all parameters
+
+print('conv1.bias.grad before backward')
+print(net.conv1.bias.grad)
+
+loss.backward()
+
+print('conv1.bias.grad after backward')
+print(net.conv1.bias.grad)

### 2017-11-23 Iroko/iroko.py
Fix the plotting and remove the accidental CLI command.

@@ -213,7 +213,6 @@ def FatTreeTest(args, controller):
     info('** Waiting for switches to connect to the controller\n')
     sleep(2)
     hosts = net.hosts
-    CLI(net)
     trafficGen(args, hosts, net)
     net.stop()
 

### 2017-11-23 Iroko/plot_rate.py

@@ -40,17 +40,14 @@ def get_bisection_bw(input_file, pat_iface):
 
     rate = {}
     column = 2
-
     for row in data:
         try:
             ifname = row[1]
         except:
             break
-
         if ifname not in ['eth0', 'lo']:
             if not rate.has_key(ifname):
                 rate[ifname] = []
-
             try:
                 rate[ifname].append(float(row[column]) * 8.0 / (1 << 20))
             except:
@@ -60,7 +57,7 @@ def get_bisection_bw(input_file, pat_iface):
         if pat_iface.match(k):
             avg_rate = avg(rate[k][10:-10])
             vals.append(avg_rate)
-
+            print k
     return fsum(vals)
 
 
@@ -84,7 +81,7 @@ def plot_results(args):
         bb['nonblocking'].append(vals / fbb)
 
     # sw = '[0-3]h[0-1]h1'
-    sw = '[1-3]00[1-9]'
+    sw = '300[1-9]'
     for t in traffics:
         print("ECMP:", t)
         input_file = args.files + '/fattree-ecmp/%s/rate.txt' % t
@@ -108,9 +105,9 @@ def plot_results(args):
         bb['hedera'].append(vals / fbb / 2)
 
     ind = np.arange(n_t)
-    width = 0.2
+    width = 0.15
     fig = plt.figure(1)
-    fig.set_size_inches(18.5, 6.5)
+    fig.set_size_inches(8.5, 6.5)
 
     for i in range(num_plot):
         fig.set_size_inches(24, 12)
@@ -131,13 +128,12 @@ def plot_results(args):
         p2 = plt.bar(ind + 3.5 * width, bb['hedera'][i * n_t:(i + 1) * n_t], width=width, color='green')
 
         # FatTree + Iroko
-        p2 = plt.bar(ind + 2.5 * width, bb['iroko'][i * n_t:(i + 1) * n_t], width=width, color='yellow')
+        p3 = plt.bar(ind + 2.5 * width, bb['iroko'][i * n_t:(i + 1) * n_t], width=width, color='yellow')
 
         # FatTree + ECMP
-        p3 = plt.bar(ind + 1.5 * width, bb['ecmp'][i * n_t:(i + 1) * n_t], width=width, color='brown')
[…]

### 2017-11-23 Iroko/hedera/pox/pox/lib/pxpcap/strip_openflow.py
Add missing lib files...

@@ -1,0 +1,93 @@
+# Copyright 2012,2013 James McCauley
+#
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at:
+#
+#     http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""
+A hacky tool to grab packet in/out data from OpenFlow traffic.
+
+Assumes packets are 1:1 with OF messages (as if captured using the
+openflow.debug component).
+
+ --infile=<filename>   Input file
+ --outfile=<filename>  Output file
+ --out-only            Don't include packet_ins
+ --in-only             Don't include packet_outs
+ --openflow-port=<num> Specify OpenFlow TCP port
+"""
+
+#TODO: Clean this up, follow multiple control traffic streams, decode
+#      TCP, etc.
+
+from pox.core import core
+import pox.openflow.libopenflow_01 as of
+import pox.lib.packet as pkt
+from pox.lib.util import dpidToStr
+import pox.lib.pxpcap.parser as pxparse
+import pox.lib.pxpcap.writer as pxwriter
+
+log = core.getLogger()
+
+from pox.lib.pxpcap.writer import PCapRawWriter
+
+_writer = None
+_of_port = 6653
+_in_only = False
+_out_only = False
+
+_pis = 0
+_pos = 0
+
+def pi_cb (data, parser):
+  global _pis, _pos
+  packet = pkt.ethernet(data)
+  if packet.find('tcp'):
+    if packet.find('tcp').dstport == _of_port or \
+       packet.find('tcp').srcport == _of_port:
+      p = packet.find('tcp').payload
+      assert p[0] == '\x01'
+      t = ord(p[1])
+      packet_length = ord(p[2]) << 8 | ord(p[3])
+      if packet_length != len(p):
+        log.error("%s != %s" % (packet_length, len(p)))
+      if t == of.OFPT_PACKET_IN:
+        if _out_only: return
+        l,p = of.ofp_packet_in.unpack_new(p)
+        _pis += 1
[…]

### 2017-11-23 Iroko/ryu/ryu/lib/packet/zebra.py

@@ -1,0 +1,3483 @@
+# Copyright (C) 2017 Nippon Telegraph and Telephone Corporation.
+#
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#    http://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or
+# implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""
+Zebra protocol parser/serializer
+
+Zebra Protocol is used to communicate with the zebra daemon.
+"""
+
+import abc
+import socket
+import struct
+import logging
+from distutils.version import LooseVersion
+
+import netaddr
+import six
+
+from ryu import flags as cfg_flags  # For loading 'zapi' option definition
+from ryu.cfg import CONF
+from ryu.lib import addrconv
+from ryu.lib import ip
+from ryu.lib import stringify
+from ryu.lib import type_desc
+from . import packet_base
+from . import bgp
+from . import safi as packet_safi
+
+
+LOG = logging.getLogger(__name__)
+
[…]

### 2017-11-23 Iroko/hedera/pox/ext/DCController.py
Also add ext/* ...

@@ -1,0 +1,1 @@
+../../DCController.py

### 2017-11-23 Iroko/hedera/pox/ext/HController.py

@@ -1,0 +1,1 @@
+../../HController.py

### 2017-12-06 Iroko/LearningAgent.py
Switched to learning version of agent

@@ -58,7 +58,6 @@ class SARSA:
 
         Qcurr = self.model(state)
         Qprev = self.model(stateprev)
-        Qupdate = Qprev.data + self.alpha * (reward + self.gamma * Qcurr.data - Qprev.data)
 
         Qcurr = reward + self.gamma * Qcurr.data
         Qcurr = Variable(Qcurr)
@@ -68,7 +67,7 @@ class SARSA:
         self.optimizer.step()
         self.prevState = inputs
         self.prevAction = actions
-        return Qupdate
+        return Qcurr 
 
     def saveNetwork(self):
         torch.save(self.model.state_dict(), './modelconfig')

### 2017-12-06 Iroko/iroko_controller.py

@@ -283,10 +283,17 @@ if __name__ == '__main__':
             Agent.updateHostsBandwidth(interface, free_bandwidths[interface], drops[interface] )
             # A supposedly more eloquent way of doing it
             reward = 1
+            if(drops[interface] > 0.0):
+                reward = -1
+            else:
+                reward = 1
+
+            
             Agent.updateCritic(interface, data, reward)
             Agent.updateActor(interface, reward)
             # txrate = random.randint(1310720, 2621440)
-            ic.send_cntrl_pckt(interface, Agent.getHostsBandwidth(interface))
+            #ic.send_cntrl_pckt(interface, Agent.getHostsBandwidth(interface))
+            ic.send_cntrl_pckt(interface, Agent.getHostsPredictedBandwidth(interface))
             # ic.send_cntrl_pckt(interface, txrate)
 
         Agent.predictBandwidthOnHosts()

### 2017-12-06 Documents/Report/design.tex
Added report to documents

--- 2017-10-10 Proposal/design.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Design}
+\section{Design / Proposed Approach}
 \label{sec:design}
 
-Simple System
+\subsection{Simple System}
 In our initial simple TCPToken design, a centralized controller regulates all node traffic by provisioning end-hosts with tokens.
-These tokens act as the “currency” of the system.
-The total amount is fixed to the aggregate bandwidth available in the network and the traffic window size of sending end hosts is calculated based on the current token availability.
+These tokens act as the "currency" of the system.
+The total amount is fixed to some factor of the aggregate bandwidth available in the network, and the traffic window size of sending end hosts is calculated based on the host's current token availability.
 If a node or switch is overburdened, it may notify the scheduler, which will adjust the traffic window of responsible sending nodes remotely.
-Any management or control operations such as token distribution and updates are priority-queued to guarantee a fast and low-latency response.
-In the case of an end-host requiring more bandwidth it can also notify the controller, which may comply based on priority and availability.
+Management and control operations including token distribution and updates are priority-queued to guarantee a fast and low-latency response.
+If an end-host requires more bandwidth, it can also notify the controller, which may comply based on priority and availability.
 Initially, the controller will compute optimal route configuration based on the topology and link bandwidth using a simple heuristic bin-packing approach.
-End-hosts will be initialized with a fixed low-to-medium bandwidth guarantee, which will be adjusted over time.
+End-hosts will be initialized with a fixed low-to-medium bandwidth guarantee that will be adjusted over time.
 
-Advanced System
+\subsection{Advanced System}
 The aforementioned system is intended only as a proof of concept.
@@
 Ideally, all management in the TCPToken system should be asynchronous and preemptive. 
-1) Paying for traffic
+\subsubsection{ Paying for traffic}
 In a complete design, tokens will be used as a transport pass instead of simply acting as traffic shaping parameters.
@@
 
-2) Predicting traffic
+\subsubsection{Predicting traffic}
 ML-based scheduling is an emerging research field, also known as Knowledge-Defined Networking. Datacenter traffic follows patterns, which can be predicted using statistical methods.
@@
 
-Fault-Tolerance and Scalability
+\subsubsection{Fault-Tolerance and Scalability}
 To handle fault-tolerance we plan to integrate a fallback-mechanism to conventional TCP, if the node has not received a TCPToken in a prefixed amount of time.
 End-hosts will only be marginally affected by a controller or link failure, as they are guaranteed a fixed amount of bandwidth based on their token availability.
-To avoid overburdening the controller, we envision a distributed hive of schedulers each managing their own region in the final design.
-Add ONOS CITATION In case of failure, a shadow node or neighbouring device may take over computation temporarily, until the main node has recovered.
-CITATION
+To avoid overburdening the controller, we envision a distributed hive of schedulers (e.g., Onos) each managing their own region in the final design.~\cite{onos} In case of failure, a shadow node or neighbouring device may take over computation temporarily, until the main node has recovered.~\cite{disco}

### 2017-12-06 Documents/Report/motivation.tex
Added report to documents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivation}
\label{sec:motivation}
Sed venenatis odio a augue porttitor, vel blandit tortor luctus. Pellentesque condimentum ante lacus, imperdiet commodo odio mattis a. Donec non consectetur libero. Donec efficitur erat nec dui convallis, sed efficitur nulla pharetra. Nullam sed feugiat est. Vestibulum est nisi, venenatis ac augue molestie, malesuada dignissim dolor. Aenean nisl ante, accumsan nec sagittis quis, euismod eleifend purus. Duis laoreet neque eros, non convallis nunc viverra eget. Sed eleifend neque neque, non faucibus justo vulputate vel. Fusce aliquam ultricies ligula ut efficitur. Nunc ut neque tristique, posuere sem vitae, consectetur nisl.

Donec pretium ut enim ac congue. Suspendisse nisl est, malesuada sit amet posuere a, gravida vitae lacus. Suspendisse et risus mauris. Nam imperdiet sit amet augue sed sodales. Quisque ante orci, aliquet non interdum nec, gravida sed enim. Nunc interdum, purus et scelerisque malesuada, diam lorem efficitur orci, et consequat ligula augue sed lacus. Morbi malesuada eleifend diam eu rhoncus. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.

### 2017-12-06 Documents/Report/relatedworks.tex
Added report to documents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
\label{sec:related}

Most of the work in congestion control relies on a centralised scheduler which periodically gets updates about the flows in the network and reserves specific paths for long running flows.

MicroTE~\cite{microte} does predictive OpenFlow routing and scheduling using a central SDN controller. Hedera~\cite{hedera} schedules large so called "elephant" flows by gathering switch statistics. However it pays poor attention to short term flows and network latency. FastPass~\cite{fastpass} on the other hand, proposes a zero-queue network and promises a very low latency network. The feasibility of this proposal remains questionable when considering the high cost of implementation.

Recent proposals on decentralised scheduling like ExpressPass~\cite{expresspass} have heavily influenced our research. ExpressPass posits a credit based system similar to TCPTokens in order to control the congestion in the network. Effecting flow rerouting using ML techniques in KDN~\cite{kdn} and other works shows promising signs of further improvements in this area.

### 2017-12-09 Iroko/DDPG/Actor.py
Newer, hopfully better agent

@@ -1,0 +1,15 @@
+import torch.nn as nn
+import torch.nn.functional as F
+
+class Actor(nn.Module):
+    def __init__(self, state = 54, actions = 18,  hidden1=81, hidden2= 40):
+        super(Actor, self).__init__()
+        self.hidden1 = nn.Linear(state, hidden1)
+        self.hidden2 = nn.Linear(hidden1, hidden2)
+        self.outputs = nn.Linear(hidden2, actions)
+
+    def forward(self, x):
+        x = F.elu(self.hidden1(x))
+        x = F.elu(self.hidden2(x))
+        return F.hardtanh(self.outputs(x))
+        

### 2017-12-09 Iroko/DDPG/AsyncDDPG.py

@@ -1,0 +1,90 @@
+from Actor import Actor
+from Critic import Critic
+from TargetActorCritic import TargetActorCritic
+
+import torch
+from torch.autograd import Variable
+import torch.optim as optim
+import random
+import copy
+class AsyncDDPG(object):
+    def __init__(self, gamma, s, a, learningRate = 1e-3,criticpath=None, actorpath=None):
+        self.gamma =gamma
+        self.actor = Actor(state= s, actions = a, hidden1 = 180, hidden2 = 87)
+        self.critic = Critic(state = s, actions = a, hidden1 =250, hidden2 = 100)
+        if(not(criticpath== None)):
+            self.critic.load_state_dict(torch.load(criticpath))
+        if(not(actorpath==None)):
+            self.actor.load_state_dict(torch.load(actorpath))
+        
+        self.actorOptimizer = optim.Adam(self.actor.parameters(),learningRate)
+        self.criticOptimizer = optim.Adam(self.critic.parameters(),learningRate)
+        #more a dimensionality thing
+        self.state = s
+        self.action = a 
+        self.count = 0
+
+    def PerformUpdate(self, batchsize, target):
+        #Mildly important, according to https://github.com/vy007vikas/PyTorch-ActorCriticRL
+        # the criterion on the actor is this: sum(-Q(s,a)) I'm assuming this is over the batch....
+        self.actorOptimizer.zero_grad() 
+        self.criticOptimizer.zero_grad()
+ 
+        batch = target.getBatchMemory(batchsize) 
+
+        Q = torch.zeros(len(batch),self.state + self.action )
+        Qprime = torch.zeros(len(batch),self.state + self.action )
+        rewards = torch.zeros(len(batch), 1)
+
+        # This loop should generate all Q values for the batch
+        i = 0
+        for sample in batch:
+            Q[i,:]= torch.cat((sample['s'], sample['a']))
+            transition = target.targetActor(Variable(sample['sprime'],volatile=True)).data
+            Qprime[i,:]  = torch.cat((sample['sprime'], transition),dim=0)
+            rewards[i,0] = sample['r'][0]
+            i += 1
+
+        #Critic Update
+        Qprime = self.gamma * target.targetCritic(Variable(Qprime)).data + rewards
+        Qprime = Variable(Qprime)
+
+        Q = self.critic(Variable(Q))
+        criterion = torch.nn.MSELoss()
+        loss = criterion(Q, Qprime)
[…]

### 2017-12-09 Iroko/DDPG/Critic.py

@@ -1,0 +1,16 @@
+import torch
+import torch.nn as nn
+import torch.nn.functional as F
+
+class Critic(nn.Module):
+    def __init__(self, state = 54, actions = 18, hidden1=100, hidden2 = 30):
+        super(Critic, self).__init__()
+        self.hidden1 = nn.Linear(state + actions, hidden1)
+        self.hidden2 = nn.Linear(hidden1, hidden2)
+        self.outputs = nn.Linear(hidden2, 1)
+
+    def forward(self, x):
+        x = F.elu(self.hidden1(x))
+        x = F.elu(self.hidden2(x))
+        return self.outputs(x)
+ 

### 2017-12-09 Iroko/DDPG/DDPG.py

@@ -1,0 +1,116 @@
+from Actor import Actor
+from Critic import Critic
+from ReplayMemory import ReplayMemory
+import torch
+from torch.autograd import Variable
+import torch.optim as optim
+import random
+class DDPG:
+    def __init__(self, gamma, memory, s, a, tau, learningRate = 1e-6,criticpath=None, actorpath=None):
+        self.gamma =gamma
+        self.memory = ReplayMemory(memory)
+        self.actor = Actor(state= s, actions = a)
+        self.critic = Critic(state = s, actions = a)
+        if(not(criticpath== None)):
+            self.critic.load_state_dict(torch.load(criticpath))
+        if(not(actorpath==None)):
+            self.actor.load_state_dict(torch.load(actorpath))
+        self.targetActor = Actor(state= s, actions = a)
+        self.targetActor.load_state_dict(self.actor.state_dict())
+        self.targetCritic = Critic(state= s, actions = a)
+        self.targetCritic.load_state_dict(self.critic.state_dict())
+        self.tau = tau
+        
+        self.actorOptimizer = optim.Adam(self.actor.parameters(),learningRate)
+        self.criticOptimizer = optim.Adam(self.critic.parameters(),learningRate)
+        #more a dimensionality thing
+        self.state = s
+        self.action = a 
+
+    def processNoise(self):
+        #this should be something more eloquent....
+        ret = torch.rand(self.action) 
+        for i in range(0, self.action):
+            r = random.random()
+            if ( r <= .33):
+                ret[i] = ret[i]
+            elif ( .33 < r and r <=.66):
+                ret[i] = 0
+            else:
+                ret[i] = -ret[i]
+        return ret 
+
+    def selectAction(self, state):
+        #remember, state better be an autograd Variable
+        ret = self.targetActor(Variable(state)).data
+        if(random.random() > .5):
+            ret = ret + self.processNoise()
+        return torch.clamp(ret, -1.0, 1.0) 
+
+    def addToMemory(self, state, action, reward, stateprime):
+        self.memory.push(state, action, reward, stateprime)
+    def primedToLearn(self):
+        return self.memory.isFull()
+
+    def PerformUpdate(self, batchsize):
+        #Mildly important, according to https://github.com/vy007vikas/PyTorch-ActorCriticRL
[…]

### 2017-12-09 Iroko/DDPG/ReplayMemory.py

@@ -1,0 +1,26 @@
+import random
+
+class ReplayMemory:
+    def __init__(self, size = 40, dims = 41):
+        self.memory = []
+        self.indx = 0
+        self.size = size 
+
+    def push(self, s, a, r, sprime):
+        mem = {'s': s, 'a': a, 'r': r, 'sprime': sprime}
+        if(len(self.memory) < self.size):
+            self.memory.append(mem)
+        else:
+            self.memory[self.indx] = mem
+        self.indx = (self.indx + 1) % self.size
+    def batch(self, batchSize):
+        return random.sample(self.memory, batchSize)
+
+    def singleSample(self):
+        return random.sample(self.memory, 1)
+
+    def len(self):
+        return len(self.memory)
+
+    def isFull(self):
+        return len(self.memory) == self.size

### 2017-12-09 Iroko/DDPG/TargetActorCritic.py

@@ -1,0 +1,65 @@
+from Actor import Actor
+from Critic import Critic
+from ReplayMemory import ReplayMemory
+import torch
+from torch.autograd import Variable
+import torch.optim as optim
+import copy
+import random
+class TargetActorCritic:
+    def __init__(self, actor, critic, memory, s, a, tau, epsilon=0.5):
+        self.memory = ReplayMemory(memory)
+        self.targetActor = copy.deepcopy(actor) 
+        self.targetCritic = copy.deepcopy(critic) 
+        self.tau = tau
+        self.epsilon = epsilon
+        #more a dimensionality thing
+        self.state = s
+        self.action = a 
+
+    def processNoise(self):
+        #this should be something more eloquent....
+        ret = torch.rand(self.action) / 2.0 
+        for i in range(0, self.action):
+            r = random.random()
+            if ( r <= .33):
+                ret[i] = ret[i]
+            elif ( .33 < r and r <=.66):
+                ret[i] = 0
+            else:
+                ret[i] = -ret[i]
+        return ret 
+
+    def selectAction(self, state):
+        #remember, state better be an autograd Variable
+        ret = self.targetActor(Variable(state)).data
+        if(random.random() > self.epsilon):
+            ret = ret + self.processNoise()
+
+        return torch.clamp(ret, 0.0, 1.0) 
+
+    def addToMemory(self, state, action, reward, stateprime):
+        self.memory.push(state, action, reward, stateprime)
+
+    def getBatchMemory(self, batchsize):
+        return self.memory.batch(batchsize)
+
+    def primedToLearn(self):
+        return self.memory.isFull()
+
+    def UpdateTargetNetworks(self, critic, actor): 
+        criticDict = critic.state_dict()
+        tCriticDict = self.targetCritic.state_dict()
+        for param in criticDict.keys():
+            tCriticDict[param] = tCriticDict[param] * (1 - self.tau) + criticDict[param] * self.tau
+
+        actorDict = actor.state_dict()
+        tActorDict = self.targetActor.state_dict()
+        for param in actorDict.keys():
+            tActorDict[param] = tActorDict[param] * (1 - self.tau) + actorDict[param] * self. tau
+
+        self.targetCritic.load_state_dict(tCriticDict)
+        self.targetActor.load_state_dict(tActorDict)
+
+
+

### 2017-12-09 Iroko/DDPG/UnitTestReplayMemory.py

@@ -1,0 +1,42 @@
+import unittest
+import torch
+from ReplayMemory import ReplayMemory
+
+class UnitTestReplayMemory(unittest.TestCase):
+
+
+    def testDefaultInit(self):
+        RepMem = ReplayMemory()
+        self.assertEqual(RepMem.size, 40)
+
+    def testPush(self):
+        RepMem = ReplayMemory()
+        actions = ['a','b', 'c']
+        for i in range(0, RepMem.size):
+            RepMem.push(i, actions[i % 3],  1, i + 1)
+            memory = RepMem.memory[i]
+            self.assertEqual(memory['a'], actions[i%3])
+            self.assertEqual(memory['s'], i)
+            self.assertEqual(memory['sprime'], i + 1)
+            self.assertEqual(memory['r'], 1)
+
+        for i in range(RepMem.size, RepMem.size * 2):
+            RepMem.push(i, actions[i % 3],  1, i + 1)
+            memory = RepMem.memory[i % RepMem.size]
+            self.assertEqual(memory['a'], actions[i%3])
+            self.assertEqual(memory['s'], i)
+            self.assertEqual(memory['sprime'], i + 1)
+            self.assertEqual(memory['r'], 1)
+
+        
+    def testPushInxes(self):
+        RepMem = ReplayMemory()
+
+        for i in range(0, 200000):
+            RepMem.push(1,1,1,1)
+            self.assertEqual(RepMem.indx, ((i + 1) % RepMem.size))
+        self.assertTrue(RepMem.isFull())
+
+
+if __name__ == '__main__':
+    unittest.main()

### 2017-12-09 Iroko/DDPG/testscript.py

@@ -1,0 +1,21 @@
+from Actor import Actor
+from Critic import Critic
+import torch
+from torch.autograd import Variable
+import torch.nn as nn
+
+
+actor = Actor()
+state = Variable(torch.rand(41))
+print(actor)
+
+print("output of network")
+action = actor(state)
+print(action)
+
+critic = Critic()
+
+stateaction = Variable(torch.cat((state.data,action.data)))
+
+Qval = critic(stateaction)
+print(Qval)

### 2017-12-09 Iroko/LearningAgentv2.py

@@ -1,0 +1,123 @@
+import torch
+import torch.nn
+from torch.autograd import Variable
+import random
+import numpy as np
+import math
+from DDPG.DDPG import DDPG
+
+
+class LearningAgentv2:
+    def __init__(self,gamma = .9, lam = .1,  memory= 100 ,  initMax = 0, defaultmax = 10e6):
+        self.hosts = {}
+        self.hostcount = 0
+        self.gamma = gamma
+        self.lam = lam
+        self.defaultmax = defaultmax
+        self.initmax = initMax
+        self.controller = DDPG(gamma,memory,5 + 16, 1,tau=.25) 
+        
+
+
+    def initializePorts(self, ports):
+        # cleans out old ports and creates maps for new ones
+        self.hosts = {}
+        for key in ports.keys():
+            self.hosts[key] = self.createHost()
+
+    def createHost(self):
+        self.hostcount += 1
+        host = dict(alloctBandwidth = self.initmax, e =  0, predictedAllocation = self.initmax)
+        #host['controller'] = self.generateController() 
+        host['action'] = 0.0 
+        #lol, what is good data abstraction really though? 
+        host['state'] = torch.zeros(self.controller.state)
+        #host['modifier'] = 0
+        host['id'] = self.hostcount - 1
+        return host
+
+    def update(self, interface, data, reward):
+        # assumes data is a pytorch.tensor
+        hostvector = torch.Tensor(self.hostcount)
+        hostvector[self.hosts[interface]['id']] = 1.0
+        stateprime = torch.cat((data, hostvector), dim=0)
+ 
+        state = self.hosts[interface]['state'] 
+        action = self.hosts[interface]['action']
+        action = torch.Tensor([action])
+        reward = torch.Tensor([reward])
+        self.controller.addToMemory(state, action, reward, stateprime)
+        if (self.controller.primedToLearn()): 
+            print('Learning is happening')
+            self.controller.PerformUpdate(32)
+            self.controller.UpdateTargetNetworks()
+            self.controller.saveActorCritic()
+        self.hosts[interface]['state'] = stateprime
+
+    def predictBandwidthOnHosts(self):
+        for host in self.hosts.keys():
+            self.predictBandwidthOnHost(host)
+
+    def predictBandwidthOnHost(self, interface):
+        state = self.hosts[interface]['state']
+        adjustment = self.controller.selectAction(state)[0]
[…]

### 2017-12-09 Iroko/iroko_controller.py

@@ -14,6 +14,7 @@ import numpy as np
 import torch
 import random
 from LearningAgent import LearningAgent
+from LearningAgentv2 import LearningAgentv2
 
 MAX_CAPACITY = 5e6   # Max capacity of link
 TOSHOW = True
@@ -266,7 +267,8 @@ if __name__ == '__main__':
     # ic.run()
 
     stats = StatsCollector()
-    Agent = LearningAgent(initMax=MAX_CAPACITY)
+    #Agent = LearningAgent(initMax=MAX_CAPACITY)
+    Agent = LearningAgentv2(initMax=MAX_CAPACITY)
     Agent.initializePorts(i_h_map)
     stats._set_interfaces()
     prevdrops = [] 
@@ -282,15 +284,15 @@ if __name__ == '__main__':
             # My Naive way to update bandwidth
             Agent.updateHostsBandwidth(interface, free_bandwidths[interface], drops[interface] )
             # A supposedly more eloquent way of doing it
-            reward = 1
-            if(drops[interface] > 0.0):
+            reward = 0
+            if(drops[interface]  > 0.0):
                 reward = -1
             else:
                 reward = 1
 
-            
-            Agent.updateCritic(interface, data, reward)
-            Agent.updateActor(interface, reward)
+            Agent.update(interface,data,reward) 
+            #Agent.updateCritic(interface, data, reward)
+            #Agent.updateActor(interface, reward)
             # txrate = random.randint(1310720, 2621440)
             #ic.send_cntrl_pckt(interface, Agent.getHostsBandwidth(interface))
             ic.send_cntrl_pckt(interface, Agent.getHostsPredictedBandwidth(interface))
@@ -300,8 +302,8 @@ if __name__ == '__main__':
         # update the allocated bandwidth
         # wait for update to happen
 
-        Agent.displayAllHosts()
-        Agent.displayALLHostsBandwidths()
+        #Agent.displayAllHosts()
+        #Agent.displayALLHostsBandwidths()
         Agent.displayALLHostsPredictedBandwidths()
         Agent.displayAdjustments()
 

### 2017-12-09 Iroko/LearningAgentv2.py
updates to clamp max bandwidth and saved parameters

@@ -48,7 +48,6 @@ class LearningAgentv2:
         reward = torch.Tensor([reward])
         self.controller.addToMemory(state, action, reward, stateprime)
         if (self.controller.primedToLearn()): 
-            print('Learning is happening')
             self.controller.PerformUpdate(32)
             self.controller.UpdateTargetNetworks()
             self.controller.saveActorCritic()
@@ -63,7 +62,8 @@ class LearningAgentv2:
         adjustment = self.controller.selectAction(state)[0]
         adjustment = max(min(adjustment, .90), -0.90)
         allocation = self.hosts[interface]['predictedAllocation']
-        allocation += max(min(adjustment * allocation, self.defaultmax), 0)
+        allocation += adjustment * allocation
+        allocation = max(min(allocation, self.defaultmax), 0)
         self.hosts[interface]['predictedAllocation'] = allocation 
         self.hosts[interface]['action'] = adjustment 
 

### 2017-12-09 Documents/Report/Iroko_Report.tex
Update report structure.

--- 2017-10-10 Proposal/paper.tex
@@
 Datacenter networks have become a hotbed for research in the recent past.
-Many works have been published on improving the  bisection bandwidth, reducing cost, and latency.
-As a core contribution of this research, centralized scheduling has entered the stage as dominant strategy to manage flows.
+Many works have been published on improving bisection bandwidth, cost, and latency.
+As a core contribution of this research, centralized scheduling has entered the stage as a dominant strategy to manage flows.
 Benefitting from a global view and full control over the network, centralized schedulers are able to enforce fine-grained traffic control, frequently achieving near-optimal bandwidth optimization.
@@
 Many algorithms are designed to only respond to packet loss or significant increase in latency, not to actively prevent it.
-
 In this project, we plan to explore the opportunities of employing a preemptive and tightly controlling central network scheduler.
-Using tokens as a global tool to enforce traffic limiting and policies, the scheduler is able to proactively control the flow of traffic as well as respond to application requests.
-"Knowledge" and predictive analysis in networks is a growing trend in research which we intend to leverage in this system.
+Using tokens as a global tool to enforce traffic limits and admission control, the scheduler is able to proactively steer the flow of traffic.
+The notion of "knowledge" and predictive analysis in networks is a growing trend in research, which we intend to leverage in this system.
 
@@
 Our tool of choice to model our system is Mininet, a rapid prototyping emulator for datacenter networks.
-We will compare our "TCPToken" system against established TCP-congestion systems such as Hedera and DCTCP.
-Based on the findings and measurement results, we will reassess the feasibility and prospects of a token-based, predictive congestion control algorithm.
+We will compare our "Iroko" system against established TCP-congestion systems such as Hedera and DCTCP.
+Based on the findings and measurement results, we will reassess the feasibility and prospects of a predictive congestion control algorithm.
 \end{abstract}
@@
 
-%\input{intro}
-%
-%\input{relatedworks}
-%\input{design}
+\input{intro}
+%\input{relatedwork}
+\input{design}
 % %\input{motivation}
-%\input{implementation}
-%\input{eval}
-%\input{timeline}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-\label{sec:intro}
-
-Initially, research in network congestion was dominated by the assumption of a decentralised, autonomous network - end-hosts only have control over the amount of traffic they send, and are unaware of the intentions or traffic rates of their peers.
-Similarly, switches and routers are unaware of global traffic patterns and only forward based on their local notion of optimality.
-
-In line with these assumptions, TCP was designed to optimize traffic globally on a simple fairness principle: nodes react to packet loss and assume that others will behave accordingly.
-An equilibrium is reached when all end-hosts achieve a traffic rate that, in sum, conforms to the maximum available bandwidth of the most congested link.
-TCP works well  in scenarios where many different distrustful participants compete for limited bandwidth.
-Still, TCP is a \textit{reactive protocol}; the fact that packet loss and latency increases occur in the network already indicates a problem.
-Packet loss is primarily due to overflowing queues in forwarding elements, implying that traffic has not been optimally distributed.
-Ideally, a network should always be "zero-queue", i.e., latency will merely be induced by propagation, and not queuing delay.
-
-Queueing has generally not been a dominating issue in wide-area and enterprise networks, as traffic is sufficiently distributed and diverse, with only few "hot" target hosts.~\cite{hedera, microte} Traffic optimization is a substantial challenge; network operators have no control over the individual network elements nor its participants.
-Under these conditions, TCP and its extensions can be considered a best-effort solution.
-
-Developments in the past decade have changed the general networking environment.
-Datacenters have emerged as an exciting new research frontier, posing novel design challenges and opportunities.Driven by minimization of costs and maximization of compute power, datacenters must  run at maximum utilization to achieve an optimal compute/cost ratio.
-Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load network, requiring more sophisticated approaches to solve congestion control.
-On the other hand, operators now have the ability to freely control and adapt their network architecture, leading to highly customized systems and fine-grained optimization.
-As a result, Software-Defined Networking (SDN) has emerged as a new networking paradigm.
-
-Moving away from the principle of distributed communication and routing, SDN introduces the notion of "centralized management".
-A single controller with global knowledge is able to automatically modify and adapt the forwarding tables of all switches in the network, while notifying end hosts of changes in the network.
-These two new trends in system design facilitated impactful new innovation opportunities in the space of TCP congestion research.
-Traffic can now be managed in a  \textit{centralised} fashion based on \textit{global knowledge} of the entire topology and traffic patterns.
-
-A new line of centralised schedulers has emerged that  can achieve close to optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
-However, these schedulers are still  \textit{reactive}  in nature.
-The central controller responds to changes in the network or requests by applications, which may cost valuable roundtrip latency.
-Often, short-term flows or bursts are unaccounted for, which causes undesirable packet loss and backpropagating congestion.
-
-A much more desirable solution is a global, centralised arbiter which is able to predict and fairly distribute flows in the network before bursts or congestion occur.
-By treating the network's compute and forwarding power as a single finite resource, a controller acts like the OS scheduler distributing CPU time slices to processes.
-This design approach follows SDN's aspiration of introducing operating systems abstractions to the networking domain space.
-
-In this project, we plan to explore the possibilities of a centralised, proactive flow scheduler. We ask ourselves the following research questions:
-\begin{enumerate}
-	\item Is it possible to design a centralised token-based scheduling network?
-	\item Is it possible to predict traffic and preemptively schedule flows and token distribution in a datacenter context?
-	\item Using this approach, are we able to achieve better performance and utilization than existing solutions?
-\end{enumerate}
-
-In the scope of this course, we attempt to answer question one and design a simple token-based scheduler in Mininet.
-If successful, we will benchmark our results and evaluate the level of utilization compared to contemporary scheduling systems.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}
-\label{sec:related}
-
-Most of the work in congestion control relies on a centralised scheduler which periodically gets updates about the flows in the network and reserves specific paths for long running flows.
-
-MicroTE~\cite{microte} does predictive OpenFlow routing and scheduling using a central SDN controller.
-Hedera~\cite{hedera} schedules large so called "elephant" flows by gathering switch statistics.
-However it pays poor attention to short term flows and network latency.
-FastPass~\cite{fastpass} on the other hand, proposes a zero-queue network and promises a very low latency network.
-The feasibility of this proposal remains questionable when considering the high cost of implementation.
-
-Recent proposals on decentralised scheduling like ExpressPass~\cite{expresspass} have heavily influenced our research.
-ExpressPass posits a credit based system similar to TCPTokens in order to control the congestion in the network.
-Effecting flow rerouting using ML techniques in KDN~\cite{kdn} and other works shows promising signs of further improvements in this area.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Design / Proposed Approach}
-\label{sec:design}
-
-\subsection{Simple System}
-In our initial simple TCPToken design, a centralized controller regulates all node traffic by provisioning end-hosts with tokens.
-These tokens act as the "currency" of the system.
-The total amount is fixed to some function of the bisection and aggregate bandwidths available in the network, and the traffic window size of sending end hosts is calculated based on the host's current token availability.
-If a node or switch is overburdened, it may notify the scheduler, which will adjust the traffic window of responsible sending nodes remotely.
-Management and control operations including token distribution and updates are priority-queued to guarantee a fast and low-latency response.
-If an end-host requires more bandwidth, it can also notify the controller, which may comply based on priority and availability.
-Initially, the controller will compute optimal route configuration based on the topology and link bandwidth using a simple heuristic bin-packing approach.
-End-hosts will be initialized with a fixed low-to-medium bandwidth guarantee that will be adjusted over time.
-
-\subsection{Advanced System}
-The aforementioned system is intended only as a proof of concept.
-It is still a  synchronous and reactive approach, dependent on notifications and requests by overburdened switches and underprovisioned hosts.
-Ideally, all management in the TCPToken system should be asynchronous and preemptive. 
-\subsubsection{ Paying for traffic}
-In a complete design, tokens will be used as a transport pass instead of simply acting as traffic shaping parameters.
-On a per-flow basis, nodes will be able to attach tokens to traffic as a form of payment.
-Enforcement is performed on the switch level, any flow that is not "paid" by a token will be dropped.
-Nodes may be handed different types of tokens to serve load-balancing and traffic shaping purposes.
-Applications operating on traditionally short-lived flows and bursts may sign their traffic with a different token type than long-lived permanent processes.
-Switches will forward traffic and balance flows accordingly with only minor interference from the central controller.
-
-\subsubsection{Predicting traffic}
-ML-based scheduling is an emerging research field, also known as Knowledge-Defined Networking. Datacenter traffic follows patterns, which can be predicted using statistical methods.
-Combined with using tokens as a traffic engineering enforcement, a preemptive scheduling strategy in datacenters may be feasible.
-
-We do not intend to implement the advanced system for this project, but are using it as a guideline to motivate our research.
-Based on the experiences and knowledge gained from implementing the basic system, we will reevaluate the feasibility and practicality of the second concept.
-
-\subsubsection{Fault-Tolerance and Scalability}
-To handle fault-tolerance we plan to integrate a fallback-mechanism to conventional TCP, if the node has not received a TCPToken in a prefixed amount of time.
-End-hosts will only be marginally affected by a controller or link failure, as they are guaranteed a fixed amount of bandwidth based on their token availability.
-
-To avoid overburdening the controller, we envision a distributed hive of schedulers (e.g., Onos) each managing their own region in the final design.~\cite{onos} In case of failure, a shadow node or neighbouring device may take over computation temporarily, until the main node has recovered.~\cite{disco}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Implementation}
-\label{sec:implementation}
-We intend to emulate our system in Mininet~\cite{mininet} to observe traffic patterns and infer a suitable token algorithm.
-Mininet has proven itself to be a viable tool to model new congestion control algorithms~\cite{mininet_learning}, and will help us prototype our concept efficiently.
-We will build a custom SDN controller that interacts with traditional OpenFlow software switches as well as end-hosts.
-End-hosts will run a custom real-world traffic generation script which adjusts based on information packets sent by the controller.
-If time permits, we may expand our implementation to MaxiNet, which can emulate large-scale network stress tests of thousands of nodes on multiple physical hosts.
-
-We initially considered a second implementation alternative in C/C++ based on the FastPass~\cite{fastpass} source code.
-This would provide us with a fully deployable system which we could fork our implementation from.
-A major advantage of this approach is the ability to test scenarios and traffic algorithms using real software code.
-However, several concerns made us favour a Mininet emulation instead.
-Firstly, FastPass relies on DPDK integration, which requires actual hardware interfaces.
-The central arbiter in the FastPass design would need to run on a dedicated machine, which increases prototyping and development complexity substantially.
-
-Secondly, the FastPass code is highly specialized and optimized research code with only little available documentation.
-Modifying and evolving the source code will require thorough understanding of kernel and networking development, a significant time-sink.
-For a class project, these may be major initial hurdles, taking away from the research aspect of the design concept.
-Consequently, we have decided to pursue an approach which allows us to quickly develop an understanding of the problem without being obstructed by engineering work.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Evaluation}
-\label{sec:evaluation}
-To evaluate the effectiveness of our system, we plan to measure against existing centralized as well as decentralized solutions.
-We will run datacenter traffic on a fat-tree topology of various depth and observe the effect each congestion algorithm will have on the overall bandwidth utilization and latency.
-The centralized design will be based on Hedera~\cite{hedera}, a common and influential datacenter scheduler.
-
-The decentralized congestion control mechanism will be DCTCP~\cite{dctcp}, a state-of-the-art TCP congestion algorithm.
-As baseline we will also compare to the default TCP congestion algorithm to estimate how our scheduler fares against the default case.
-As our initial simple design is very constrained in its ability to route and control traffic we expect to achieve less optimality against DCTCP and Hedera, but gain a substantial advantage over common TCP.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Timeline}
-\label{sec:timeline}
-\begin{itemize}
-	\item \underline{October 15th:} Finalize draft and related work / background section.
-	\item \underline{ October 22nd:} Iterate over literature and thoroughly analyze related work. Set up an initial prototyping environment in Mininet on which each team member can work on.
-	\item \underline{October 30th:} Finalize the initial simple design and explore naive congestion modelling in Mininet. Integrate existing Mininet emulations of Hedera and DCTCP.
-	\item \underline{November 12th:} Develop initial naive TCPTokens prototype involving controller and end-host design.
-	\item \underline{November 19th:} Develop intelligent centralised congestion control heuristic.
-	\item \underline{November 26th:} Run evaluations and microbenchmarks.
-	\item \underline{December 3rd:} Iterate and improve the scheduler, explore topology and bandwidth constraints. Test at scale.
-	\item \underline{December 10th:} Write report and prepare presentation.
-\end{itemize}
+\input{implementation}
+\input{eval}
 
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{paper}
+\bibliography{Iroko_Proposal}
 

### 2017-12-09 Documents/Report/design.tex
Update report structure.

--- 2017-10-09 Proposal/design.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Design}
+\section{Design / Proposed Approach}
 \label{sec:design}
+
+\subsection{Overview and Goals}
+In our initial simple system, "Iroko", a centralized controller regulates all 
+node traffic by rate-limiting end-hosts. We have opted for a centralized 
+manager in favor of a distributed protocol to leverage the advantages of a 
+global network view.~\footnote{For now, we do not expect our design to scale up 
+to thousands of switches. Iroko is intended for small to mid-tier size data 
+centers, which may benefit from a simplistic, centralized scheduling model.}
+By polling each switch individually for port and utilization statistics we are 
+able to infer a global traffic matrix, which we can amalgamate with static 
+routing and topology information.
+
+Iroko guarantees minimal congestion and low average access latency. Reducing 
+network-global packet loss and jitter is the priority and objective function of 
+the arbiter, which will enforce these goals by restricting host bandwidth. In 
+an ideal Iroko system, packet-loss will only rarely, if ever, occur. In respect 
+to the free lunch theorem we aim to trade off maximum bandwidth utilization for 
+optimal system stability and reliable latency.
+
+It is important to note that routing decisions are not in the current scope of 
+Iroko. While it may certainly beneficial to include dynamic and adaptive 
+routing decisions in a predictive scheduler, we do not include these features 
+in the current system due to time and complexity limitations. Iroko operates on 
+static flow routes defined and generated by ECMP. It is assumed that these 
+routes will not vary substantially and that the computed ECMP forwarding hash 
+is consistent.
+
+\subsection{Controlling traffic flow}
+Rate-limiting is performed on the granularity of the IP protocol, allowing us 
+to restrict traffic to specific hot regions and to package flows into groups.
+End-hosts are guaranteed a limited amount of bandwidth which they may send to 
+each destination address. The traffic window size for any flow is calculated 
+based on the host's current traffic availability. Total bandwidth of a group of 
+flows may never exceed the imposed bandwidth limitation, but nodes are free to 
+distribute the allocated resources on singular subflows.
+
+The central arbiter enforces these bandwidth and access limitations by 
+assigning "tokens" to nodes. These tokens are plain information packets 
+specifying the affected destination address flow, its bandwidth guarantee, and 
+the expiration time. Tokens act as the rate-limiter of the system and form a 
+queue each end-host will cycle through.
+
+When a node opens a flow, it will look up the current bandwidth restriction for 
+the destination IP in its token database and calculate the maximum congestion 
+window possible for the particular TCP connection. In our simple design, flows 
+are always assigned a fair fraction.
+
+Once a token has expired for a particular flow-group, the restrictions of the 
+next token in the queue will be applied. This mechanism is the underlying basis 
+for a predictive access control algorithm and does not require any application 
+level modification. On end-hosts, only the transport layer services will have 
+to be modified.
+
+\subsection{Adapting traffic flow}
+
+Initially, the controller will compute optimal route configuration based on the 
+topology and link bandwidth using a simple heuristic bin-packing approach. 
+End-hosts will be initialized with a fixed low-to-medium bandwidth guarantee 
+that will be adjusted over time. This bandwidth guarantee will be below the 
+host's proportional bi-section bandwidth that would be used to reach any other 
+host. That is, if every end host transfers using the full bandwidth initially 
+allocated to it,  there would be no dropped packets due to congestion.
+
+Of course, this would vastly underutilize network resources, thus the 
+controller will dynamically adjust these allocations to ensure that hosts that 
+require the bandwidth are allocated it, while host that are not currently 
+utilizing all their allocated bandwidth have their allocated bandwidth reduced. 
+To avoid starving hosts, some small amount of bandwidth must always be 
+allocated to a given host even if it not utilizing network resources. Thus, the 
+network, by design, will never reach 100\% utilization.
+
+Conversely however one should rarely see dropped packets due to network 
+congestion. Furthermore, the controller must be able to react quickly to 
+changing bandwidth needs for hosts. Ideally, using a predictive method.
+
+\begin{figure}
+\centering
+\includegraphics[width=1\linewidth]{Topology3}
+\caption{Simple sketch of the Iroko architecture. A central scheduler maintains 
+a consistent view of the current network activity and computes the optimal 
+end-host bandwidth distribution. Bandwidth is enforced at end-hosts on a per-IP 
+basis.}
+\label{fig:Topology3}
+\end{figure}
+
+\subsection{Statistical System}
+One potential implementation for our arbiter is to use reinforcement learning. 
+In reinforcement learning, an agent (in our case the arbiter) performs actions 
+in the environment and receives a reward signal from the environment to adjust 
+its decisions in the next epoch ~\cite{Sutton:1998:IRL:551283}. The reward is 
+generally used to represent the goal the agent is trying to achieve. The agent 
+can freely choose whatever actions are necessary to achieve this goal 
+~\cite{Sutton:1998:IRL:551283}.
+
+The environment in this case is defined as the statistics collected from each 
+switch. This decision is based on the fact that this data is easy to collect 
+and provides a current snapshot of the performance of the network. We wish to 
+optimize over the packet loss, and define the reward to be good if packet loss 
+is decreasing (+1 reward); bad if it is increasing ( -1 reward); and acceptable 
+if packet loss remains the same within a threshold (0 reward). Using this 
+representation, we take full advantage of all the information provided in our 
+data center; This representation is replicable and can be applied beyond the 
+scope of this project.
+
+Given the data center assumption, the number of end-hosts is known in the 
+topology. We can define our arbiter's actions as a discretized representation 
+among an n-dimensional vector where each cell is one of three values: increase 
+decrease, or maintain the allocated bandwidth. This representation offers 
+sufficient complexity. There are $3^n$ possible actions to perform and to 
+achieve optimal results the agent must explore the environment extensively, 
+precluding a significantly larger action space.
+
+A potential improvement to our representation is a hierarchical learning agent. 
+In this case, our arbiter would manage agents for each host and these 
+sub-agents would manage their respective host whether to increase, decrease, or 
+maintain bandwidth for their particular host. This problem can be referred to 
+as hierarchical reinforcement learning~\cite{Borga1993}, but may prove to be 
+beyond the scope of our current proposed solution, although has the added 
+benefit of solving a set of sub-problems to address the global problem of 
+packet loss and simplifies the representation spaces in the sub-problems.
+
+Thus we frame our problem as one which can now be solved using techniques in 
+reinforcement learning. A classic algorithm to consider is the use of 
+SARSA~\cite{Sutton:1998:IRL:551283}, which is an online temporal difference 
+algorithm used in classical reinforcement learning problems. One advantage of 
+SARSA is its online nature~\cite{Sutton:1998:IRL:551283} which promotes more 
+conservative decisions. Given full representation of our environment and 
+actions, we can store information in an artificial neural network which is 
+another popular choice in reinforcement learning literature. This model would 
+store the reward signal information that is used to make future action choices, 
+and is a theoretically proven function approximator which has made intractable 
+environmental representations manageable~\cite{Sutton:1998:IRL:551283}.
+
+%Using this model, we hope to achieve near 0% packet loss and provide a 
+%%solution that gets somebody fired at google for not having tried it first.

### 2017-12-09 Documents/Report/eval.tex
Update report structure.

--- 2017-10-10 Proposal/paper.tex
@@
 \label{sec:evaluation}
-To evaluate the effectiveness of our system, we plan to measure against existing centralized as well as decentralized solutions.
-We will run datacenter traffic on a fat-tree topology of various depth and observe the effect each congestion algorithm will have on the overall bandwidth utilization and latency.
-The centralized design will be based on Hedera~\cite{hedera}, a common and influential datacenter scheduler.
 
-The decentralized congestion control mechanism will be DCTCP~\cite{dctcp}, a state-of-the-art TCP congestion algorithm.
-As baseline we will also compare to the default TCP congestion algorithm to estimate how our scheduler fares against the default case.
-As our initial simple design is very constrained in its ability to route and control traffic we expect to achieve less optimality against DCTCP and Hedera, but gain a substantial advantage over common TCP.
+In the absence of a commercial data center, the implementation of our network 
+design is going to be done on top of mininet with simulated FatTree topologies 
+of varying size. We stress test with iPerf and simulate data center traffic 
+using tcpreplay and packet traces provided by~\cite{traffic} .
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Timeline}
-\label{sec:timeline}
-\begin{itemize}
-	\item \underline{October 15th:} Finalize draft and related work / background section.
-	\item \underline{ October 22nd:} Iterate over literature and thoroughly analyze related work. Set up an initial prototyping environment in Mininet on which each team member can work on.
-	\item \underline{October 30th:} Finalize the initial simple design and explore naive congestion modelling in Mininet. Integrate existing Mininet emulations of Hedera and DCTCP.
-	\item \underline{November 12th:} Develop initial naive TCPTokens prototype involving controller and end-host design.
-	\item \underline{November 19th:} Develop intelligent centralised congestion control heuristic.
-	\item \underline{November 26th:} Run evaluations and microbenchmarks.
-	\item \underline{December 3rd:} Iterate and improve the scheduler, explore topology and bandwidth constraints. Test at scale.
-	\item \underline{December 10th:} Write report and prepare presentation.
-\end{itemize}
+To evaluate the general effectiveness of our system, we plan to measure against 
+existing centralized as well as decentralized solutions.
+The centralized design will be based on Hedera~\cite{hedera}, a common and 
+influential datacenter scheduler. The decentralized congestion control 
+mechanism will be DCTCP~\cite{dctcp}, a state-of-the-art TCP congestion 
+algorithm.
 
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{paper}
+Since we are primarily concerned about reducing the latency and packet drops 
+while keeping utilisation at maximum, the measurements to get a good insight 
+into how well the design can perform are as follows:
+\begin{enumerate}
 
-\end{document}
+\item Latency: We are aiming for a low latency network which means that 99th 
+percentile latency in the network across all flows should as low as possible. 
+Latency should be measured when all hosts are sending packets at the maximum 
+limit and also with random traffic patterns. There should be low latency even 
+during sudden bursts as the transmission rate is limited by a base value.
+
+\item Packet drop rate: Ideally this metric will approach zero, as the 
+objective of Iroko is to minimize packet loss.
+
+\item Fairness: Fairness can be tested by introducing a new host to a 
+completely saturated network and increasing the transmission rate to see if all 
+the hosts gets fair share of the total bandwidth. We are assuming all flows 
+should be at equal priority. Differential priority is out of scope.
+
+\item Responsiveness: This metric depends on the predictive power and 
+efficiency of Iroko. It may be interesting to analyze the response time of the 
+central scheduler compared to decentralized DCTCP and TCP. In addition, it is 
+valuable to identify at what data center and flow size the central computing 
+node of the network may become a bottleneck.
+
+\item Network utilization: We will measure the used bandwidth and the total 
+load on the hosts. These metrics can be measured by varying the load from each 
+host. Load to utilization ratio should be 1 until the saturation point and 
+after that utilization should stay stable at the maximum bisection bandwidth.
+It is also important to check that there is no starvation happening in any 
+host. This can be measured by simulating random traffic patterns and plotting 
+residual bandwidth and excess load. We expect overall utilization to be lower 
+compared to Hedera and DCTCP. Although Iroko will maximize goodput, we do not 
+measure it on grounds of simplicity.
+
+\end{enumerate}
+
+As baseline we will also compare to the default TCP congestion algorithm to 
+estimate how our scheduler fares against the default case. As our initial 
+simple design is very constrained in its ability to route and control traffic 
+we expect to achieve less overall optimality against DCTCP and Hedera, but 
+still gain a substantial advantage over common TCP.

### 2017-12-09 Documents/Report/implementation.tex
Update report structure.

--- 2017-10-10 Proposal/paper.tex
@@
-\documentclass[sigconf]{acmart}
-
-\usepackage{booktabs} % For formal tables
-\usepackage[normalem]{ulem}
-% Copyright
-\setcopyright{none}
-
-\input{annotations}
-
-\begin{document}
-
-\title{TCPTokens: Introducing Currency into Datacenter Congestion Control}
-
-\author{Anand Jayarajan}
-\affiliation{%
-  \institution{University of British Columbia}
-}
-\email{anandj@cs.ubc.ca}
-
-\author{Michael Przystupa}
-\affiliation{%
-  \institution{University of British Columbia}
-}
-\email{michael.przystupa@gmail.com}
-
-\author{Robert Reiss}
-\affiliation{%
-  \institution{University of British Columbia}
-}
-\email{rreiss@cs.ubc.ca}
-
-\author{Fabian Ruffy}
-\affiliation{%
-  \institution{University of British Columbia}
-}
-\email{fruffy@cs.ubc.ca}
-
-\begin{abstract}
-Datacenter networks have become a hotbed for research in the recent past.
-Many works have been published on improving the  bisection bandwidth, reducing cost, and latency.
-As a core contribution of this research, centralized scheduling has entered the stage as dominant strategy to manage flows.
-Benefitting from a global view and full control over the network, centralized schedulers are able to enforce fine-grained traffic control, frequently achieving near-optimal bandwidth optimization.
-However, modern congestion control algorithms and schedulers are still fundamentally reactive.
-Many algorithms are designed to only respond to packet loss or significant increase in latency, not to actively prevent it.
-
-In this project, we plan to explore the opportunities of employing a preemptive and tightly controlling central network scheduler.
-Using tokens as a global tool to enforce traffic limiting and policies, the scheduler is able to proactively control the flow of traffic as well as respond to application requests.
-"Knowledge" and predictive analysis in networks is a growing trend in research which we intend to leverage in this system.
-
-We investigate the potential in such a new form of interactive congestion control and analyze it against state-of-the-art solutions.
-Our tool of choice to model our system is Mininet, a rapid prototyping emulator for datacenter networks.
-We will compare our "TCPToken" system against established TCP-congestion systems such as Hedera and DCTCP.
-Based on the findings and measurement results, we will reassess the feasibility and prospects of a token-based, predictive congestion control algorithm.
-\end{abstract}
-
-\keywords{TCP, Congestion Control, SDN, Datacenter}
-
-\maketitle
-
-%\input{intro}
-%
-%\input{relatedworks}
-%\input{design}
-% %\input{motivation}
-%\input{implementation}
-%\input{eval}
-%\input{timeline}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-\label{sec:intro}
-
-Initially, research in network congestion was dominated by the assumption of a decentralised, autonomous network - end-hosts only have control over the amount of traffic they send, and are unaware of the intentions or traffic rates of their peers.
-Similarly, switches and routers are unaware of global traffic patterns and only forward based on their local notion of optimality.
-
-In line with these assumptions, TCP was designed to optimize traffic globally on a simple fairness principle: nodes react to packet loss and assume that others will behave accordingly.
-An equilibrium is reached when all end-hosts achieve a traffic rate that, in sum, conforms to the maximum available bandwidth of the most congested link.
-TCP works well  in scenarios where many different distrustful participants compete for limited bandwidth.
-Still, TCP is a \textit{reactive protocol}; the fact that packet loss and latency increases occur in the network already indicates a problem.
-Packet loss is primarily due to overflowing queues in forwarding elements, implying that traffic has not been optimally distributed.
-Ideally, a network should always be "zero-queue", i.e., latency will merely be induced by propagation, and not queuing delay.
-
-Queueing has generally not been a dominating issue in wide-area and enterprise networks, as traffic is sufficiently distributed and diverse, with only few "hot" target hosts.~\cite{hedera, microte} Traffic optimization is a substantial challenge; network operators have no control over the individual network elements nor its participants.
-Under these conditions, TCP and its extensions can be considered a best-effort solution.
-
-Developments in the past decade have changed the general networking environment.
-Datacenters have emerged as an exciting new research frontier, posing novel design challenges and opportunities.Driven by minimization of costs and maximization of compute power, datacenters must  run at maximum utilization to achieve an optimal compute/cost ratio.
-Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load network, requiring more sophisticated approaches to solve congestion control.
-On the other hand, operators now have the ability to freely control and adapt their network architecture, leading to highly customized systems and fine-grained optimization.
-As a result, Software-Defined Networking (SDN) has emerged as a new networking paradigm.
-
-Moving away from the principle of distributed communication and routing, SDN introduces the notion of "centralized management".
-A single controller with global knowledge is able to automatically modify and adapt the forwarding tables of all switches in the network, while notifying end hosts of changes in the network.
-These two new trends in system design facilitated impactful new innovation opportunities in the space of TCP congestion research.
-Traffic can now be managed in a  \textit{centralised} fashion based on \textit{global knowledge} of the entire topology and traffic patterns.
-
-A new line of centralised schedulers has emerged that  can achieve close to optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
-However, these schedulers are still  \textit{reactive}  in nature.
-The central controller responds to changes in the network or requests by applications, which may cost valuable roundtrip latency.
-Often, short-term flows or bursts are unaccounted for, which causes undesirable packet loss and backpropagating congestion.
-
-A much more desirable solution is a global, centralised arbiter which is able to predict and fairly distribute flows in the network before bursts or congestion occur.
-By treating the network's compute and forwarding power as a single finite resource, a controller acts like the OS scheduler distributing CPU time slices to processes.
-This design approach follows SDN's aspiration of introducing operating systems abstractions to the networking domain space.
-
-In this project, we plan to explore the possibilities of a centralised, proactive flow scheduler. We ask ourselves the following research questions:
-\begin{enumerate}
-	\item Is it possible to design a centralised token-based scheduling network?
-	\item Is it possible to predict traffic and preemptively schedule flows and token distribution in a datacenter context?
-	\item Using this approach, are we able to achieve better performance and utilization than existing solutions?
-\end{enumerate}
-
-In the scope of this course, we attempt to answer question one and design a simple token-based scheduler in Mininet.
-If successful, we will benchmark our results and evaluate the level of utilization compared to contemporary scheduling systems.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}
-\label{sec:related}
-
-Most of the work in congestion control relies on a centralised scheduler which periodically gets updates about the flows in the network and reserves specific paths for long running flows.
-
-MicroTE~\cite{microte} does predictive OpenFlow routing and scheduling using a central SDN controller.
-Hedera~\cite{hedera} schedules large so called "elephant" flows by gathering switch statistics.
-However it pays poor attention to short term flows and network latency.
-FastPass~\cite{fastpass} on the other hand, proposes a zero-queue network and promises a very low latency network.
-The feasibility of this proposal remains questionable when considering the high cost of implementation.
-
-Recent proposals on decentralised scheduling like ExpressPass~\cite{expresspass} have heavily influenced our research.
-ExpressPass posits a credit based system similar to TCPTokens in order to control the congestion in the network.
-Effecting flow rerouting using ML techniques in KDN~\cite{kdn} and other works shows promising signs of further improvements in this area.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Design / Proposed Approach}
-\label{sec:design}
-
-\subsection{Simple System}
-In our initial simple TCPToken design, a centralized controller regulates all node traffic by provisioning end-hosts with tokens.
-These tokens act as the "currency" of the system.
-The total amount is fixed to some function of the bisection and aggregate bandwidths available in the network, and the traffic window size of sending end hosts is calculated based on the host's current token availability.
-If a node or switch is overburdened, it may notify the scheduler, which will adjust the traffic window of responsible sending nodes remotely.
-Management and control operations including token distribution and updates are priority-queued to guarantee a fast and low-latency response.
-If an end-host requires more bandwidth, it can also notify the controller, which may comply based on priority and availability.
-Initially, the controller will compute optimal route configuration based on the topology and link bandwidth using a simple heuristic bin-packing approach.
-End-hosts will be initialized with a fixed low-to-medium bandwidth guarantee that will be adjusted over time.
-
-\subsection{Advanced System}
-The aforementioned system is intended only as a proof of concept.
-It is still a  synchronous and reactive approach, dependent on notifications and requests by overburdened switches and underprovisioned hosts.
-Ideally, all management in the TCPToken system should be asynchronous and preemptive. 
-\subsubsection{ Paying for traffic}
-In a complete design, tokens will be used as a transport pass instead of simply acting as traffic shaping parameters.
-On a per-flow basis, nodes will be able to attach tokens to traffic as a form of payment.
-Enforcement is performed on the switch level, any flow that is not "paid" by a token will be dropped.
-Nodes may be handed different types of tokens to serve load-balancing and traffic shaping purposes.
-Applications operating on traditionally short-lived flows and bursts may sign their traffic with a different token type than long-lived permanent processes.
-Switches will forward traffic and balance flows accordingly with only minor interference from the central controller.
-
-\subsubsection{Predicting traffic}
-ML-based scheduling is an emerging research field, also known as Knowledge-Defined Networking. Datacenter traffic follows patterns, which can be predicted using statistical methods.
-Combined with using tokens as a traffic engineering enforcement, a preemptive scheduling strategy in datacenters may be feasible.
-
-We do not intend to implement the advanced system for this project, but are using it as a guideline to motivate our research.
-Based on the experiences and knowledge gained from implementing the basic system, we will reevaluate the feasibility and practicality of the second concept.
-
-\subsubsection{Fault-Tolerance and Scalability}
-To handle fault-tolerance we plan to integrate a fallback-mechanism to conventional TCP, if the node has not received a TCPToken in a prefixed amount of time.
-End-hosts will only be marginally affected by a controller or link failure, as they are guaranteed a fixed amount of bandwidth based on their token availability.
-
-To avoid overburdening the controller, we envision a distributed hive of schedulers (e.g., Onos) each managing their own region in the final design.~\cite{onos} In case of failure, a shadow node or neighbouring device may take over computation temporarily, until the main node has recovered.~\cite{disco}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{sec:implementation}
-We intend to emulate our system in Mininet~\cite{mininet} to observe traffic patterns and infer a suitable token algorithm.
-Mininet has proven itself to be a viable tool to model new congestion control algorithms~\cite{mininet_learning}, and will help us prototype our concept efficiently.
-We will build a custom SDN controller that interacts with traditional OpenFlow software switches as well as end-hosts.
-End-hosts will run a custom real-world traffic generation script which adjusts based on information packets sent by the controller.
-If time permits, we may expand our implementation to MaxiNet, which can emulate large-scale network stress tests of thousands of nodes on multiple physical hosts.
+We intend to emulate our system in Mininet~\cite{mininet} to observe traffic 
+patterns and infer a suitable token algorithm. Mininet has proven itself to be 
+a viable tool to model new congestion control 
+algorithms~\cite{mininet_learning}, and will help us prototype our concept 
+efficiently. We will build a custom SDN controller that interacts with 
+traditional OpenFlow software switches as well as end-hosts. End-hosts will run 
+a custom real-world traffic generation script which adjusts based on 
+information packets sent by the controller.
+If time permits, we may expand our implementation to MaxiNet, which can emulate 
+large-scale network stress tests on multiple physical hosts.
 
-We initially considered a second implementation alternative in C/C++ based on the FastPass~\cite{fastpass} source code.
-This would provide us with a fully deployable system which we could fork our implementation from.
-A major advantage of this approach is the ability to test scenarios and traffic algorithms using real software code.
+We initially considered a second implementation alternative in C/C++ based on 
+the FastPass~\cite{fastpass} source code. This would provide us with a fully 
+deployable system which we could fork our implementation from. A major 
+advantage of this approach is the ability to test scenarios and traffic 
+algorithms using real software code.
 However, several concerns made us favour a Mininet emulation instead.
-Firstly, FastPass relies on DPDK integration, which requires actual hardware interfaces.
-The central arbiter in the FastPass design would need to run on a dedicated machine, which increases prototyping and development complexity substantially.
+Firstly, FastPass relies on DPDK integration, which requires actual hardware 
+interfaces. The central arbiter in the FastPass design would need to run on a 
+dedicated machine, which increases prototyping and development complexity 
+substantially.
 
-Secondly, the FastPass code is highly specialized and optimized research code with only little available documentation.
-Modifying and evolving the source code will require thorough understanding of kernel and networking development, a significant time-sink.
-For a class project, these may be major initial hurdles, taking away from the research aspect of the design concept.
-Consequently, we have decided to pursue an approach which allows us to quickly develop an understanding of the problem without being obstructed by engineering work.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Evaluation}
-\label{sec:evaluation}
-To evaluate the effectiveness of our system, we plan to measure against existing centralized as well as decentralized solutions.
-We will run datacenter traffic on a fat-tree topology of various depth and observe the effect each congestion algorithm will have on the overall bandwidth utilization and latency.
-The centralized design will be based on Hedera~\cite{hedera}, a common and influential datacenter scheduler.
-
-The decentralized congestion control mechanism will be DCTCP~\cite{dctcp}, a state-of-the-art TCP congestion algorithm.
-As baseline we will also compare to the default TCP congestion algorithm to estimate how our scheduler fares against the default case.
-As our initial simple design is very constrained in its ability to route and control traffic we expect to achieve less optimality against DCTCP and Hedera, but gain a substantial advantage over common TCP.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Timeline}
-\label{sec:timeline}
-\begin{itemize}
-	\item \underline{October 15th:} Finalize draft and related work / background section.
-	\item \underline{ October 22nd:} Iterate over literature and thoroughly analyze related work. Set up an initial prototyping environment in Mininet on which each team member can work on.
-	\item \underline{October 30th:} Finalize the initial simple design and explore naive congestion modelling in Mininet. Integrate existing Mininet emulations of Hedera and DCTCP.
-	\item \underline{November 12th:} Develop initial naive TCPTokens prototype involving controller and end-host design.
-	\item \underline{November 19th:} Develop intelligent centralised congestion control heuristic.
-	\item \underline{November 26th:} Run evaluations and microbenchmarks.
-	\item \underline{December 3rd:} Iterate and improve the scheduler, explore topology and bandwidth constraints. Test at scale.
-	\item \underline{December 10th:} Write report and prepare presentation.
-\end{itemize}
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{paper}
-
-\end{document}
+Secondly, the FastPass code is highly specialized and optimized research code 
+with only little available documentation. Modifying and evolving the source 
+code will require thorough understanding of kernel and networking development, 
+a significant time-sink. For a class project, these may be major initial 
+hurdles, taking away from the research aspect of the design concept.
+Consequently, we have decided to pursue an approach which allows us to quickly 
+develop an understanding of the problem without being obstructed by engineering 
+work.

### 2017-12-09 Documents/Report/intro.tex
Update report structure.

--- 2017-10-10 Proposal/paper.tex
@@
-\documentclass[sigconf]{acmart}
-
-\usepackage{booktabs} % For formal tables
-\usepackage[normalem]{ulem}
-% Copyright
-\setcopyright{none}
-
-\input{annotations}
-
-\begin{document}
-
-\title{TCPTokens: Introducing Currency into Datacenter Congestion Control}
-
-\author{Anand Jayarajan}
-\affiliation{%
-  \institution{University of British Columbia}
-}
-\email{anandj@cs.ubc.ca}
-
-\author{Michael Przystupa}
-\affiliation{%
-  \institution{University of British Columbia}
-}
-\email{michael.przystupa@gmail.com}
-
-\author{Robert Reiss}
-\affiliation{%
-  \institution{University of British Columbia}
-}
-\email{rreiss@cs.ubc.ca}
-
-\author{Fabian Ruffy}
-\affiliation{%
-  \institution{University of British Columbia}
-}
-\email{fruffy@cs.ubc.ca}
-
-\begin{abstract}
-Datacenter networks have become a hotbed for research in the recent past.
-Many works have been published on improving the  bisection bandwidth, reducing cost, and latency.
-As a core contribution of this research, centralized scheduling has entered the stage as dominant strategy to manage flows.
-Benefitting from a global view and full control over the network, centralized schedulers are able to enforce fine-grained traffic control, frequently achieving near-optimal bandwidth optimization.
-However, modern congestion control algorithms and schedulers are still fundamentally reactive.
-Many algorithms are designed to only respond to packet loss or significant increase in latency, not to actively prevent it.
-
-In this project, we plan to explore the opportunities of employing a preemptive and tightly controlling central network scheduler.
-Using tokens as a global tool to enforce traffic limiting and policies, the scheduler is able to proactively control the flow of traffic as well as respond to application requests.
-"Knowledge" and predictive analysis in networks is a growing trend in research which we intend to leverage in this system.
-
-We investigate the potential in such a new form of interactive congestion control and analyze it against state-of-the-art solutions.
-Our tool of choice to model our system is Mininet, a rapid prototyping emulator for datacenter networks.
-We will compare our "TCPToken" system against established TCP-congestion systems such as Hedera and DCTCP.
-Based on the findings and measurement results, we will reassess the feasibility and prospects of a token-based, predictive congestion control algorithm.
-\end{abstract}
-
-\keywords{TCP, Congestion Control, SDN, Datacenter}
-
-\maketitle
-
-%\input{intro}
-%
-%\input{relatedworks}
-%\input{design}
-% %\input{motivation}
-%\input{implementation}
-%\input{eval}
-%\input{timeline}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
+\section{Introduction / Background}
 \label{sec:intro}
 
-Initially, research in network congestion was dominated by the assumption of a decentralised, autonomous network - end-hosts only have control over the amount of traffic they send, and are unaware of the intentions or traffic rates of their peers.
-Similarly, switches and routers are unaware of global traffic patterns and only forward based on their local notion of optimality.
+%Initially, research in network congestion was dominated by the assumption of a 
+%decentralized, autonomous network - end-hosts only have control over the 
+%amount 
+%of traffic they send, and are unaware of the intentions or traffic rates of 
+%their peers. Similarly, switches and routers are unaware of global traffic 
+%patterns and only forward based on their local notion of optimality.
+%
+%In line with these assumptions, TCP was designed to optimize traffic globally 
+%on a simple fairness principle: nodes react to packet loss and assume that 
+%others will behave accordingly. An equilibrium is reached when all end-hosts 
+%achieve a traffic rate that, in sum, conforms to the maximum available 
+%bandwidth of the most congested link.
+%TCP works well  in scenarios where many different distrustful participants 
+%compete for limited bandwidth. Still, TCP is a \textit{reactive protocol}; the 
+%fact that packet loss and latency increases occur in the network already 
+%indicates a problem. Packet loss is primarily due to overflowing queues in 
+%forwarding elements or mismatched hardware capabilities, implying that traffic 
+%has not been optimally distributed. Ideally, a network should always be 
+%"zero-queue", i.e., latency will merely be induced by propagation, and not 
+%queuing delay.
+%
+%Queueing has commonly not been a dominating issue in wide-area and enterprise 
+%networks, as traffic is sufficiently distributed and diverse, with only few 
+%"hot" target hosts.~\cite{hedera, microte} Traffic optimization is a 
+%substantial challenge; network operators have no control over the individual 
+%network elements nor its participants. Under these conditions, TCP and its 
+%extensions can be considered a best-effort solution to globally maximize 
+%utilization.
 
-In line with these assumptions, TCP was designed to optimize traffic globally on a simple fairness principle: nodes react to packet loss and assume that others will behave accordingly.
-An equilibrium is reached when all end-hosts achieve a traffic rate that, in sum, conforms to the maximum available bandwidth of the most congested link.
-TCP works well  in scenarios where many different distrustful participants compete for limited bandwidth.
-Still, TCP is a \textit{reactive protocol}; the fact that packet loss and latency increases occur in the network already indicates a problem.
-Packet loss is primarily due to overflowing queues in forwarding elements, implying that traffic has not been optimally distributed.
-Ideally, a network should always be "zero-queue", i.e., latency will merely be induced by propagation, and not queuing delay.
+Developments in the past decade have changed the general networking 
+environment. Data centers have emerged as an exciting new research frontier, 
+posing novel design challenges and opportunities. Driven by minimization of 
+costs and maximization of compute power, data centers must run at the highest 
+possible utilization to achieve the ideal compute/cost ratio.
+Optimizing the distribution of traffic and simultaneously guaranteeing fairness 
+is a perennial challenge for any network operator.
+Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the 
+eventual collapse of a high-load network, requiring sophisticated 
+approaches to solve congestion control.
+Despite the innovation potential of data centers, TCP has dominated as 
+the congestion control protocol of choice.
+TCP is a \textit{reactive protocol}, responding to indicators of congestion and 
+latency in the network. However, the fact that packet loss and latency 
+surges occur in the network, already portends a problem. Packet loss is 
+incurred by overflowing queues in forwarding elements or mismatched 
+hardware capabilities, implying that traffic has not been optimally 
+distributed. Ideally, a network should always be "zero-queue", i.e., latency 
+will merely be induced by propagation, and not queuing delay.
 
-Queueing has generally not been a dominating issue in wide-area and enterprise networks, as traffic is sufficiently distributed and diverse, with only few "hot" target hosts.~\cite{hedera, microte} Traffic optimization is a substantial challenge; network operators have no control over the individual network elements nor its participants.
-Under these conditions, TCP and its extensions can be considered a best-effort solution.
+With the advent of Software-Defined Networking (SDN), operators now have the 
+ability to freely control and adapt their network architecture, leading to 
+highly customized systems and fine-grained optimization.~\cite{sdn_road}
 
-Developments in the past decade have changed the general networking environment.
-Datacenters have emerged as an exciting new research frontier, posing novel design challenges and opportunities.Driven by minimization of costs and maximization of compute power, datacenters must  run at maximum utilization to achieve an optimal compute/cost ratio.
-Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load network, requiring more sophisticated approaches to solve congestion control.
-On the other hand, operators now have the ability to freely control and adapt their network architecture, leading to highly customized systems and fine-grained optimization.
-As a result, Software-Defined Networking (SDN) has emerged as a new networking paradigm.
+Moving away from the principle of distributed communication and routing, SDN 
+introduces the notion of "centralized management". A single controller with 
+global knowledge is able to automatically modify and adapt the forwarding 
+tables of all switches in the network, while notifying end hosts of changes in 
+the network.
+These two new trends in system design; full architectural control and 
+centralized management, facilitated new opportunities in the space of TCP 
+congestion research. Traffic can now be managed in a  \textit{centralized} 
+fashion based on \textit{global knowledge} of the entire topology and traffic 
+patterns.
 
-Moving away from the principle of distributed communication and routing, SDN introduces the notion of "centralized management".
-A single controller with global knowledge is able to automatically modify and adapt the forwarding tables of all switches in the network, while notifying end hosts of changes in the network.
-These two new trends in system design facilitated impactful new innovation opportunities in the space of TCP congestion research.
-Traffic can now be managed in a  \textit{centralised} fashion based on \textit{global knowledge} of the entire topology and traffic patterns.
+A new line of centralized schedulers has emerged that can achieve close to 
+optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
+However, these schedulers are still  \textit{reactive}  in nature. The central 
+controller responds to changes in the network or requests by applications, 
+which may cost valuable round-trip latency. Often, short-term flows or bursts 
+are unaccounted for, which causes undesirable packet loss and back-propagating 
+congestion.
 
-A new line of centralised schedulers has emerged that  can achieve close to optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
-However, these schedulers are still  \textit{reactive}  in nature.
-The central controller responds to changes in the network or requests by applications, which may cost valuable roundtrip latency.
-Often, short-term flows or bursts are unaccounted for, which causes undesirable packet loss and backpropagating congestion.
+The idea of admission control and service guarantees in networks is not 
+new.~\cite{access_limit, access_limit2}. However, such designs traditionally 
+aim to assure quality and bandwidth guarantees in a contentious, decentralized, 
+and untrusted environments such as the internet. In a datacenter, these 
+conditions do not apply. End-hosts are generally considered reliable and 
+restricted in behavior, which allows for great simplification of enforcement 
+and prioritization policies.
 
-A much more desirable solution is a global, centralised arbiter which is able to predict and fairly distribute flows in the network before bursts or congestion occur.
-By treating the network's compute and forwarding power as a single finite resource, a controller acts like the OS scheduler distributing CPU time slices to processes.
-This design approach follows SDN's aspiration of introducing operating systems abstractions to the networking domain space.
+A desirable solution is a global, centralized arbiter which is able to predict 
+and fairly distribute flows in the network before bursts or congestion even 
+occur. By treating the network's compute and forwarding power as a single 
+finite resource, a controller acts akin to a OS scheduler distributing CPU time 
+slices to processes. This design approach follows SDN's aspiration of 
+introducing operating systems abstractions to the networking domain space.
 
-In this project, we plan to explore the possibilities of a centralised, proactive flow scheduler. We ask ourselves the following research questions:
+In this project, we plan to explore the possibilities of a centralized, 
+proactive flow scheduler. We ask ourselves the following research questions:
 \begin{enumerate}
-	\item Is it possible to design a centralised token-based scheduling network?
-	\item Is it possible to predict traffic and preemptively schedule flows and token distribution in a datacenter context?
-	\item Using this approach, are we able to achieve better performance and utilization than existing solutions?
+    \item What are the requirements for such a centralized, predictive 
+    scheduler to succeed?
+    \item Is it possible to preemptively regulate a network by analyzing global 
+    traffic patterns?
+    \item What latency, packet loss and utilization are we able to achieve?
 \end{enumerate}
 
-In the scope of this course, we attempt to answer question one and design a simple token-based scheduler in Mininet.
-If successful, we will benchmark our results and evaluate the level of utilization compared to contemporary scheduling systems.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}
-\label{sec:related}
-
-Most of the work in congestion control relies on a centralised scheduler which periodically gets updates about the flows in the network and reserves specific paths for long running flows.
-
-MicroTE~\cite{microte} does predictive OpenFlow routing and scheduling using a central SDN controller.
-Hedera~\cite{hedera} schedules large so called "elephant" flows by gathering switch statistics.
-However it pays poor attention to short term flows and network latency.
-FastPass~\cite{fastpass} on the other hand, proposes a zero-queue network and promises a very low latency network.
-The feasibility of this proposal remains questionable when considering the high cost of implementation.
-
-Recent proposals on decentralised scheduling like ExpressPass~\cite{expresspass} have heavily influenced our research.
-ExpressPass posits a credit based system similar to TCPTokens in order to control the congestion in the network.
-Effecting flow rerouting using ML techniques in KDN~\cite{kdn} and other works shows promising signs of further improvements in this area.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Design / Proposed Approach}
-\label{sec:design}
-
-\subsection{Simple System}
-In our initial simple TCPToken design, a centralized controller regulates all node traffic by provisioning end-hosts with tokens.
-These tokens act as the "currency" of the system.
-The total amount is fixed to some function of the bisection and aggregate bandwidths available in the network, and the traffic window size of sending end hosts is calculated based on the host's current token availability.
-If a node or switch is overburdened, it may notify the scheduler, which will adjust the traffic window of responsible sending nodes remotely.
-Management and control operations including token distribution and updates are priority-queued to guarantee a fast and low-latency response.
-If an end-host requires more bandwidth, it can also notify the controller, which may comply based on priority and availability.
-Initially, the controller will compute optimal route configuration based on the topology and link bandwidth using a simple heuristic bin-packing approach.
-End-hosts will be initialized with a fixed low-to-medium bandwidth guarantee that will be adjusted over time.
-
-\subsection{Advanced System}
-The aforementioned system is intended only as a proof of concept.
-It is still a  synchronous and reactive approach, dependent on notifications and requests by overburdened switches and underprovisioned hosts.
-Ideally, all management in the TCPToken system should be asynchronous and preemptive. 
-\subsubsection{ Paying for traffic}
-In a complete design, tokens will be used as a transport pass instead of simply acting as traffic shaping parameters.
-On a per-flow basis, nodes will be able to attach tokens to traffic as a form of payment.
-Enforcement is performed on the switch level, any flow that is not "paid" by a token will be dropped.
-Nodes may be handed different types of tokens to serve load-balancing and traffic shaping purposes.
-Applications operating on traditionally short-lived flows and bursts may sign their traffic with a different token type than long-lived permanent processes.
-Switches will forward traffic and balance flows accordingly with only minor interference from the central controller.
-
-\subsubsection{Predicting traffic}
-ML-based scheduling is an emerging research field, also known as Knowledge-Defined Networking. Datacenter traffic follows patterns, which can be predicted using statistical methods.
-Combined with using tokens as a traffic engineering enforcement, a preemptive scheduling strategy in datacenters may be feasible.
-
-We do not intend to implement the advanced system for this project, but are using it as a guideline to motivate our research.
-Based on the experiences and knowledge gained from implementing the basic system, we will reevaluate the feasibility and practicality of the second concept.
-
-\subsubsection{Fault-Tolerance and Scalability}
-To handle fault-tolerance we plan to integrate a fallback-mechanism to conventional TCP, if the node has not received a TCPToken in a prefixed amount of time.
-End-hosts will only be marginally affected by a controller or link failure, as they are guaranteed a fixed amount of bandwidth based on their token availability.
-
-To avoid overburdening the controller, we envision a distributed hive of schedulers (e.g., Onos) each managing their own region in the final design.~\cite{onos} In case of failure, a shadow node or neighbouring device may take over computation temporarily, until the main node has recovered.~\cite{disco}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Implementation}
-\label{sec:implementation}
-We intend to emulate our system in Mininet~\cite{mininet} to observe traffic patterns and infer a suitable token algorithm.
-Mininet has proven itself to be a viable tool to model new congestion control algorithms~\cite{mininet_learning}, and will help us prototype our concept efficiently.
-We will build a custom SDN controller that interacts with traditional OpenFlow software switches as well as end-hosts.
-End-hosts will run a custom real-world traffic generation script which adjusts based on information packets sent by the controller.
-If time permits, we may expand our implementation to MaxiNet, which can emulate large-scale network stress tests of thousands of nodes on multiple physical hosts.
-
-We initially considered a second implementation alternative in C/C++ based on the FastPass~\cite{fastpass} source code.
-This would provide us with a fully deployable system which we could fork our implementation from.
-A major advantage of this approach is the ability to test scenarios and traffic algorithms using real software code.
-However, several concerns made us favour a Mininet emulation instead.
-Firstly, FastPass relies on DPDK integration, which requires actual hardware interfaces.
-The central arbiter in the FastPass design would need to run on a dedicated machine, which increases prototyping and development complexity substantially.
-
-Secondly, the FastPass code is highly specialized and optimized research code with only little available documentation.
-Modifying and evolving the source code will require thorough understanding of kernel and networking development, a significant time-sink.
-For a class project, these may be major initial hurdles, taking away from the research aspect of the design concept.
-Consequently, we have decided to pursue an approach which allows us to quickly develop an understanding of the problem without being obstructed by engineering work.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Evaluation}
-\label{sec:evaluation}
-To evaluate the effectiveness of our system, we plan to measure against existing centralized as well as decentralized solutions.
-We will run datacenter traffic on a fat-tree topology of various depth and observe the effect each congestion algorithm will have on the overall bandwidth utilization and latency.
-The centralized design will be based on Hedera~\cite{hedera}, a common and influential datacenter scheduler.
-
-The decentralized congestion control mechanism will be DCTCP~\cite{dctcp}, a state-of-the-art TCP congestion algorithm.
-As baseline we will also compare to the default TCP congestion algorithm to estimate how our scheduler fares against the default case.
-As our initial simple design is very constrained in its ability to route and control traffic we expect to achieve less optimality against DCTCP and Hedera, but gain a substantial advantage over common TCP.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Timeline}
-\label{sec:timeline}
-\begin{itemize}
-	\item \underline{October 15th:} Finalize draft and related work / background section.
-	\item \underline{ October 22nd:} Iterate over literature and thoroughly analyze related work. Set up an initial prototyping environment in Mininet on which each team member can work on.
-	\item \underline{October 30th:} Finalize the initial simple design and explore naive congestion modelling in Mininet. Integrate existing Mininet emulations of Hedera and DCTCP.
-	\item \underline{November 12th:} Develop initial naive TCPTokens prototype involving controller and end-host design.
-	\item \underline{November 19th:} Develop intelligent centralised congestion control heuristic.
-	\item \underline{November 26th:} Run evaluations and microbenchmarks.
-	\item \underline{December 3rd:} Iterate and improve the scheduler, explore topology and bandwidth constraints. Test at scale.
-	\item \underline{December 10th:} Write report and prepare presentation.
-\end{itemize}
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{paper}
-
-\end{document}
+In the scope of this course, we attempt to answer these questions and design a 
+simple predictive scheduler in Mininet. If successful, we will benchmark our 
+results and evaluate the level of utilization compared to contemporary 
+scheduling systems.

### 2017-12-09 Iroko/LearningAgentv2.py
Learning agent now loads actor and critic

@@ -15,7 +15,7 @@ class LearningAgentv2:
         self.lam = lam
         self.defaultmax = defaultmax
         self.initmax = initMax
-        self.controller = DDPG(gamma,memory,5 + 16, 1,tau=.25) 
+        self.controller = DDPG(gamma,memory,5 + 16, 1,tau=.25, criticpath='critic', actorpath='actor') 
         
 
 

### 2017-12-10 Documents/Report/intro.tex
More graphs

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction / Background}
+\section{Introduction}
 \label{sec:intro}
@@
 is a perennial challenge for any network operator.
-Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the 
-eventual collapse of a high-load network, requiring sophisticated 
-approaches to solve congestion control.
+
+Inefficient routing and congestion control can quickly lead to 
+bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load 
+network, requiring more sophisticated solutions. In fact, recent insights into 
+datacenter network traffic have shown that buffer overruns and small 
+short-lived bursts are the primary factor of loss.\cite{fb_dc,msr_dc}
 Despite the innovation potential of data centers, TCP has dominated as 
 the congestion control protocol of choice.
-TCP is a \textit{reactive protocol}, responding to indicators of congestion and 
-latency in the network. However, the fact that packet loss and latency 
-surges occur in the network, already portends a problem. Packet loss is 
-incurred by overflowing queues in forwarding elements or mismatched 
-hardware capabilities, implying that traffic has not been optimally 
-distributed. Ideally, a network should always be "zero-queue", i.e., latency 
+TCP is a \textit{reactive protocol}, responding to indicators of packet drops
+latency in the network. However, the fact that loss and latency 
+surge in the network, already portends a problem. Overflowing queues in 
+forwarding elements or mismatched hardware capabilities imply that traffic has 
+not been optimally managed.
+Ideally, a network should always be "zero-queue", i.e., latency 
 will merely be induced by propagation, and not queuing delay.
 
+Recent research insights and contributions may aid in achieving such a data 
+center network.
+
+\paragraph{Software-Defined Networking}
 With the advent of Software-Defined Networking (SDN), operators now have the 
@@
 the network.
-These two new trends in system design; full architectural control and 
-centralized management, facilitated new opportunities in the space of TCP 
-congestion research. Traffic can now be managed in a  \textit{centralized} 
-fashion based on \textit{global knowledge} of the entire topology and traffic 
-patterns.
+This full architectural control and centralized management facilitated new 
+opportunities in the space of TCP congestion research. Traffic can now be 
+managed in a  \textit{centralized} fashion based on \textit{global knowledge} 
+of the entire topology and traffic patterns.
 
-A new line of centralized schedulers has emerged that can achieve close to 
-optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
+\paragraph{Centralized Scheduling and Congestion Control}
+Following the rise of SDN, a new line of centralized schedulers that can 
+achieve close to optimal bandwidth utilization has emerged.~\cite{hedera, 
+fastpass, microte, b4, dionysus}
 However, these schedulers are still \textit{reactive} in nature. The central 
@@
 are unaccounted for, which causes undesirable packet loss and back-propagating 
-congestion.
+congestion.\cite{perc}
 
+\paragraph{Admission-control systems}
+In parallel, admission-control-based approaches have gained traction as a 
+viable alternative to the conventional burst-and-backoff mechanism of 
+TCP.~\cite{expresspass, fastpass, perc}
 The idea of admission control and service guarantees in networks is not 
-new.~\cite{access_limit, access_limit2}. However, such designs traditionally 
+new.~\cite{access_limit, access_limit2}. These designs traditionally 
 aim to assure quality and bandwidth guarantees in a contentious, decentralized, 
@@
 conditions do not apply. End-hosts are generally considered reliable and 
-restricted in behavior, which allows for great simplification of enforcement 
-and prioritization policies.
+restricted in behaviour, which allows for great simplification of enforcement 
+and prioritization policies. Admission control systems give administrators the 
+benefit of tight control over the datacenter, mitigating the effects of 
+queue-buildup and bursts.
 
-A desirable solution is a global, centralized arbiter which is able to predict 
-and fairly distribute flows in the network before bursts or congestion even 
-occur. By treating the network's compute and forwarding power as a single 
-finite resource, a controller acts akin to a OS scheduler distributing CPU time 
-slices to processes. This design approach follows SDN's aspiration of 
-introducing operating systems abstractions to the networking domain space.
+\paragraph{Predictability of traffic}
+Remy\cite{remy} has shown that it is possible to learn optimal congestion 
+control parameters for various network topologies, but 
+Due the scarcity of public data center traffic sets, it is unclear how 
+predictable traffic truly is.
+
+A desirable solution is a global, centralized arbiter, able to predict and 
+fairly distribute flows in the network before bursts or congestion even occur. 
+By treating the network's compute and forwarding power as a single finite 
+resource, a controller acts akin to a OS scheduler distributing CPU time slices 
+to processes.
 

### 2017-12-10 Documents/Report/motivation.tex
More graphs

@@
 \label{sec:motivation}
-Sed venenatis odio a augue porttitor, vel blandit tortor luctus.
-Pellentesque condimentum ante lacus, imperdiet commodo odio mattis a.
-Donec non consectetur libero.
-Donec efficitur erat nec dui convallis, sed efficitur nulla pharetra.
-Nullam sed feugiat est.
-Vestibulum est nisi, venenatis ac augue molestie, malesuada dignissim dolor.
-Aenean nisl ante, accumsan nec sagittis quis, euismod eleifend purus.
-Duis laoreet neque eros, non convallis nunc viverra eget.
-Sed eleifend neque neque, non faucibus justo vulputate vel.
-Fusce aliquam ultricies ligula ut efficitur.
-Nunc ut neque tristique, posuere sem vitae, consectetur nisl.
+Several recent contributions in networking research
 
-Donec pretium ut enim ac congue.
-Suspendisse nisl est, malesuada sit amet posuere a, gravida vitae lacus.
-Suspendisse et risus mauris.
-Nam imperdiet sit amet augue sed sodales.
-Quisque ante orci, aliquet non interdum nec, gravida sed enim.
-Nunc interdum, purus et scelerisque malesuada, diam lorem efficitur orci, et consequat ligula augue sed lacus.
-Morbi malesuada eleifend diam eu rhoncus.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
+\paragraph{Software-Defined Networking}
+With the advent of Software-Defined Networking (SDN), operators now have the 
+ability to freely control and adapt their network architecture, leading to 
+highly customized systems and fine-grained optimization.~\cite{sdn_road}
+
+Moving away from the principle of distributed communication and routing, SDN 
+introduces the notion of "centralized management". A single controller with 
+global knowledge is able to automatically modify and adapt the forwarding 
+tables of all switches in the network, while notifying end hosts of changes in 
+the network.
+This full architectural control and centralized management facilitated new 
+opportunities in the space of TCP congestion research. Traffic can now be 
+managed in a  \textit{centralized} fashion based on \textit{global knowledge} 
+of the entire topology and traffic patterns.
+
+\paragraph{Centralized Scheduling and Congestion Control}
+A new line of centralized schedulers has emerged that can achieve close to 
+optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
+However, these schedulers are still  \textit{reactive}  in nature. The central 
+controller responds to changes in the network or requests by applications, 
+which may cost valuable round-trip latency. Often, short-term flows or bursts 
+are unaccounted for, which causes undesirable packet loss and back-propagating 
+congestion.
+
+\paragraph{Admission-control systems}
+In parallel, admission-control-based approaches have gained traction as a 
+viable alternative to the conventional burst-and-backoff mechanism of 
+TCP.~\cite{expresspass, fastpass, perc}
+The idea of admission control and service guarantees in networks is not 
+new.~\cite{access_limit, access_limit2}. However, such designs traditionally 
+aim to assure quality and bandwidth guarantees in a contentious, decentralized, 
+and untrusted environments such as the internet. In a datacenter, these 
+conditions do not apply. End-hosts are generally considered reliable and 
+restricted in behaviour, which allows for great simplification of enforcement 
+and prioritization policies.
+
+\paragraph{Predictability of traffic}

### 2017-12-10 Documents/Report/intro.tex
More intro writing.

@@
 \paragraph{Predictability of traffic}
-Remy\cite{remy} has shown that it is possible to learn optimal congestion 
-control parameters for various network topologies, but 
-Due the scarcity of public data center traffic sets, it is unclear how 
-predictable traffic truly is.
+Although it is unclear how predictable data center traffic truly is, various 
+works have shown that repeated patterns may 
+exist.\cite{msr_dc,fb_dc,traffic,microte}
+If data center traffic exhibits predictable patterns over long time, it may be 
+possible to devise a proactive mechanism which predicts the traffic matrix of 
+the next nth iterations, and adjusts traffic accordingly.
+
+%Winstein and Balakrishnan,~2013\cite{remy} demonstrated that it is possible to 
+%learn optimal TCP congestion control parameters for various network 
+%topologies. 
+%Benson et al.,~2014\cite{microte} analyzed data center traffic and inferred 
+%that around 30\%
 

### 2017-12-10 Documents/Report/conclusion.tex
Added discussion, future work and conclusion. Sparse thoughts for first two.

--- 2017-12-09 Documents/Report/intro.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction / Background}
-\label{sec:intro}
-
-%Initially, research in network congestion was dominated by the assumption of a 
-%decentralized, autonomous network - end-hosts only have control over the 
-%amount 
-%of traffic they send, and are unaware of the intentions or traffic rates of 
-%their peers. Similarly, switches and routers are unaware of global traffic 
-%patterns and only forward based on their local notion of optimality.
-%
-%In line with these assumptions, TCP was designed to optimize traffic globally 
-%on a simple fairness principle: nodes react to packet loss and assume that 
-%others will behave accordingly. An equilibrium is reached when all end-hosts 
-%achieve a traffic rate that, in sum, conforms to the maximum available 
-%bandwidth of the most congested link.
-%TCP works well  in scenarios where many different distrustful participants 
-%compete for limited bandwidth. Still, TCP is a \textit{reactive protocol}; the 
-%fact that packet loss and latency increases occur in the network already 
-%indicates a problem. Packet loss is primarily due to overflowing queues in 
-%forwarding elements or mismatched hardware capabilities, implying that traffic 
-%has not been optimally distributed. Ideally, a network should always be 
-%"zero-queue", i.e., latency will merely be induced by propagation, and not 
-%queuing delay.
-%
-%Queueing has commonly not been a dominating issue in wide-area and enterprise 
-%networks, as traffic is sufficiently distributed and diverse, with only few 
-%"hot" target hosts.~\cite{hedera, microte} Traffic optimization is a 
-%substantial challenge; network operators have no control over the individual 
-%network elements nor its participants. Under these conditions, TCP and its 
-%extensions can be considered a best-effort solution to globally maximize 
-%utilization.
-
-Developments in the past decade have changed the general networking 
-environment. Data centers have emerged as an exciting new research frontier, 
-posing novel design challenges and opportunities. Driven by minimization of 
-costs and maximization of compute power, data centers must run at the highest 
-possible utilization to achieve the ideal compute/cost ratio.
-Optimizing the distribution of traffic and simultaneously guaranteeing fairness 
-is a perennial challenge for any network operator.
-Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the 
-eventual collapse of a high-load network, requiring sophisticated 
-approaches to solve congestion control.
-Despite the innovation potential of data centers, TCP has dominated as 
-the congestion control protocol of choice.
-TCP is a \textit{reactive protocol}, responding to indicators of congestion and 
-latency in the network. However, the fact that packet loss and latency 
-surges occur in the network, already portends a problem. Packet loss is 
-incurred by overflowing queues in forwarding elements or mismatched 
-hardware capabilities, implying that traffic has not been optimally 
-distributed. Ideally, a network should always be "zero-queue", i.e., latency 
-will merely be induced by propagation, and not queuing delay.
-
-With the advent of Software-Defined Networking (SDN), operators now have the 
-ability to freely control and adapt their network architecture, leading to 
-highly customized systems and fine-grained optimization.~\cite{sdn_road}
-
-Moving away from the principle of distributed communication and routing, SDN 
-introduces the notion of "centralized management". A single controller with 
-global knowledge is able to automatically modify and adapt the forwarding 
-tables of all switches in the network, while notifying end hosts of changes in 
-the network.
-These two new trends in system design; full architectural control and 
-centralized management, facilitated new opportunities in the space of TCP 
-congestion research. Traffic can now be managed in a  \textit{centralized} 
-fashion based on \textit{global knowledge} of the entire topology and traffic 
-patterns.
-
-A new line of centralized schedulers has emerged that can achieve close to 
-optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
-However, these schedulers are still  \textit{reactive}  in nature. The central 
-controller responds to changes in the network or requests by applications, 
-which may cost valuable round-trip latency. Often, short-term flows or bursts 
-are unaccounted for, which causes undesirable packet loss and back-propagating 
-congestion.
-
-The idea of admission control and service guarantees in networks is not 
-new.~\cite{access_limit, access_limit2}. However, such designs traditionally 
-aim to assure quality and bandwidth guarantees in a contentious, decentralized, 
-and untrusted environments such as the internet. In a datacenter, these 
-conditions do not apply. End-hosts are generally considered reliable and 
-restricted in behavior, which allows for great simplification of enforcement 
-and prioritization policies.
-
-A desirable solution is a global, centralized arbiter which is able to predict 
-and fairly distribute flows in the network before bursts or congestion even 
-occur. By treating the network's compute and forwarding power as a single 
-finite resource, a controller acts akin to a OS scheduler distributing CPU time 
-slices to processes. This design approach follows SDN's aspiration of 
-introducing operating systems abstractions to the networking domain space.
-
-In this project, we plan to explore the possibilities of a centralized, 
-proactive flow scheduler. We ask ourselves the following research questions:
-\begin{enumerate}
-    \item What are the requirements for such a centralized, predictive 
-    scheduler to succeed?
-    \item Is it possible to preemptively regulate a network by analyzing global 
-    traffic patterns?
-    \item What latency, packet loss and utilization are we able to achieve?
-\end{enumerate}
-
-In the scope of this course, we attempt to answer these questions and design a 
-simple predictive scheduler in Mininet. If successful, we will benchmark our 
-results and evaluate the level of utilization compared to contemporary 
-scheduling systems.
+\section{Conclusion}
+\label{sec:conclusion}

### 2017-12-10 Documents/Report/discussion.tex
Added discussion, future work and conclusion. Sparse thoughts for first two.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
\label{sec:discussion}

Rather predictably, Iroko struggles on random traffic patterns, however we
do not see this as particularly concerning. Firstly \textit{Iroko} still outperforms
both DCTCP and ECMP implementations. Further, as these traffic patterns
were designed by the authors of the Hedera paper \cite{hedera}, it remains
likely these patterns are somewhat amenable to the Hedera implementation.
Finally random traffic patterns seem unlikely to be prevalent in a
datacenter context. While the general predictability of datacenter
traffic remains an area of open research, our results show that
if there is sufficient predictability, we do get results better than
random, and for truly random data, Iroko does no worse generally than
incumbent solutions.

Our machine learning approach shows promise, but machine learning remains
largely a black-box discipline. This nature makes it difficult for us
to determine precisely which parameters to use, and what drives
the improvements we've seen. Our current machine learning approach
is remarkably light-weight and does not require vast resources
in our test network. A heavier handed approach may yield better
results, however these may come at the cost of scalability.

### 2017-12-10 Documents/Report/future_work.tex
Added discussion, future work and conclusion. Sparse thoughts for first two.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future Work}
\label{sec:future}

Predicting datacenter traffic online with machine learning shows promise, and we believe that 
future widely deployed datacenter traffic control mechanisms will include
machine learning in their toolkit. However machine learning alone may not
be sufficient to achieve the best possible results. While we  
compared our results to Hedera, the work remains largely
orthogonal to our approach, and scheduling flows would likely 
complement machine learning based bandwidth control.

For large, long-running workloads that require significant 
network resources at the termination such as MAP-REDUCE
workloads, host may have very predictable completion times
for these jobs. In such cases, hosts could report these 
estimated completion times to the central controller,
allowing the controller to account for them pre-emptively
rather than trying to infer them which may take a couple
of epochs, and therefore reduce efficiency. Such a
system would be complementary to \textit{Iroko} and
allow for finer grained control.

Finally employing a token or credit based system similar
to that explored in \cite{expresspass} might yield excellent
results when coupled with our machine learning based predictive
central controller. This would allow hosts more ability to manage
their own bandwidth with intertemporal substitution of flows,
while still enforcing centralized predictive control.

### 2017-12-10 Documents/Report/Iroko_Report.tex
More graphs
report updated
more changes to intro

@@
 
-\title{TCPTokens: Introducing Currency into Datacenter Congestion Control}
+\title{Iroko: Predictive Datacenter Congestion Control}
 
@@
 \begin{abstract}
-Datacenter networks have become a hotbed for research in the recent past.
-Many works have been published on improving bisection bandwidth, cost, and latency.
-As a core contribution of this research, centralized scheduling has entered the stage as a dominant strategy to manage flows.
-Benefitting from a global view and full control over the network, centralized schedulers are able to enforce fine-grained traffic control, frequently achieving near-optimal bandwidth optimization.
-However, modern congestion control algorithms and schedulers are still fundamentally reactive.
-Many algorithms are designed to only respond to packet loss or significant increase in latency, not to actively prevent it.
-In this project, we plan to explore the opportunities of employing a preemptive and tightly controlling central network scheduler.
-Using tokens as a global tool to enforce traffic limits and admission control, the scheduler is able to proactively steer the flow of traffic.
-The notion of "knowledge" and predictive analysis in networks is a growing trend in research, which we intend to leverage in this system.
+Datacenter networks have become a hotbed of network research in the past 
+decade. Many works have been published on improving bisection bandwidth, cost, 
+and latency. As a core contribution of this research, centralized scheduling 
+has entered the stage as a dominant strategy to manage flows. Benefitting from 
+a global view and full control over the network, centralized schedulers are 
+able to enforce fine-grained traffic control, frequently achieving near-optimal 
+bandwidth optimization. However, modern congestion control algorithms and 
+schedulers are still fundamentally reactive. Many algorithms are designed to 
+only respond to packet loss or significant increase in latency, not to actively 
+prevent it.
+In this project, we plan to explore the opportunities of employing a preemptive 
+and tightly controlling central network scheduler. By enforcing traffic limits 
+and admission control, the scheduler is able to proactively steer the flow of 
+traffic. The notion of "knowledge" and predictive analysis in networks is a 
+growing trend in research, which we intend to leverage in this system.
 
-We investigate the potential in such a new form of interactive congestion control and analyze it against state-of-the-art solutions.
-Our tool of choice to model our system is Mininet, a rapid prototyping emulator for datacenter networks.
-We will compare our "Iroko" system against established TCP-congestion systems such as Hedera and DCTCP.
-Based on the findings and measurement results, we will reassess the feasibility and prospects of a predictive congestion control algorithm.
+We investigate the potential in such a new form of interactive congestion 
+control and analyze it against state-of-the-art solutions. Our tool of choice 
+to model our system is Mininet, a rapid prototyping emulator for datacenter 
+networks. We will compare our "Iroko" system against established TCP-congestion 
+systems such as Hedera and DCTCP. Based on the findings, measurement results, 
+and effectiveness of our reinforcement learning technique, we will reassess the 
+feasibility and prospects of a predictive congestion control algorithm.
 \end{abstract}
@@
 \input{intro}
-%\input{relatedwork}
+%\input{motivation}
 \input{design}
-% %\input{motivation}
+
 \input{implementation}
 \input{eval}
+\input{discussion}
+\input{future_work}
+\input{conclusion}
 
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{Iroko_Proposal}
+\bibliography{Iroko_Report}
 

### 2017-12-10 Documents/Report/intro.tex
more changes to intro

--- 2017-12-09 Documents/Report/intro.tex
@@
 is a perennial challenge for any network operator.
-Inefficient routing can quickly lead to bufferbloat~\cite{bufferbloat} and the 
-eventual collapse of a high-load network, requiring sophisticated 
-approaches to solve congestion control.
+
+Inefficient routing and congestion control can quickly lead to 
+bufferbloat~\cite{bufferbloat} and the eventual collapse of a high-load 
+network, requiring more sophisticated solutions. In fact, recent insights into 
+datacenter network traffic have shown that buffer overruns and small 
+short-lived bursts are the primary factor of loss.\cite{fb_dc,msr_dc}
 Despite the innovation potential of data centers, TCP has dominated as 
-the congestion control protocol of choice.
-TCP is a \textit{reactive protocol}, responding to indicators of congestion and 
-latency in the network. However, the fact that packet loss and latency 
-surges occur in the network, already portends a problem. Packet loss is 
-incurred by overflowing queues in forwarding elements or mismatched 
-hardware capabilities, implying that traffic has not been optimally 
-distributed. Ideally, a network should always be "zero-queue", i.e., latency 
+the congestion control algorithm of choice.
+TCP is a \textit{reactive protocol}, responding to indicators of packet drops
+latency. However, this surge of loss and latency in the network already 
+portends a problem. Overflowing queues in forwarding elements or mismatched 
+hardware capabilities imply that traffic has not been optimally managed.
+Ideally, a network should always be "zero-queue", i.e., latency 
 will merely be induced by propagation, and not queuing delay.
+
+While TCP has not fundamentally changed in
 
@@
 the network.
-These two new trends in system design; full architectural control and 
-centralized management, facilitated new opportunities in the space of TCP 
-congestion research. Traffic can now be managed in a  \textit{centralized} 
+This full architectural control facilitated new opportunities in the space of 
+TCP congestion research. Traffic can now be managed in a  \textit{centralized} 
 fashion based on \textit{global knowledge} of the entire topology and traffic 
@@
 
-A new line of centralized schedulers has emerged that can achieve close to 
-optimal bandwidth utilization.~\cite{hedera, fastpass, microte, b4, dionysus}
+Following SDN's rise in popularity, a new line of centralized schedulers that 
+can achieve close to optimal bandwidth utilization has emerged.~\cite{hedera, 
+fastpass, microte, b4, dionysus}
 However, these schedulers are still \textit{reactive} in nature. The central 
@@
 are unaccounted for, which causes undesirable packet loss and back-propagating 
-congestion.
+congestion.~\cite{perc}
 
-The idea of admission control and service guarantees in networks is not 
-new.~\cite{access_limit, access_limit2}. However, such designs traditionally 
+In parallel, admission-control-based approaches have gained traction as a 
+viable alternative to the conventional burst-and-backoff mechanism of 
+TCP.~\cite{expresspass, fastpass, perc}
+While idea of admission control and service guarantees in networks is not 
+new,~\cite{access_limit, access_limit2} these designs traditionally 
 aim to assure quality and bandwidth guarantees in a contentious, decentralized, 
-and untrusted environments such as the internet. In a datacenter, these 
+and untrusted environments such as the internet. In a data center, these 
 conditions do not apply. End-hosts are generally considered reliable and 
-restricted in behavior, which allows for great simplification of enforcement 
-and prioritization policies.
+restricted in behaviour, which allows for great simplification of enforcement 
+and prioritization policies. Admission control systems can give administrators 
+the benefit of tight control over the data center, mitigating the effects of 
+queue-buildup and bursts.
 
-A desirable solution is a global, centralized arbiter which is able to predict 
-and fairly distribute flows in the network before bursts or congestion even 
-occur. By treating the network's compute and forwarding power as a single 
-finite resource, a controller acts akin to a OS scheduler distributing CPU time 
-slices to processes. This design approach follows SDN's aspiration of 
-introducing operating systems abstractions to the networking domain space.
+Although it is unclear how predictable data center traffic truly is, various 
+works have shown that repeated patterns may 
+exist.\cite{msr_dc,fb_dc,traffic,microte}
+If data center traffic exhibits predictable patterns over long time, it may be 
+possible to devise a proactive mechanism which predicts the traffic matrix of 
+the next nth iterations, and adjusts traffic accordingly.
 
-In this project, we plan to explore the possibilities of a centralized, 
-proactive flow scheduler. We ask ourselves the following research questions:
+%Winstein and Balakrishnan,~2013\cite{remy} demonstrated that it is possible to 
+%learn optimal TCP congestion control parameters for various network 
+%topologies. 
+%Benson et al.,~2014\cite{microte} analyzed data center traffic and inferred 
+%that around 30\%
+
+An ideal culmination of these insights may be a global, 
+centralized arbiter, able to predict and fairly distribute flows in the 
+network. By only allowing a safe sending rate for each participant it minimizes 
+the amount of bursts and loss possible.
+Treating the network's compute and forwarding power as a single finite 
+resource, the controller acts akin to a OS scheduler distributing CPU time 
+slices to processes.
+
+This project report details our exploration of such a centralized, 
+proactive flow scheduler. We asked ourselves the following research questions:
 \begin{enumerate}
@@
 \end{enumerate}
-
-In the scope of this course, we attempt to answer these questions and design a 
-simple predictive scheduler in Mininet. If successful, we will benchmark our 
-results and evaluate the level of utilization compared to contemporary 
-scheduling systems.
+In the scope of this report, we outline the design and implementation of our 
+flow scheduler and learning approach. We perform an analysis of its 
+effectiveness compared to DCTCP and Hedera and conclude with a discussion of 
+the benefits and drawbacks of centralized learning and admission control in the 
+data center context.

### 2017-12-10 Documents/Report/intro.tex
stuff.

@@
 
-While TCP has not fundamentally changed in
+TCP's algorithm as best-effort solution to globally maximize utilization has 
+not fundamentally changed over the last thirty years. However, new research 
+advancement raise the question if TCP and its assumptions about networks are 
+still a viable option.~\cite{pcc,bbr,perc}
 
@@
 While idea of admission control and service guarantees in networks is not 
-new,~\cite{access_limit, access_limit2} these designs traditionally 
+new~\cite{access_limit, access_limit2}, these designs traditionally 
 aim to assure quality and bandwidth guarantees in a contentious, decentralized, 
@@
 exist.\cite{msr_dc,fb_dc,traffic,microte}
-If data center traffic exhibits predictable patterns over long time, it may be 
+If data center traffic exhibits predictable patterns, it may be 
 possible to devise a proactive mechanism which predicts the traffic matrix of 
-the next nth iterations, and adjusts traffic accordingly.
+next iterations, and adjusts traffic accordingly.
 
@@
 
-An ideal culmination of these insights may be a global, 
+An ideal culmination of these insights is a global, 
 centralized arbiter, able to predict and fairly distribute flows in the 

### 2017-12-10 Documents/Report/related_work.tex
stuff.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
\label{sec:related}
Network traffic congestion control and fairness has been an active topic of 
research for several decades and exhibits a vast range of work. We draw the 
majority of our inspiration from the following four categories.
\subsection{End-to-End TCP optimization}
\subsection{Centralized Scheduling and Congestion Control}
\subsection{Network Admission-Control}
\subsection{Datacenter Predictability and Learning}

### 2017-12-10 Iroko/plot_qlen.py
Queue length plotting script

@@ -1,0 +1,152 @@
+'''
+@author: Milad Sharif(msharif@stanford.edu)
+'''
+
+from monitor.helper import *
+from math import fsum
+import numpy as np
+
+parser = argparse.ArgumentParser()
+parser.add_argument('--input', '-f', dest='files', required=True, help='Input rates')
+parser.add_argument('--out', '-o', dest='out', default=None,
+                    help="Output png file for the plot.")
+
+args = parser.parse_args()
+
+
+''' Output of bwm-ng has the following format:
+    unix_timestamp;iface_name;bytes_out;bytes_in;bytes_total;packets_out;packets_in;packets_total;errors_out;errors_in
+    '''
+
+traffics = ['stag_prob_0_2_3_data', 'stag_prob_1_2_3_data', 'stag_prob_2_2_3_data',
+            'stag_prob_0_5_3_data', 'stag_prob_1_5_3_data', 'stag_prob_2_5_3_data', 'stride1_data',
+            'stride2_data', 'stride4_data', 'stride8_data', 'random0_data', 'random1_data', 'random2_data',
+            'random0_bij_data', 'random1_bij_data', 'random2_bij_data', 'random_2_flows_data',
+            'random_3_flows_data', 'random_4_flows_data', 'hotspot_one_to_one_data']
+
+labels = ['stag0(0.2,0.3)', 'stag1(0.2,0.3)', 'stag2(0.2,0.3)', 'stag0(0.5,0.3)',
+          'stag1(0.5,0.3)', 'stag2(0.5,0.3)', 'stride1', 'stride2',
+          'stride4', 'stride8', 'rand0', 'rand1', 'rand2', 'randbij0',
+          'randbij1', 'randbij2', 'randx2', 'randx3', 'randx4', 'hotspot']
+
+'''
+traffics=['stag_prob_0_2_3_data']
+
+labels=['stag0(0.2,0.3)']
+'''
+
+def get_qlen(input_file, pat_iface):
+    pat_iface = re.compile(pat_iface)
+
+    data = read_list(input_file)
+
+    rate = {}
+    column = 2
+    for row in data:
+        try:
+            ifname = row[0]
+        except:
+            break
+        if ifname not in ['eth0', 'lo']:
+            if not rate.has_key(ifname):
+                rate[ifname] = []
+            try:
[…]

### 2017-12-10 Documents/Report/implementation.tex
draft of implementation section, plus additional papers referenced

@@
 work.
+
+\subsection{Topology Simulation}
+We chose to use  Mininet to prototype Iroko.
+Mininet is a python emulation API which leverages the linux kernel to provide a lightweight virtualization of a data center in a single machine ~\cite{mininet}.
+This is an important advantage, because it allows us to get meaningful results with limited amounts of hardware.
+Other methods of testing Iroko would require more resources, or require specialized code specific to the testbed.
+Mininet avoids these problems and allowed us to develop a prototype which could be deployed into a real data center.
+
+Mininet’s lightweight virtualization also allows the API to scale to bigger topologies which otherwise couldn’t be prototyped easily ~\cite{mininet}.
+This feature will be useful for further future research, because we can then test our controller on other topologies beyond the ones we look in this current work.
+
+We created  fat-tree topology consisting of 20 switches and 16 hosts in the Mininet environment.
+The 20 switches are configured into pods as described in < cite paper about Fat-tree>.
+This design choice was to allow a direct comparison to Hedera  ~\cite{hedera}, another global arbiter which re-routes elephant flows using information from the data center.
+In their work they also used a fat-tree topology as we have described.
+This topology could be further expanded in future work to determine the scalability of our controller.
+
+\subsection{Data Collection }
+	Using a mapping from links to hosts, we were able to directly infer information from a host using data collected from the links in the second tier of our topology.
+We used the awk scripting language and the linux tc qdisc command to collect information from the links.
+Awk is a built in text parsing language available on most Linux distributions < cite?>.
+We used a script which would read the /proc/net/dev file to collect the currently used  bandwidth on a link.
+From this information, we calculate the free bandwidth as the difference between the theoretical maximum bi-section link bandwidth and the used bandwidth .
+As our goal was to predict the appropriate bandwidth allocatio to each host, we assumed it to be useful data to the controller to know the current bandwidth utilization by a host.
+Using tc -s qdisc show dev <interface> and a combination of regex expressions, we collected the packet loss, interface overlimits, and interface queues.
+Each of these data provide further information to hint at the current state of the data center.
+Our goal was to eliminate dropped packets, and would  use this measurement of our controller performance.
+Overlimit packets provides information on when a link goes over it’s allocated bandwidth, and adds a potentially useful feature for our controller to learn from.
+Interface queues provide the controller with information on the queue length of the host.
+These additional statistics we hoped would provide our controller enough data to make meaningful inferences of the appropriate bandwidth allocation to each host.
+\subsection{Bandwidth Allocation}
+We implmented an applications level set of code which connects with our python code base via Sockets.
+It works by sending a packet to hosts which allocated their allowed maximum bandwidth for an epoch.
+It reads a traffic matrix shared among all the hosts and sends traffic accordingly.
+For Iroko we are rate limiting the traffic at the application level.
+So we are using UDP instead of TCP.
+
+\subsection{Iroko}
+
+The core of Irokos learning capabilities is solved using a reinforcement learning algorithm.
+Reinforcement learning is a research field in machine learning which develops techniques which can solve the reinforcement learning problem.
+The goal in reinforcement learning is to learn what decisions to make that maximize the expected future reward from the environment based on it’s current state.
+An agent follows a policy, a function which maps to decisions the agent can make in the environment, which allows the agent to explore the environment.
+To adjusts it’s decision, a reward function is used to determine how valuable the current state is to the agent ~\cite{Sutton:1998:IRL:551283}.
+The reward is generally used to represent the goal the agent is trying to achieve ~\cite{Sutton:1998:IRL:551283}.
+
+We decided to map a “state” in our data center as the current host to allocate bandwidth, along with all the associated data we decribed in the previous section.
+This representation allowed us to use a single controller for all the hosts.
+The advantage of this is that our controller could learn on all hosts in a single model instead of separate models per host which would have looked at each host’s bandwidth allocation in isolation.
+The action of our agent was the percentage to either increase or decrease the current hosts allocated bandwidth.
+We chose to use a single action which could take values betwen -.9 and .9.
+We chose this representation, because we assumed that a host should alway have some allocated bandwidth.
+The current implementation also adjust bandwidth based on the current allocated bandwidth, and could thus forseeably elimnate a hots bandwidth entirely.
+We defined  a reward of 1 if the packet loss was 0 , and a reward of -1 for any loss greater than 0.
+A more intricate reward function can be devised, and a limitation of the current reward signal is that 0 loss is also obtainable if no bandwidth is used in the current epoch.
+
+We chose to use the deep deterministic policy gradient policy algorithm as described in ~\cite{DDPG} .
+DDPG is an actor-critic model which learns the appropriate policy by taking the gradient of the actor network relative to the critic Q-value approximation.
+Our agent learns using a replay memory implemented as a circular array.
+A replay memory buffer has shown to help deep learning algorithms learn better by sampling state, action, reward, transition state pair to learn from which helps speed learning ~\cite{DQLearning}.
+The motivation for a replay buffer is to train on data which is uncorrelated, which is not normally the case in reinforcement learning problems.
+This decision might be unnecessary for a data center where some authors have suggested that data center traffic is unpredictable and may suggest that each state can then be seen to be independent, and identifcally sampled from the environment.
+We recognize these findings may also give an upper limit to the expected succes of Iroko.
+
+To learn, a reinforcement learning agent must balance exploration and exploitation ~\cite{Sutton:1998:IRL:551283}.
+Instead of using the exploration policy suggested in ~\cite{DDPG} we add uniform distributed value between 0 – 1 which will increase the predicted action with a 33% probability, decrease it with 33% probability, or not alter the value at all.
+This means we will take the predicted value (.5 * 1 + .5 * .33) =  ~.666% or 2/3 of the time, and will increase 16.
+667 % of the time or decrease 16.667 % of the time.
+Alternative exploration policies could be used, but this one was chosen for it’s simplicity.
+
+Iroko then modifies the allocated bandwidth by taking the chosen action as follows:
+new bandwidth <-- current bandwidth + (suggested allocation ) * current bandwidth

### 2017-12-10 install.sh
Fix install

@@ -1,7 +1,8 @@
 #!/bin/bash
+sudo apt-get install mininet
 sudo -H pip install pip --upgrade
 sudo -H pip install --upgrade numpy 
-sudo -H pip install http://download.pytorch.org/whl/cu75/torch-0.2.0.post3-cp27-cp27mu-manylinux1_x86_64.whl C
+sudo -H pip install http://download.pytorch.org/whl/cu75/torch-0.2.0.post3-cp27-cp27mu-manylinux1_x86_64.whl
 sudo -H pip install torchvision
 sudo apt-get install bwm-ng
 make -C Iroko/cluster_loadgen

### 2017-12-11 Documents/Report/intro.tex
Push more text.

@@
 datacenter network traffic have shown that buffer overruns and small 
-short-lived bursts are the primary factor of loss.\cite{fb_dc,msr_dc}
+short-lived bursts are the primary factor of loss.\cite{fb_dc,msr_dc,dctcp}
 Despite the innovation potential of data centers, TCP has dominated as 
@@
 
-%Winstein and Balakrishnan,~2013\cite{remy} demonstrated that it is possible to 
-%learn optimal TCP congestion control parameters for various network 
-%topologies. 
-%Benson et al.,~2014\cite{microte} analyzed data center traffic and inferred 
-%that around 30\%
-
 An ideal culmination of these insights is a global, 

### 2017-12-11 Documents/Report/related_work.tex
Push more text.

@@
 research for several decades and exhibits a vast range of work. We draw the 
-majority of our inspiration from the following four categories.
+majority of our inspiration from the following four categories of research.
 \subsection{End-to-End TCP optimization}
-\subsection{Centralized Scheduling and Congestion Control}
+Many different TCP optimizations tailored to various environments and use-cases 
+exist. In general, these extensions share TCP's fundamental requirements of 
+scalability, decentralization, and fairness. Optimizations are generally 
+static, only on the end-host, and implemented "offline".~\cite{throwdown} This 
+family of modifications includes tweaks to the congestion window size and 
+growth, signal detection, and optimization goal(e.g., bandwidth, latency, or 
+convergence).~\cite{tcp_family}
+In this project we compare to two particular variants of the traditional TCP 
+protocol. Data Center TCP~\cite{dctcp}
+
+\subsection{Centralized Scheduling}
+SDN
 \subsection{Network Admission-Control}
-\subsection{Datacenter Predictability and Learning}
+\subsection{Predicting and Learning Network Traffic}
+Winstein and Balakrishnan,~2013\cite{remy} demonstrated that it is possible to 
+learn optimal TCP congestion control parameters for various network 
+topologies. 
+Benson et al.,~2014\cite{microte} analyzed data center traffic and inferred 
+that around 30\%

### 2017-12-11 Documents/Report/implementation.tex

--- 2017-12-09 Documents/Report/implementation.tex
@@
 work.
+
+\subsection{Topology Simulation}
+We chose to use  Mininet, a python data center emulation API, to prototype Iroko.
+Mininet's implementation leverages the Linux kernel to provide a lightweight virtualization of a data center in a single machine ~\cite{mininet}.
+This is an important advantage, because it allows us to create a prototype we can test using limited amounts of hardware.
+Other methods of testing Iroko would require more resources, or require specialized code specific to the testbed which cannot deployed in a real system ~\cite{mininet}.
+Mininet avoids these problems and allowed us to develop a prototype of Iroko which could be deployed into a real data center.
+
+Mininet’s lightweight virtualization also allows the API to scale better, and to support bigger topologies which otherwise couldn’t be prototyped easily ~\cite{mininet}.
+This feature will be useful for future research, because we can then test Iroko on other topologies beyond the ones we look use in this current work.
+
+We created  fat-tree topology consisting of 20 switches and 16 hosts in the Mininet environment.
+The 20 switches are configured into 5 pods as described in ~\cite{fattree}.
+This design choice was to allow a direct comparison to Hedera  ~\cite{hedera}, a global arbiter which uses flow statistics to re-route elephant flows.
+In their work they also used a fat-tree topology as we have described, and this allows us to directly evaluate against Hedera's performance .
+
+\subsection{Data Collection }
+We collected several different metrics from the links connected to each host.
+We leverage the fact we know which host connects to which link in the data center.
+We used the awk scripting language and the Linux tc qdisc command to collect information from the links.
+
+Awk is a text parsing language available on most Linux distributions by default.
+Our script read from each links.
+The awk script read the /proc/net/dev file to collect the currently used  bandwidth on a link.
+We then used this to calculate free bandwidth which is the difference between theoretical maximum bisection bandwidth for a link from the used bandwidth.
+
+In addition, we used $tc -s qdisc show dev <interface>$ where $<interface>$ is the current host we are polling from, and a combination of regex expressions to collect the packet loss, packet overlimits, and interface queues .
+Each of these data provide further information to hint at the current state of the data center.
+Our goal was to eliminate dropped packets, and use this to help Iroko make predictions.
+Overlimit packets provides information on when a link goes over it’s allocated bandwidth, and adds a potentially useful feature for our controller to learn from.
+Interface queues provide Iroko with information on the queue length of the host.
+These additional statistics we hoped would provide our controller enough data to make meaningful inferences of the appropriate bandwidth allocation to each host.
+
+\subsection{Bandwidth Allocation}
+Iroko rate limits traffic at the applications level, and thus uses UDP instead of TCP to communicate with hosts.
+It reads a traffic matrix shared among all the hosts and sends network traffic accordingly.
+Iroko also manages the current allocated bandwidth per host by sending a packet which updates their current allocated bandwidth.
+
+\subsection{Iroko}
+
+In order to make inferences about the appropriate bandwidth allocation per host, we treat bandwidth allocation as a reinforcement learning problem.
+The goal in reinforcement learning is to maximize the expected reward given by an environment for each possible state $s$ ~\cite{Sutton:1998:IRL:551283}.
+These values are often tracked using the state-action value function $Q(s,a)$, but often it is intractable to store all $(s,a)$) pairs a table, and often a function approximator is used instead to estimate $Q(s,a)$.
+Using function $Q(s,a)$, a policy $\pi(s)$ is then learned to direct how the agent should proceed in the environment.
+
+We represent a state from the data center as an array combining a bit vector corresponding to the current host and it's allocated bandwidth, free bandawidth, packet loss, packet overlimits, and queue length:
+
+\[ [0 0 0 ... 1 0 0, bandwidth, free bandwidth, packet loss, overlimit, queue ] \]
+The advantage of this is that Iroko will learn a policy which considers all hosts together in a representation instead of using separate agents for each network which wouldn't consider bandwidth allocation to each separate host.
+
+The action Iroko takes at each step is the amount to increase or decrease the current hosts allocated bandwidth.
+Iroko will generate an adjustment value clamped between $-.9 - .9$.
+We chose this value range, because we wanted to avoid starving hosts.
+Iroko then modifies the allocated bandwidth similar to the approach used in ~\cite{remy} for congestion windows by using the action with allocated bandwidth as follows:
+
+\[bandwidth_{t+1}   <-- bandwidth_t +  a_t * bandwidth_t\]
+
+where a is the chosen bandwidth adjustment, t is the previous epoch, and t+1 is the new allocation.
+This representation could be  modified in a number of ways, such as forcing a minimum expected bandwidth, or add the capacity to starve a host.
+
+For a reinforcement learning agent to learn a policy, a reward function is used to generate the value of a particular state, and is used to represent the goal of an agent ~\cite{Sutton:1998:IRL:551283}.
+We defined  a reward of 1 if the packet loss was 0 , and a reward of -1 for any loss greater than 0.
+M  A more intricate reward functions can be devised, and onea limitation of the current reward signal is that 0 loss can occuris also obtainable if no bandwidth is used in the current epoch which could confuse the agent.
+
+We chose to use the deep deterministic policy gradient policy algorithm (DDPG) as described in ~\cite{DDPG} .
+DDPG is an actor-critic model which uses neural networks to approximate $Q(s,a)$, and learns the policy by taking the actor network gradients relative to the critic $Q(s,a)$ approximations ~\cite{DDPG}.
+Our agent learns using a replay memory, which is a technique that helps deep reinforcement learning algorithms learn by making the data uncorrelated via random batch sampling ~\cite{DQLearning}.
+This decision might be unnecessary for a data center where some authors have suggested that data center traffic is inherently unpredictable ~\cite{microte} and may suggest that each state can be viewed as uncorrelated inherently.
+We recognize these findings may also give an upper limit to the expected performance of Iroko.
+
+To find the optimal values, an agent must balance exploiting the current best actions with exploring an environment with a new action which may lead to bigger future rewards ~\cite{Sutton:1998:IRL:551283}.
+Instead of using the exploration approach used in ~\cite{DDPG}, we use $\epsilon-greedy\ learning$, where an agent takes the best action with probability $\epsilon$ and a random action with probability $1 - \epsilon$.
+Our random action is then defined as follows:
+
+\[
+noise(x) = 
+\begin{cases}
+
+uniform(0,1), &\text{P(x = increase) = .333},\\
+uniform(-1,0), &\text{P(x = decrease) = .3333},\\
+0, & \text{(x = unchanged) = .3333}
+\end{cases}
+\]
+
+where uniform(a,b) refers to taking a random value in the range between [a, b]
+This means we will take the best predicted value $(.5 * 1 + .5 * .33) =  ~.66.6\%$  and will increase $P(x= increase) * P(\epsilon) = 16.
+667\%$ of the time or decrease $P(x= decrease) * P(\epsilon) = 16.
+667\%$ of the time.
+It should be noted that the action value is still clamped between -0.9 and 0.9 regardless of the noise introduced.
+Alternative exploration policies could be used, but this one was chosen due it’s simplicity, and theoretical capability to explore the whole action space if allowed to run indefinitely.

### 2017-12-11 Documents/Report/implementation.tex
implementation changes

@@
 \section{Implementation}
-\label{sec:implementation}
-We intend to emulate our system in Mininet~\cite{mininet} to observe traffic 
-patterns and infer a suitable token algorithm. Mininet has proven itself to be 
+\subsection{Overview}
+
+We emulate our system in Mininet~\cite{mininet} to observe traffic 
+patterns and to appropriately rate-limit end-hosts. Mininet has proven itself to be 
 a viable tool to model new congestion control 
-algorithms~\cite{mininet_learning}, and will help us prototype our concept 
-efficiently. We will build a custom SDN controller that interacts with 
-traditional OpenFlow software switches as well as end-hosts. End-hosts will run 
+algorithms~\cite{mininet_learning}, and helped us prototype our concept 
+efficiently. We built our custom SDN controller Iroko which interacts with 
+traditional OpenFlow software switches as well as end-hosts. End-hosts run 
 a custom real-world traffic generation script which adjusts based on 
 information packets sent by the controller.
-If time permits, we may expand our implementation to MaxiNet, which can emulate 
-large-scale network stress tests on multiple physical hosts.
-
-We initially considered a second implementation alternative in C/C++ based on 
-the FastPass~\cite{fastpass} source code. This would provide us with a fully 
-deployable system which we could fork our implementation from. A major 
-advantage of this approach is the ability to test scenarios and traffic 
-algorithms using real software code.
-However, several concerns made us favour a Mininet emulation instead.
-Firstly, FastPass relies on DPDK integration, which requires actual hardware 
-interfaces. The central arbiter in the FastPass design would need to run on a 
-dedicated machine, which increases prototyping and development complexity 
-substantially.
-
-Secondly, the FastPass code is highly specialized and optimized research code 
-with only little available documentation. Modifying and evolving the source 
-code will require thorough understanding of kernel and networking development, 
-a significant time-sink. For a class project, these may be major initial 
-hurdles, taking away from the research aspect of the design concept.
-Consequently, we have decided to pursue an approach which allows us to quickly 
-develop an understanding of the problem without being obstructed by engineering 
-work.
 
@@
 Mininet's implementation leverages the Linux kernel to provide a lightweight virtualization of a data center in a single machine ~\cite{mininet}.
-This is an important advantage, because it allows us to create a prototype we can test using limited amounts of hardware.
+This is an important advantage, because it allowed us to create a prototype we can test using limited amounts of hardware.
 Other methods of testing Iroko would require more resources, or require specialized code specific to the testbed which cannot deployed in a real system ~\cite{mininet}.
-Mininet avoids these problems and allowed us to develop a prototype of Iroko which could be deployed into a real data center.
+In contrast, Mininet has allowed us to develop Iroko to be deployable in a real system without re-implementation independent of testbed specific code.
 
 Mininet’s lightweight virtualization also allows the API to scale better, and to support bigger topologies which otherwise couldn’t be prototyped easily ~\cite{mininet}.
-This feature will be useful for future research, because we can then test Iroko on other topologies beyond the ones we look use in this current work.
+This feature will be useful for future research, because we can then test Iroko on other topologies beyond the fat-tree topology we looked at in this work.
 
-We created  fat-tree topology consisting of 20 switches and 16 hosts in the Mininet environment.
+In the Mininet environment, we created a fat-tree topology consisting of 20 switches and 16 hosts .
 The 20 switches are configured into 5 pods as described in ~\cite{fattree}.
 This design choice was to allow a direct comparison to Hedera  ~\cite{hedera}, a global arbiter which uses flow statistics to re-route elephant flows.
-In their work they also used a fat-tree topology as we have described, and this allows us to directly evaluate against Hedera's performance .
+In ~\cite{hedera}, Hedera was evaluted  using a fat-tree topology as we have described, and this allows us to directly compare Iroko against Hedera's performance .
 
@@
 Awk is a text parsing language available on most Linux distributions by default.
-Our script read from each links.
-The awk script read the /proc/net/dev file to collect the currently used  bandwidth on a link.
-We then used this to calculate free bandwidth which is the difference between theoretical maximum bisection bandwidth for a link from the used bandwidth.
+The awk script reads the /proc/net/dev file to collect the currently used  bandwidth on a link.
+We then used this to calculate free bandwidth which is the difference between theoretical maximum bisection bandwidth, and the reported bandwidth on a link.
 
-In addition, we used $tc -s qdisc show dev <interface>$ where $<interface>$ is the current host we are polling from, and a combination of regex expressions to collect the packet loss, packet overlimits, and interface queues .
+In addition, we used the following Linux command:
+\[$tc -s qdisc show dev <interface>$ \] 
+where $<interface>$ is the current host we are polling from.
+Using the obtained text, we then applied regex expressions to collect the packet loss, packet overlimits, and interface queues .
 Each of these data provide further information to hint at the current state of the data center.
-Our goal was to eliminate dropped packets, and use this to help Iroko make predictions.
+Our goal was to eliminate dropped packets, and used this to help Iroko make rate-limiting decisions.
 Overlimit packets provides information on when a link goes over it’s allocated bandwidth, and adds a potentially useful feature for our controller to learn from.
@@
 
-In order to make inferences about the appropriate bandwidth allocation per host, we treat bandwidth allocation as a reinforcement learning problem.
+In order to make rate-limiting decisions per host, we treat bandwidth allocation as a reinforcement learning problem.
 The goal in reinforcement learning is to maximize the expected reward given by an environment for each possible state $s$ ~\cite{Sutton:1998:IRL:551283}.
-These values are often tracked using the state-action value function $Q(s,a)$, but often it is intractable to store all $(s,a)$) pairs a table, and often a function approximator is used instead to estimate $Q(s,a)$.
-Using function $Q(s,a)$, a policy $\pi(s)$ is then learned to direct how the agent should proceed in the environment.
+These values are often tracked using the state-action value function $Q(s,a)$, and often this value is approximated using a function due to intractable memory needs ~\cite{Sutton:1998:IRL:551283}.
+Using $Q(s,a)$, a policy $\pi(s)$ is then learned to direct how the agent should proceed in the environment.$\pi(s)$ role is to decide the next action a reinforcement learning agent should take, in our case the amount to rate-limit a host.
 
-We represent a state from the data center as an array combining a bit vector corresponding to the current host and it's allocated bandwidth, free bandawidth, packet loss, packet overlimits, and queue length:
+We represent a state from the data center as an array combining a bit vector corresponding to the current host and it's allocated bandwidth, free bandwidth, packet loss, packet overlimits, and queue length:
 
-\[ [0 0 0 ... 1 0 0, bandwidth, free bandwidth, packet loss, overlimit, queue ] \]
-The advantage of this is that Iroko will learn a policy which considers all hosts together in a representation instead of using separate agents for each network which wouldn't consider bandwidth allocation to each separate host.
+\[ [0 0 0 ... 1 0 0, bandwidth, free bandwidth, packet loss, overlimit, queue length ] \]
+The advantage of this is that Iroko will learn a policy which considers all hosts together in unified representation instead of using separate agents for each network which wouldn't consider bandwidth allocation to each separate host.
 

### 2017-12-11 Iroko/LearningAgentv2.py
working on queues hopefully

@@ -15,7 +15,7 @@ class LearningAgentv2:
         self.lam = lam
         self.defaultmax = defaultmax
         self.initmax = initMax
-        self.controller = DDPG(gamma,memory,5 + 16, 1,tau=.25, criticpath='critic', actorpath='actor') 
+        self.controller = DDPG(gamma,memory,5 + 16, 1,tau=.25)#), criticpath='critic', actorpath='actor') 
         
 
 

### 2017-12-11 Iroko/iroko_controller.py

@@ -285,7 +285,8 @@ if __name__ == '__main__':
             Agent.updateHostsBandwidth(interface, free_bandwidths[interface], drops[interface] )
             # A supposedly more eloquent way of doing it
             reward = 0
-            if(drops[interface]  > 0.0):
+            #if(drops[interface]  > 0.0):
+            if(queues[interface]):
                 reward = -1
             else:
                 reward = 1

### 2017-12-11 Documents/Report/related_work.tex

--- 2017-12-10 Documents/Report/related_work.tex
@@
 research for several decades and exhibits a vast range of work. We draw the 
-majority of our inspiration from the following four categories.
+majority of our inspiration from the following four categories of research.
+
 \subsection{End-to-End TCP optimization}
-\subsection{Centralized Scheduling and Congestion Control}
+Many different TCP optimizations tailored to various environments and use-cases 
+exist. In general, these extensions share TCP's fundamental requirements of 
+scalability, decentralization, and fairness. Optimizations are generally 
+static, only on the end-host, and implemented "offline".~\cite{throwdown} This 
+family of modifications includes tweaks to the congestion window size and 
+growth, signal detection, and optimization goal(e.g., bandwidth, latency, or 
+convergence).~\cite{tcp_family}
+
+In this project, we compare Iroko to two particular TCP protocols of interest.  
+Data Center TCP~\cite{dctcp}, a congestion control algorithm specialized to 
+reduce bufferbloat in data center networks. DCTCP leverages Explicit Congestion 
+Notifications (ECNs) as early warning signal to identify queue buildup and 
+adjusts its sending rate to the amount of ECN-marked packets on the receiving 
+side. As part of the static TCP congestion control family, DCTCP is fully 
+decentralized, does not limit host sending rates in advance and relies on a 
+static congestion control model.
+
+The second algorithm we compare to is TCP CUBIC~\cite{cubic}, the default 
+congestion control protocol of the Linux kernel. CUBIC uses a cubic window size 
+function and serves as the baseline for Iroko.
+
+\subsection{Centralized Scheduling}
+Hedera~\cite{hedera} is an early example of a centralized scheduler working in 
+synergy with an Equal-Cost-Multipath (ECMP) routing environment. Hedera 
+monitors switch traffic and identifies long-running elephant flows which may 
+collide with short-lived, bursty traffic. Hedera overrides the ECMP routing 
+entry and moves the elephant flow to an underutilized link, thus reducing the 
+potential for flow collision and congestion.
+
+MicroTE~\cite{microte} is another central scheduler working under the 
+assumption that a subset of short-term traffic is predictable. The controller 
+isolates flows which it determined to be predictable and install a custom 
+optimal routing entry. The remaining unpredictable traffic is routed according 
+to ECMP striping.
+
+Hedera and MicroTE rely on TCP to converge to optimal sending rates and do not 
+perform any global sending rate optimizations.
+Iroko as a rate-limiting arbiter is not intended to compete with centralized 
+flow schedulers and may in fact benefit from building on existing routing 
+frameworks.
+
 \subsection{Network Admission-Control}
-\subsection{Datacenter Predictability and Learning}
+Recent contributions have shown that purely admission-based approaches 
+may work just as well as, if not better than, the conventional TCP fairness 
+principle. A large focus of these systems is to eliminate queuing and fast flow 
+convergence coupled with minimal round trip latency.
+FastPass~\cite{fastpass} employs an arbiter mechanism, in which hosts request 
+access to the network resources. For any new flow, the FastPass shim layer will 
+send a "admission request" to a central controller, which chooses an optimal 
+allocated time slot and path. Once an applications has been assigned this path, 
+it is guaranteed to experience zero queuing latency and jitter during 
+transmission. This "zero-loss" network provides application with a stable and 
+reliable transmission experience. However, FastPass still works reactively and 
+only operates in incoming requests. Queueing is thus pushed to end-hosts and 
+the arbiter itself.
+
+Another recent example of admission control is the 
+ExpressPass~\cite{expresspass} system. As opposed to FastPass, ExpressPass is 
+fully distributed with the explicit intention to scale. ExpressPass uses a 
+token system enforced by switches in the network. Receiving hosts provide 
+senders with tokens as currency, allowing them to send traffic. If the network 
+experiences congestion, switches drop the tokens given by the receiver, thus 
+limiting the amount of packets a host can send. In ExpressPass, congestion 
+control is managed by the network elements instead of end-hosts.
+
+Iroko builds heavily on the contributions of both FastPass and ExpressPass. 
+However, we intend to go further and schedule rate-limits ahead of time in 
+order to reduce sending delays and host starvation of admission-control systems 
+substantially.
+
+\subsection{Predicting and Learning Network Traffic}
+A new direction of networking research is study on the predictability and 
+trainability of network traffic.~\cite{throwdown, learning_tcp}
+A common assumption is that data center traffic is largely unpredictable and 
+untrainable.~\cite{fb_dc}
+However, Winstein and Balakrishnan,~2013\cite{remy} demonstrated that it is 
+possible to learn near-optimal TCP congestion control parameters for various 
+network topologies. The system, Remy, represents an offline learning algorithm 
+inferring optimal TCP parameters based on pre-supplied topology and 
+traffic-model.
+While Remy only performed offline training, PCC~\cite{pcc,throwdown} aims to 
+develop a dynamic and actively improving congestion control algorithm. PCC 
+makes no prior assumption about the network and continuously reiterates over 
+its host-local optimization function to find optimal parameters.
+Furthermore, Benson et al.,~2014\cite{microte} analyzed data center traffic and 
+inferred that around 30\% of short-lived traffic is predictable over a period 
+of 1-2 seconds.
+
+In this work, we assume that traffic may be trainable to a degree that is 
+useful. We conjecture on the basis of the aforementioned work that it is 
+possible to identify recurring patterns in a subset of traffic and adjust flow 
+and congestion control pre-emptively.

### 2017-12-11 Documents/Report/eval.tex
Evaluation: first draft

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Evaluation}
-\label{sec:evaluation}
+\section{Evaluation} \label{sec:evaluation}
 
-In the absence of a commercial data center, the implementation of our network 
-design is going to be done on top of mininet with simulated FatTree topologies 
-of varying size. We stress test with iPerf and simulate data center traffic 
-using tcpreplay and packet traces provided by~\cite{traffic} .
+This section describes the evaluation of Iroko using our testbed built on top of
+Mininet simulator. The main goal is to compare the performance against the
+existing centralized and decentralized techniques like Hedera~\cite{hedera} and
+DCTCP~\cite{dctcp} respectively.
 
-To evaluate the general effectiveness of our system, we plan to measure against 
-existing centralized as well as decentralized solutions.
-The centralized design will be based on Hedera~\cite{hedera}, a common and 
-influential datacenter scheduler. The decentralized congestion control 
-mechanism will be DCTCP~\cite{dctcp}, a state-of-the-art TCP congestion 
-algorithm.
+\subsection{Benchmark suite}
+
+For evaluation purposes we are targeting a small scale data center of 16 hosts
+and 20 switches with FatTree topology. We took measurements for Hedera, DCTCP,
+ECMP and with Non-blocking switch on the same test setup. ECMP measurement is
+considered as the base line. Non-blocking is a hypothetical single switch fully
+connected setup and is considered as the theoretical maximum we are aiming to
+achieve.
+
+For traffic simulation, we are using the same traffic matrices used in
+Hedera~\cite{hedera}. We took measurements for the variations of the following 3
+traffic patterns:
+
+\begin{enumerate} \item Stride$\left(i\right)$: A host with index $x$ sends to
+the host with index $(x + i)mod(num hosts)$.  \item Staggered Prob $\left(EdgeP,
+PodP\right)$: A host sends to another host in the same edge switch with
+probability $EdgeP$, and to its same pod with probability $PodP$, and to the
+rest of the network with probability $1-EdgeP - PodP$.  \item Random: A host
+sends to any other host in the	network with uniform probability. We include
+bijective mappings and ones where hotspots are present.  \end{enumerate}
 
 Since we are primarily concerned about reducing the latency and packet drops
-while keeping utilisation at maximum, the measurements to get a good insight 
-into how well the design can perform are as follows:
+while keeping utilization at maximum, we measured bisection bandwidth and
+    average queue length on switches. The experiments last for 60 seconds for
+    each benchmark and we consider only the middle 40 seconds.
+
+\subsection{Bisection Bandwidth}
+
+\begin{figure*} 
+	\centering 
+	\includegraphics[width=1\linewidth]{rate.png} 
+	\caption{}
+	\label{fig:rate} 
+\end{figure*}
+
+In Figure \ref{fig:rate} we show the aggregate bisection bandwidth. Iroko
+consistently provides better utilization compared to ECMP and DCTCP in almost
+all experiments. As expected non-blocking is providing the best performance out
+of all four techniques. Utilization of non-blocking is lower for traffic
+patterns with high incast. Hedera perform extremely well in all experiments
+except for some where Iroko outperforms Hedera. Interestingly enough, Hedera
+even outperforms non-blocking limit in stride8 experiment. ECMP does reasonably
+well when there is high probability of local communication for example smaller
+strides, but start to deteriorate as the stride length increases because of hash
+collisions.
+
+Another interesting observation is that DCTCP gives abysmal utilization for some
+experiments. Further investigation into this revealed that this is because of
+DCTCP's aggressive congestion window size reduction strategy on incast. The
+proposed solution for incast problem is to use dynamic buffer allocation in
+switches which is not implemented in our test setup. General performance of
+DCTCP is not satisfactory either. For some experiments it even under performs
+ECMP.  According to DCTCP paper it is optimized for some specific data center
+traffic patterns like web search, soft real-time applications, recommendation
+systems etc which we don't include in our experiments.
+
+\subsection{More Benchmarks}
+
+Apart from utilization and loss rate following are the other benchmarks that we
+are interested. We have not measured these yet due to time concerns.
+
 \begin{enumerate}
 
-\item Latency: We are aiming for a low latency network which means that 99th 
-percentile latency in the network across all flows should as low as possible. 
-Latency should be measured when all hosts are sending packets at the maximum 
-limit and also with random traffic patterns. There should be low latency even 
-during sudden bursts as the transmission rate is limited by a base value.
+\item \textbf{99th percentile latency:} Since we are aiming for a low latency
+which means that 99th percentile latency in the network across all flows should
+be as low as possible.
 
-\item Packet drop rate: Ideally this metric will approach zero, as the 
-objective of Iroko is to minimize packet loss.
+\item \textbf{Fairness:} Fairness across flows can be tested by introducing a
+new host to a completely saturated network and increasing the transmission rate
+to see if all the hosts gets fair share of the total bandwidth. Currently we are
+assuming all flows should be at equal priority. Differential priority is out of
+scope.
 
-\item Fairness: Fairness can be tested by introducing a new host to a 
-completely saturated network and increasing the transmission rate to see if all 
-the hosts gets fair share of the total bandwidth. We are assuming all flows 
-should be at equal priority. Differential priority is out of scope.
+\item \textbf{Responsiveness:} This metric evaluates the predictive power and
+efficiency of Iroko. It would be interesting to analyze the response time of the
+central scheduler compared to decentralized DCTCP and TCP.
 
-\item Responsiveness: This metric depends on the predictive power and 
-efficiency of Iroko. It may be interesting to analyze the response time of the 
-central scheduler compared to decentralized DCTCP and TCP. In addition, it is 
-valuable to identify at what data center and flow size the central computing 
-node of the network may become a bottleneck.
-
-\item Network utilization: We will measure the used bandwidth and the total 
-load on the hosts. These metrics can be measured by varying the load from each 
-host. Load to utilization ratio should be 1 until the saturation point and 
-after that utilization should stay stable at the maximum bisection bandwidth.
-It is also important to check that there is no starvation happening in any 
-host. This can be measured by simulating random traffic patterns and plotting 
-residual bandwidth and excess load. We expect overall utilization to be lower 
-compared to Hedera and DCTCP. Although Iroko will maximize goodput, we do not 
-measure it on grounds of simplicity.
+\item \textbf{Starvation:} Since we are rate limiting at the host level, Iroko
+should make sure that no flows are starving to send data. This can be measured
+by plotting residual bandwidth and excess load.
 
 \end{enumerate}
-
-As baseline we will also compare to the default TCP congestion algorithm to 
-estimate how our scheduler fares against the default case. As our initial 
-simple design is very constrained in its ability to route and control traffic 
-we expect to achieve less overall optimality against DCTCP and Hedera, but 
-still gain a substantial advantage over common TCP.

### 2017-12-11 Iroko/monitor/plot_queue.py
Use CDF for plot_qlen.py

@@ -69,17 +69,25 @@ to_plot = []
 
 def get_style(i):
     if i == 0:
+        return {'color': 'brown'}
+    elif i == 1:
         return {'color': 'red'}
+    elif i == 2:
+        return {'color': 'magenta'}
+    elif i == 3:
+        return {'color': 'green'}
+    elif i == 4:
+        return {'color': 'royalblue'}
     else:
         return {'color': 'black', 'ls': '-.'}
 
 
 for i, f in enumerate(args.files):
     data = read_list(f)
-    xaxis = map(float, col(0, data))
+    xaxis = map(float, col(1, data))
     start_time = xaxis[0]
     xaxis = map(lambda x: x - start_time, xaxis)
-    qlens = map(float, col(1, data))
+    qlens = map(float, col(2, data))
     if args.summarise or args.cdf:
         to_plot.append(qlens[10:-10])
     else:
@@ -93,6 +101,8 @@ plt.grid(True)
 #plt.yticks(yaxis, ylabels)
 # plt.ylim((0,1100))
 plt.ylim((args.miny, args.maxy))
+fig = plt.figure(1)
+fig.set_size_inches(8.5, 6.5)
 
 if args.summarise:
     plt.xlabel("Link Rates")
@@ -117,8 +127,8 @@ elif args.cdf:
         plt.plot(xs, ys, label=args.legend[i], lw=2, **get_style(i))
         plt.ylabel("Fraction")
         plt.xlabel("Packets")
-        plt.ylim((0, 1.0))
-        plt.legend(args.legend, loc="upper left")
+        plt.ylim((0.5, 1.0))
+        plt.legend(args.legend, loc="best")
         plt.title("")
 else:
     plt.xlabel("Seconds")

### 2017-12-11 Iroko/plot_qlen.py

@@ -5,11 +5,10 @@
 from monitor.helper import *
 from math import fsum
 import numpy as np
+from os import system
 
 parser = argparse.ArgumentParser()
 parser.add_argument('--input', '-f', dest='files', required=True, help='Input rates')
-parser.add_argument('--out', '-o', dest='out', default=None,
-                    help="Output png file for the plot.")
 
 args = parser.parse_args()
 
@@ -35,118 +34,17 @@ traffics=['stag_prob_0_2_3_data']
 labels=['stag0(0.2,0.3)']
 '''
 
-def get_qlen(input_file, pat_iface):
-    pat_iface = re.compile(pat_iface)
-
-    data = read_list(input_file)
-
-    rate = {}
-    column = 2
-    for row in data:
-        try:
-            ifname = row[0]
-        except:
-            break
-        if ifname not in ['eth0', 'lo']:
-            if not rate.has_key(ifname):
-                rate[ifname] = []
-            try:
-                rate[ifname].append(float(row[column]) / 100)
-            except:
-                break
-    vals = []
-    for k in rate.keys():
-        if pat_iface.match(k):
-            avg_qlen = avg(rate[k][10:-10])
-            vals.append(avg_qlen)
-    return max(vals)
-
-
 def plot_results(args):
 
-    fbb = 16. * 10  # 160 mbps
-
-    num_plot = 2
-    num_t = 20
-    n_t = num_t / num_plot
-
-    bb = {'nonblocking': [], 'hedera': [], 'iroko': [], 'dctcp': [], 'ecmp': []}
-
     # sw = '4h1h1'
-    sw = '1001'
-    for t in traffics:
-        print("Nonblocking:", t)
-        input_file = args.files + '/nonblocking/%s/qlen.txt' % t
-        vals = get_qlen(input_file, sw)
-        print(vals)
-        bb['nonblocking'].append(vals / fbb)
+    for i,t in enumerate(traffics):
+        nb_input = args.files + '/nonblocking/%s/qlen.txt' % t
+        ecmp_input = args.files + '/fattree-ecmp/%s/qlen.txt' % t
+        dctcp_input = args.files + '/fattree-dctcp/%s/qlen.txt' % t
+        iroko_input = args.files + '/fattree-iroko/%s/qlen.txt' % t
[…]

### 2017-12-11 Iroko/LearningAgentv2.py
uses saved files now

@@ -15,7 +15,7 @@ class LearningAgentv2:
         self.lam = lam
         self.defaultmax = defaultmax
         self.initmax = initMax
-        self.controller = DDPG(gamma,memory,5 + 16, 1,tau=.25)#), criticpath='critic', actorpath='actor') 
+        self.controller = DDPG(gamma,memory,5 + 16, 1,tau=.25, criticpath='critic', actorpath='actor') 
         
 
 

### 2017-12-11 Documents/Report/implementation.tex
Implementation updates

@@
 Mininet's implementation leverages the Linux kernel to provide a lightweight virtualization of a data center in a single machine ~\cite{mininet}.
-This is an important advantage, because it allowed us to create a prototype we can test using limited amounts of hardware.
-Other methods of testing Iroko would require more resources, or require specialized code specific to the testbed which cannot deployed in a real system ~\cite{mininet}.
-In contrast, Mininet has allowed us to develop Iroko to be deployable in a real system without re-implementation independent of testbed specific code.
-
-Mininet’s lightweight virtualization also allows the API to scale better, and to support bigger topologies which otherwise couldn’t be prototyped easily ~\cite{mininet}.
-This feature will be useful for future research, because we can then test Iroko on other topologies beyond the fat-tree topology we looked at in this work.
+This allowed us to create a prototype of Iroko we can test quickly using limited amounts of hardware.
+Other methods of testing Iroko would require more resources, or specialized code specific to the testbed which cannot deployed in a real system ~\cite{mininet}.
+In contrast, Mininet has allowed us to develop Iroko such that we could deploy it in a real system without re-implementation independent of testbed specific code.
 
@@
 This design choice was to allow a direct comparison to Hedera  ~\cite{hedera}, a global arbiter which uses flow statistics to re-route elephant flows.
-In ~\cite{hedera}, Hedera was evaluted  using a fat-tree topology as we have described, and this allows us to directly compare Iroko against Hedera's performance .
+In ~\cite{hedera}, Hedera was evaluated  using a fat-tree topology as we have described, and this allows us to directly compare Iroko against Hedera's performance .
+
+Mininet’s lightweight virtualization also allows the API to scale such as to support bigger topologies which otherwise couldn’t be prototyped easily ~\cite{mininet}.
+This feature will be useful for future research, because we can then test Iroko on other topologies beyond the fat-tree topology we looked at in this work.
 
 \subsection{Data Collection }
-We collected several different metrics from the links connected to each host.
-We leverage the fact we know which host connects to which link in the data center.
-We used the awk scripting language and the Linux tc qdisc command to collect information from the links.
+Since we knew prior to testing which port connected to each host, we used different link statistics from each switch port connected to host to make predictions.
+This was accomplished using the awk scripting language and the Linux tc qdisc command to collect information.
 
 Awk is a text parsing language available on most Linux distributions by default.
-The awk script reads the /proc/net/dev file to collect the currently used  bandwidth on a link.
+Our awk script reads the /proc/net/dev file on each interface to collect the currently used  bandwidth on a link.
 We then used this to calculate free bandwidth which is the difference between theoretical maximum bisection bandwidth, and the reported bandwidth on a link.
@@
 where $<interface>$ is the current host we are polling from.
-Using the obtained text, we then applied regex expressions to collect the packet loss, packet overlimits, and interface queues .
+Using the obtained text, we then applied regex expressions to collect the packet loss, packet overlimits, and interface queues.
 Each of these data provide further information to hint at the current state of the data center.
@@
 Overlimit packets provides information on when a link goes over it’s allocated bandwidth, and adds a potentially useful feature for our controller to learn from.
-Interface queues provide Iroko with information on the queue length of the host.
+Interface queues provide Iroko with information on the queue length of the host, which measures how many packets are buffered to be processed.
 These additional statistics we hoped would provide our controller enough data to make meaningful inferences of the appropriate bandwidth allocation to each host.
@@
 In order to make rate-limiting decisions per host, we treat bandwidth allocation as a reinforcement learning problem.
-The goal in reinforcement learning is to maximize the expected reward given by an environment for each possible state $s$ ~\cite{Sutton:1998:IRL:551283}.
-These values are often tracked using the state-action value function $Q(s,a)$, and often this value is approximated using a function due to intractable memory needs ~\cite{Sutton:1998:IRL:551283}.
+The goal in reinforcement learning is to maximize the expected reward given by an environment for each possible state $s$ by choosing the best action $a$ to take at each time step~\cite{Sutton:1998:IRL:551283}.
+These values are can be tracked using the state-action value function $Q(s,a)$, and often this value is approximated using a function due to intractable memory needs ~\cite{Sutton:1998:IRL:551283}.
 Using $Q(s,a)$, a policy $\pi(s)$ is then learned to direct how the agent should proceed in the environment.$\pi(s)$ role is to decide the next action a reinforcement learning agent should take, in our case the amount to rate-limit a host.
@@
 Iroko will generate an adjustment value clamped between $-.9 - .9$.
-We chose this value range, because we wanted to avoid starving hosts.
+We chose this value range, because we wanted to avoid host starvation.
 Iroko then modifies the allocated bandwidth similar to the approach used in ~\cite{remy} for congestion windows by using the action with allocated bandwidth as follows:

### 2017-12-11 Iroko/topo_non_block.py
More graphs.

@@ -32,7 +32,7 @@ import sys
 parentdir = os.path.dirname(os.path.dirname(os.path.abspath(__file__)))
 sys.path.insert(0, parentdir)
 #import iperf_peers
-MAX_QUEUE = 1000
+MAX_QUEUE = 50
 
 
 class NonBlocking(Topo):

### 2017-12-11 Documents/Report/implementation.tex
Intro changes.

@@
 
-In addition, we used the following Linux command:
-\[$tc -s qdisc show dev <interface>$ \] 
-where $<interface>$ is the current host we are polling from.
-Using the obtained text, we then applied regex expressions to collect the packet loss, packet overlimits, and interface queues.
-Each of these data provide further information to hint at the current state of the data center.
-Our goal was to eliminate dropped packets, and used this to help Iroko make rate-limiting decisions.
-Overlimit packets provides information on when a link goes over it’s allocated bandwidth, and adds a potentially useful feature for our controller to learn from.
-Interface queues provide Iroko with information on the queue length of the host, which measures how many packets are buffered to be processed.
-These additional statistics we hoped would provide our controller enough data to make meaningful inferences of the appropriate bandwidth allocation to each host.
+In addition, we used the following Linux command:\\
+\texttt{tc -s qdisc show dev <interface>}\\
+where \texttt{<interface>} is the current host we are polling from. Using the 
+obtained text, we then applied regex expressions to collect the packet loss, 
+packet overlimits, and interface queues. Each of these data provide further 
+information to hint at the current state of the data center. Our goal was to 
+eliminate dropped packets, and used this to help Iroko make rate-limiting 
+decisions. Overlimit packets provides information on when a link goes over it’s 
+allocated bandwidth, and adds a potentially useful feature for our controller 
+to learn from. Interface queues provide Iroko with information on the queue 
+length of the host, which measures how many packets are buffered to be 
+processed.  These additional statistics we hoped would provide our controller 
+enough data to make meaningful inferences of the appropriate bandwidth 
+allocation to each host.
 

### 2017-12-11 Documents/Report/implementation.tex
changes to design and implementation

--- 2017-12-11 Documents/Report/implementation.tex
@@
 This allowed us to create a prototype of Iroko we can test quickly using limited amounts of hardware.
-Other methods of testing Iroko would require more resources, or specialized code specific to the testbed which cannot deployed in a real system ~\cite{mininet}.
+Other methods of testing Iroko would require more resources, or specialized code specific to the testbed which cannot be deployed in a real system ~\cite{mininet}.
 In contrast, Mininet has allowed us to develop Iroko such that we could deploy it in a real system without re-implementation independent of testbed specific code.
@@
 \subsection{Data Collection }
-Since we knew prior to testing which port connected to each host, we used different link statistics from each switch port connected to host to make predictions.
+Since we knew prior to testing which port connected to each host, we collected different data from port connections to each end-host.
 This was accomplished using the awk scripting language and the Linux tc qdisc command to collect information.
@@
 In addition, we used the following Linux command:
-\[$tc -s qdisc show dev <interface>$ \] 
+texttt{tc\ -s\ qdisc\ show\ dev\ <interface>}
 where $<interface>$ is the current host we are polling from.
@@
 
-\subsection{Iroko}
+\subsection{Reinforcement Learning Algorithm}
 
-In order to make rate-limiting decisions per host, we treat bandwidth allocation as a reinforcement learning problem.
-The goal in reinforcement learning is to maximize the expected reward given by an environment for each possible state $s$ by choosing the best action $a$ to take at each time step~\cite{Sutton:1998:IRL:551283}.
-These values are can be tracked using the state-action value function $Q(s,a)$, and often this value is approximated using a function due to intractable memory needs ~\cite{Sutton:1998:IRL:551283}.
-Using $Q(s,a)$, a policy $\pi(s)$ is then learned to direct how the agent should proceed in the environment.$\pi(s)$ role is to decide the next action a reinforcement learning agent should take, in our case the amount to rate-limit a host.
-
-We represent a state from the data center as an array combining a bit vector corresponding to the current host and it's allocated bandwidth, free bandwidth, packet loss, packet overlimits, and queue length:
-
-\[ [0 0 0 ... 1 0 0, bandwidth, free bandwidth, packet loss, overlimit, queue length ] \]
-The advantage of this is that Iroko will learn a policy which considers all hosts together in unified representation instead of using separate agents for each network which wouldn't consider bandwidth allocation to each separate host.
-
-The action Iroko takes at each step is the amount to increase or decrease the current hosts allocated bandwidth.
-Iroko will generate an adjustment value clamped between $-.9 - .9$.
-We chose this value range, because we wanted to avoid host starvation.
-Iroko then modifies the allocated bandwidth similar to the approach used in ~\cite{remy} for congestion windows by using the action with allocated bandwidth as follows:
-
-\[bandwidth_{t+1}   <-- bandwidth_t +  a_t * bandwidth_t\]
-
-where a is the chosen bandwidth adjustment, t is the previous epoch, and t+1 is the new allocation.
-This representation could be  modified in a number of ways, such as forcing a minimum expected bandwidth, or add the capacity to starve a host.
-
-For a reinforcement learning agent to learn a policy, a reward function is used to generate the value of a particular state, and is used to represent the goal of an agent ~\cite{Sutton:1998:IRL:551283}.
-We defined  a reward of 1 if the packet loss was 0 , and a reward of -1 for any loss greater than 0.
-M  A more intricate reward functions can be devised, and onea limitation of the current reward signal is that 0 loss can occuris also obtainable if no bandwidth is used in the current epoch which could confuse the agent.
-
-We chose to use the deep deterministic policy gradient policy algorithm (DDPG) as described in ~\cite{DDPG} .
-DDPG is an actor-critic model which uses neural networks to approximate $Q(s,a)$, and learns the policy by taking the actor network gradients relative to the critic $Q(s,a)$ approximations ~\cite{DDPG}.
-Our agent learns using a replay memory, which is a technique that helps deep reinforcement learning algorithms learn by making the data uncorrelated via random batch sampling ~\cite{DQLearning}.
-This decision might be unnecessary for a data center where some authors have suggested that data center traffic is inherently unpredictable ~\cite{microte} and may suggest that each state can be viewed as uncorrelated inherently.
+We chose to use the deep deterministic policy gradient policy algorithm (DDPG) almost entirely as described in ~\cite{DDPG} .
+DDPG is a form of actor-critic reinforcement learning algorithm ~\cite{Sutton:1998:IRL:551283} which uses neural networks to approximate $Q(s,a)$, and learns the policy by taking the actor network gradients relative to the critic $Q(s,a)$ approximations ~\cite{DDPG}.
+Our agent learns using replay memory, which is a technique that helps deep reinforcement learning algorithms learn by making the data uncorrelated via random batch sampling ~\cite{DQLearning}.
+This decision might be unnecessary for a data center where authors have suggested that data center traffic is inherently unpredictable ~\cite{microte} and this suggests that we could assume data center data is inherently uncorrelated.
 We recognize these findings may also give an upper limit to the expected performance of Iroko.
@@
 To find the optimal values, an agent must balance exploiting the current best actions with exploring an environment with a new action which may lead to bigger future rewards ~\cite{Sutton:1998:IRL:551283}.
-Instead of using the exploration approach used in ~\cite{DDPG}, we use $\epsilon-greedy\ learning$, where an agent takes the best action with probability $\epsilon$ and a random action with probability $1 - \epsilon$.
+Instead of using the exploration approach used in ~\cite{DDPG}, we use $\epsilon-greedy\ learning$, where an agent takes the best action with probability $\epsilon$ and a random action with probability $1 - \epsilon$ ~\cite{Sutton:1998:IRL:551283}.
 Our random action is then defined as follows:
@@
 Alternative exploration policies could be used, but this one was chosen due it’s simplicity, and theoretical capability to explore the whole action space if allowed to run indefinitely.
+
+We represent a state from the data center as an array combining a bit vector corresponding to the current host and it's allocated bandwidth, free bandwidth, packet loss, packet overlimits, and queue length:
+
+\[ [0 0 0 ... 1 ... 0 0, bandwidth, free bandwidth, packet loss, overlimit, queue length ] \]
+The advantage of this is that Iroko will learn a policy which considers all hosts together in unified representation instead of using separate agents for each network which wouldn't consider bandwidth allocation to each separate host.
+
+The action Iroko takes at each step is the amount to increase or decrease the current hosts allocated bandwidth.
+Iroko will generate an adjustment value clamped between $-.9 - .9$.
+We chose this value range, because we wanted to avoid host starvation.
+Iroko then modifies the allocated bandwidth similar to the approach used in ~\cite{remy} for congestion windows by using the action with allocated bandwidth as follows:
+
+\[bandwidth_{t+1}   \leftarrow bandwidth_t +  a_t * bandwidth_t\]
+
+where a is the chosen bandwidth adjustment, t is the previous epoch, and t+1 is the new allocation.
+This representation could be  modified in a number of ways, such as forcing a minimum expected bandwidth, or add the capacity to starve a host.

### 2017-12-11 Documents/Report/intro.tex

@@
 is a perennial challenge for any network operator.
-
 Inefficient routing and congestion control can quickly lead to 
@@
 short-lived bursts are the primary factor of loss.\cite{fb_dc,msr_dc,dctcp}
+
 Despite the innovation potential of data centers, TCP has dominated as 
 the congestion control algorithm of choice.
-TCP is a \textit{reactive protocol}, responding to indicators of packet drops
-latency. However, this surge of loss and latency in the network already 
-portends a problem. Overflowing queues in forwarding elements or mismatched 
-hardware capabilities imply that traffic has not been optimally managed.
+Still, TCP is a \textit{reactive protocol}, only responding to signals of 
+congestion. A substantial and noticeable increase of loss and latency in the 
+network already portends a problem. Overflowing queues in forwarding elements 
+or mismatched hardware capabilities imply that traffic has not been optimally 
+managed.
 Ideally, a network should always be "zero-queue", i.e., latency 
@@
 
-TCP's algorithm as best-effort solution to globally maximize utilization has 
-not fundamentally changed over the last thirty years. However, new research 
-advancement raise the question if TCP and its assumptions about networks are 
-still a viable option.~\cite{pcc,bbr,perc}
+TCP's algorithm is a best-effort solution to globally maximize utilization 
+under the fairness principle. While this design has not fundamentally changed 
+over the last thirty years, new research advancements raise the question if TCP 
+is still the most reliable option.~\cite{pcc,bbr,perc}
 
@@
 restricted in behaviour, which allows for great simplification of enforcement 
-and prioritization policies. Admission control systems can give administrators 
-the benefit of tight control over the data center, mitigating the effects of 
+and prioritization policies. Admission-based systems can give administrators 
+the advantage of tight control over the data center, mitigating the effects of 
 queue-buildup and bursts.
@@
 If data center traffic exhibits predictable patterns, it may be 
-possible to devise a proactive mechanism which predicts the traffic matrix of 
+possible to devise a proactive mechanism which forecasts the traffic matrix of 
 next iterations, and adjusts traffic accordingly.

### 2017-12-11 Documents/Report/related_work.tex

@@
 exist. In general, these extensions share TCP's fundamental requirements of 
-scalability, decentralization, and fairness. Optimizations are generally 
-static, only on the end-host, and implemented "offline".~\cite{throwdown} This 
-family of modifications includes tweaks to the congestion window size and 
-growth, signal detection, and optimization goal(e.g., bandwidth, latency, or 
-convergence).~\cite{tcp_family}
+scalability, decentralization, and fairness. Optimizations are generally only 
+implemented on the end-host and the model itself is static and 
+"offline".~\cite{throwdown} This family of modifications includes tweaks to the 
+congestion window size and growth, signal detection, and optimization goal 
+(e.g., bandwidth, latency, or convergence).~\cite{tcp_family}
 
 In this project, we compare Iroko to two particular TCP protocols of interest.  
-Data Center TCP~\cite{dctcp}, a congestion control algorithm specialized to 
-reduce bufferbloat in data center networks. DCTCP leverages Explicit Congestion 
-Notifications (ECNs) as early warning signal to identify queue buildup and 
-adjusts its sending rate to the amount of ECN-marked packets on the receiving 
-side. As part of the static TCP congestion control family, DCTCP is fully 
-decentralized, does not limit host sending rates in advance and relies on a 
-static congestion control model.
+The first, Data Center TCP (DCTCP)~\cite{dctcp}, is a congestion control 
+algorithm developed to reduce bufferbloat in data center networks. DCTCP 
+leverages Explicit Congestion Notifications (ECNs) as early warning signal to 
+identify queue buildup and adjusts its sending rate to the amount of ECN-marked 
+packets on the receiving side. As part of the static TCP congestion control 
+family, DCTCP is fully decentralized, does not limit host sending rates in 
+advance and relies on a static congestion control model.
 
 The second algorithm we compare to is TCP CUBIC~\cite{cubic}, the default 
-congestion control protocol of the Linux kernel. CUBIC uses a cubic window size 
-function and serves as the baseline for Iroko.
+congestion control protocol of the Linux kernel. CUBIC uses a cubic function   
+to increase window size and serves as the baseline for Iroko. It is intended to 
+work well over high bandwidth, high latency networks.
 
@@
 
-MicroTE~\cite{microte} is another central scheduler working under the 
+MicroTE~\cite{microte} is a central scheduler working under the 
 assumption that a subset of short-term traffic is predictable. The controller 
@@
 Recent contributions have shown that purely admission-based approaches 
-may work just as well as, if not better than, the conventional TCP fairness 
-principle. A large focus of these systems is to eliminate queuing and fast flow 
+may work as well as, if not better than, conventional unbounded TCP. A large 
+focus of these systems is to eliminate queuing and fast flow 
 convergence coupled with minimal round trip latency.
+
 FastPass~\cite{fastpass} employs an arbiter mechanism, in which hosts request 
@@
 traffic-model.
-While Remy only performed offline training, PCC~\cite{pcc,throwdown} aims to 
-develop a dynamic and actively improving congestion control algorithm. PCC 
-makes no prior assumption about the network and continuously reiterates over 
+While Remy only performed static offline training, PCC~\cite{pcc,throwdown} 
+aims to develop a dynamic and actively improving congestion control algorithm. 
+PCC makes no prior assumption about the network and continuously reiterates over
 its host-local optimization function to find optimal parameters.
-Furthermore, Benson et al.,~2014\cite{microte} analyzed data center traffic and 
+
+As part of the contribution of MicroTE, Benson et 
+al.,~2014~\cite{microte} analyzed data center traffic and 
 inferred that around 30\% of short-lived traffic is predictable over a period 
@@
 In this work, we assume that traffic may be trainable to a degree that is 
-useful. We conjecture on the basis of the aforementioned work that it is 
+useful. On the basis of the aforementioned work we conjecture that it may be 
 possible to identify recurring patterns in a subset of traffic and adjust flow 

### 2017-12-11 Documents/Report/design.tex

@@
 \subsection{Overview and Goals}
-In our initial simple system, "Iroko", a centralized controller regulates all 
+In our initial system, "Iroko", a centralized controller regulates all 
 node traffic by rate-limiting end-hosts. We have opted for a centralized 
@@
 
-Iroko guarantees minimal congestion and low average access latency. Reducing 
+Iroko attempts to provide minimal congestion and low average access latency. Reducing 
 network-global packet loss and jitter is the priority and objective function of 
@@
 an ideal Iroko system, packet-loss will only rarely, if ever, occur. In respect 
-to the free lunch theorem we aim to trade off maximum bandwidth utilization for 
+to the no free lunch theorem we aim to trade off maximum bandwidth utilization for 
 optimal system stability and reliable latency.
@@
 It is important to note that routing decisions are not in the current scope of 
-Iroko. While it may certainly beneficial to include dynamic and adaptive 
+Iroko. While it may certainly be beneficial to include dynamic and adaptive 
 routing decisions in a predictive scheduler, we do not include these features 
@@
 
+\subsection{Adapting traffic flow}
+
+The controller will dynamically adjust bandwidth allocations to ensure that hosts 
+requiring more bandwidth are allocated it, while host that are not currently 
+utilizing all their allocated bandwidth have their allocated bandwidth reduced. 
+To avoid starving hosts, some small amount of bandwidth must always be 
+allocated to a given host even if it not utilizing network resources. Thus, the 
+network, by design, will never reach 100\% utilization.
+
+Conversely however one should rarely see dropped packets due to network 
+congestion. Furthermore, the controller must be able to react quickly to 
+changing bandwidth needs for hosts. Ideally, using a predictive method.
+
+\subsection{Reinforcement Learning}
+In order to make predictive rate-limiting decisions per host, we treat rate-limiting as a reinforcement learning problem.
+In reinforcement learning, an agent (in our case the arbiter) performs actions
+in the environment and receives a reward signal from the environment to adjust 
+its decisions in the next epoch ~\cite{Sutton:1998:IRL:551283}. The reward is 
+generally used to represent the goal an agent is trying to achieve. The agent 
+can freely choose whatever actions are necessary to achieve this goal 
+~\cite{Sutton:1998:IRL:551283}.
+
+The environment in this case is defined as the statistics collected from each 
+switch and the current end-host to rate limit.
+This decision is based on the fact that these data are easy to collect and provides a current snapshot of the performance of the data center.
+Using this representation, we take full advantage of all the information provided in our data center.
+
+We wish to optimize over the packet loss, and define the reward to be good (+1) if there is no loss and bad (-1) if there is any loss.
+One limitation of the current reward signal is it's inability to distinguish cases when loss is 0 because no bandwidth was used in the epoch.
+This means, a more intricate reward functions can be devised for  a future work.
+
+Given the data center assumption, the number of end-hosts is known in the 
+topology. We define our arbiter's actions as a single continuous value which will either increase 
+decrease, or maintain the allocated bandwidth to different end hosts.
+A continuous value allows our agent to make direct rate-limiting decisions, and reinforcement learning algorithms have been devised to learn continuous value actions such as this ~\cite{DDPG}.
+
+Thus we frame our problem as one which can now be solved using techniques in reinforcement learning. Given full representation of our environment and 
+actions, we will store the state-action value, also referred to as the Q(s,a) function,  in an artificial neural network. The Q(s,a) function 
+stores the reward signal information that is used to make future action choices.
+We chose an artificial neural network to approximate this function, because of they are theoretically proven function approximators ~\cite{Cybenko1989} which have made intractable environmental representations manageable~\cite{Sutton:1998:IRL:551283}.
+
 \subsection{Controlling traffic flow}
-Rate-limiting is performed on the granularity of the IP protocol, allowing us 
+Our future research goal for Iroko is to rate-limit on the granularity of the IP protocol, allowing us 
 to restrict traffic to specific hot regions and to package flows into groups.
-End-hosts are guaranteed a limited amount of bandwidth which they may send to 
-each destination address. The traffic window size for any flow is calculated 
+End-hosts will be guaranteed a limited amount of bandwidth which they may send to 
+each destination address. The traffic window size for any flow will be calculated 
 based on the host's current traffic availability. Total bandwidth of a group of 
@@
 
-The central arbiter enforces these bandwidth and access limitations by 
-assigning "tokens" to nodes. These tokens are plain information packets 
+Iroko would enforce these bandwidth and access limitations by assigning "tokens" to nodes. These tokens are plain information packets 
 specifying the affected destination address flow, its bandwidth guarantee, and 
@@
 
-\subsection{Adapting traffic flow}
-
-Initially, the controller will compute optimal route configuration based on the 
-topology and link bandwidth using a simple heuristic bin-packing approach. 
-End-hosts will be initialized with a fixed low-to-medium bandwidth guarantee 
-that will be adjusted over time. This bandwidth guarantee will be below the 
-host's proportional bi-section bandwidth that would be used to reach any other 
-host. That is, if every end host transfers using the full bandwidth initially 
-allocated to it,  there would be no dropped packets due to congestion.
-
-Of course, this would vastly underutilize network resources, thus the 
-controller will dynamically adjust these allocations to ensure that hosts that 
-require the bandwidth are allocated it, while host that are not currently 
-utilizing all their allocated bandwidth have their allocated bandwidth reduced. 
-To avoid starving hosts, some small amount of bandwidth must always be 
-allocated to a given host even if it not utilizing network resources. Thus, the 
-network, by design, will never reach 100\% utilization.
-
-Conversely however one should rarely see dropped packets due to network 
-congestion. Furthermore, the controller must be able to react quickly to 
-changing bandwidth needs for hosts. Ideally, using a predictive method.
-
 \begin{figure}
@@
 
-\subsection{Statistical System}
-One potential implementation for our arbiter is to use reinforcement learning. 
-In reinforcement learning, an agent (in our case the arbiter) performs actions 
-in the environment and receives a reward signal from the environment to adjust 
-its decisions in the next epoch ~\cite{Sutton:1998:IRL:551283}. The reward is 
-generally used to represent the goal the agent is trying to achieve. The agent 
-can freely choose whatever actions are necessary to achieve this goal 
-~\cite{Sutton:1998:IRL:551283}.
-
-The environment in this case is defined as the statistics collected from each 
-switch. This decision is based on the fact that this data is easy to collect 
-and provides a current snapshot of the performance of the network. We wish to 
-optimize over the packet loss, and define the reward to be good if packet loss 
-is decreasing (+1 reward); bad if it is increasing ( -1 reward); and acceptable 
-if packet loss remains the same within a threshold (0 reward). Using this 
-representation, we take full advantage of all the information provided in our 
-data center; This representation is replicable and can be applied beyond the 
-scope of this project.
-
-Given the data center assumption, the number of end-hosts is known in the 
-topology. We can define our arbiter's actions as a discretized representation 
-among an n-dimensional vector where each cell is one of three values: increase 
-decrease, or maintain the allocated bandwidth. This representation offers 
-sufficient complexity. There are $3^n$ possible actions to perform and to 
-achieve optimal results the agent must explore the environment extensively, 
-precluding a significantly larger action space.
-
-A potential improvement to our representation is a hierarchical learning agent. 
-In this case, our arbiter would manage agents for each host and these 
-sub-agents would manage their respective host whether to increase, decrease, or 
-maintain bandwidth for their particular host. This problem can be referred to 
-as hierarchical reinforcement learning~\cite{Borga1993}, but may prove to be 
-beyond the scope of our current proposed solution, although has the added 
-benefit of solving a set of sub-problems to address the global problem of 
-packet loss and simplifies the representation spaces in the sub-problems.
-
-Thus we frame our problem as one which can now be solved using techniques in 
-reinforcement learning. A classic algorithm to consider is the use of 
-SARSA~\cite{Sutton:1998:IRL:551283}, which is an online temporal difference 
-algorithm used in classical reinforcement learning problems. One advantage of 
-SARSA is its online nature~\cite{Sutton:1998:IRL:551283} which promotes more 
-conservative decisions. Given full representation of our environment and 
-actions, we can store information in an artificial neural network which is 
-another popular choice in reinforcement learning literature. This model would 
-store the reward signal information that is used to make future action choices, 
-and is a theoretically proven function approximator which has made intractable 
-environmental representations manageable~\cite{Sutton:1998:IRL:551283}.
-
 %Using this model, we hope to achieve near 0% packet loss and provide a 

### 2017-12-11 Documents/Report/implementation.tex

@@
 
-In addition, we used the following Linux command:
-texttt{tc\ -s\ qdisc\ show\ dev\ <interface>}
-where $<interface>$ is the current host we are polling from.
-Using the obtained text, we then applied regex expressions to collect the packet loss, packet overlimits, and interface queues.
-Each of these data provide further information to hint at the current state of the data center.
-Our goal was to eliminate dropped packets, and used this to help Iroko make rate-limiting decisions.
-Overlimit packets provides information on when a link goes over it’s allocated bandwidth, and adds a potentially useful feature for our controller to learn from.
-Interface queues provide Iroko with information on the queue length of the host, which measures how many packets are buffered to be processed.
-These additional statistics we hoped would provide our controller enough data to make meaningful inferences of the appropriate bandwidth allocation to each host.
+In addition, we used the following Linux command:\\
+\texttt{tc -s qdisc show dev <interface>}\\
+where \texttt{<interface>} is the current host we are polling from. Using the 
+obtained text, we then applied regex expressions to collect the packet loss, 
+packet overlimits, and interface queues. Each of these data provide further 
+information to hint at the current state of the data center. Our goal was to 
+eliminate dropped packets, and used this to help Iroko make rate-limiting 
+decisions. Overlimit packets provides information on when a link goes over it’s 
+allocated bandwidth, and adds a potentially useful feature for our controller 
+to learn from. Interface queues provide Iroko with information on the queue 
+length of the host, which measures how many packets are buffered to be 
+processed.  These additional statistics we hoped would provide our controller 
+enough data to make meaningful inferences of the appropriate bandwidth 
+allocation to each host.
 

### 2017-12-11 Documents/Report/implementation.tex

@@
 To find the optimal values, an agent must balance exploiting the current best actions with exploring an environment with a new action which may lead to bigger future rewards ~\cite{Sutton:1998:IRL:551283}.
-Instead of using the exploration approach used in ~\cite{DDPG}, we use $\epsilon-greedy\ learning$, where an agent takes the best action with probability $\epsilon$ and a random action with probability $1 - \epsilon$ ~\cite{Sutton:1998:IRL:551283}.
+Instead of the exploration approach used in ~\cite{DDPG}, we use $\epsilon-greedy\ learning$, where an agent takes the best action with probability $\epsilon$ and a random action with probability $1 - \epsilon$ ~\cite{Sutton:1998:IRL:551283}.
 Our random action is then defined as follows:
@@
 \[
-noise(x) = 
+noise = 
 \begin{cases}

### 2017-12-11 Documents/Report/eval.tex
Evaluation: second draft

@@
 	\includegraphics[width=1\linewidth]{rate.png}
-	\caption{}
+	\caption{Average bisection bandwidth}
 	\label{fig:rate}
@@
 patterns with high incast. Hedera perform extremely well in all experiments
-except for some where Iroko outperforms Hedera. Interestingly enough, Hedera
-even outperforms non-blocking limit in stride8 experiment. ECMP does reasonably
-well when there is high probability of local communication for example smaller
-strides, but start to deteriorate as the stride length increases because of hash
-collisions.
+except for some like stride1, stag2$(0.5,0.3)$ where Iroko outperforms Hedera.
+Interestingly enough, Hedera even outperforms non-blocking limit in stride8
+experiment. ECMP does reasonably well when there is high probability of local
+communication for example smaller strides, but start to deteriorate as the
+stride length increases because of hash collisions.
 
 Another interesting observation is that DCTCP gives abysmal utilization for some
-experiments. Further investigation into this revealed that this is because of
-DCTCP's aggressive congestion window size reduction strategy on incast. The
-proposed solution for incast problem is to use dynamic buffer allocation in
-switches which is not implemented in our test setup. General performance of
-DCTCP is not satisfactory either. For some experiments it even under performs
-ECMP.  According to DCTCP paper it is optimized for some specific data center
-traffic patterns like web search, soft real-time applications, recommendation
-systems etc which we don't include in our experiments.
+experiments, for example randx2 and randx4. Further investigation into this
+revealed that this is because of DCTCP's aggressive congestion window size
+reduction strategy on incast. The proposed solution for incast problem is to use
+dynamic buffer allocation in switches which is not implemented in our test
+setup. General performance of DCTCP is not satisfactory either. For some
+experiments it even under performs ECMP.  According to DCTCP paper it is
+optimized for some specific data center traffic patterns like web search, soft
+real-time applications, recommendation systems etc which we don't include in our
+experiments.
+
+\subsection{Drop rate}
+
+\begin{figure*}
+	\centering
+	\includegraphics[width=1\linewidth]{qlen.png}
+	\caption{CDF of queue lengths}
+	\label{fig:qlen}
+\end{figure*}
+
+To analyze the drop rate we measured the number of packets built up in the queue
+on all switches for the same set of traffic patterns and plotted a cumulative
+distribution graph (Figure \ref{fig:qlen}). A high queue size is considered to
+be bad. We have fixed the maximum queue size to 50 packets.
+
+Iroko is not doing a good job here by consistently keeping $90\%$ of the queues
+empty and rest $10\%$ full. But it is interesting to note that Iroko performs
+really well for traffic patterns like stride1, stag2$(0.5,0.3)$ etc where it
+shows high utilization as well. DCTCP maintains the queues to the minimum across
+all experiments. This is not surprising considering the low utilization of
+DCTCP. ECMP and Hedera behaves almost the same. Nonblocking has the highest
+queue size out of all techniques. We believe that the CPU is the bottleneck here
+since it only uses a single switch to connect the entire network.
+
+Iroko definitely need more fine tuning for the prediction algorithm. We also
+need to investigate why the graph has a sudden jump from $0$ to $50$. Initial
+analysis shows that a small fixed set of switches are the bottleneck in all
+experiments. Our guess is that this happens because of the unfair bandwidth
+allocation across traffic flows. Analyzing fairness might provide better insight
+into this problem.
 

### 2017-12-12 Documents/Report/intro.tex
Minor fixes

@@
 posing novel design challenges and opportunities. Driven by minimization of 
-costs and maximization of compute power, data centers must run at the highest 
+costs and maximization of compute power, they must run at the highest 
 possible utilization to achieve the ideal compute/cost ratio.
-Optimizing the distribution of traffic and simultaneously guaranteeing fairness 
-is a perennial challenge for any network operator.
+
+Optimizing the distribution of traffic and simultaneously guaranteeing fairness is a perennial challenge for any network operator.
 Inefficient routing and congestion control can quickly lead to 
@@
 TCP's algorithm is a best-effort solution to globally maximize utilization 
-under the fairness principle. While this design has not fundamentally changed 
-over the last thirty years, new research advancements raise the question if TCP 
-is still the most reliable option.~\cite{pcc,bbr,perc}
+under the fairness principle. This design has not fundamentally changed 
+over the last thirty years, and new research advancements raise the question if TCP can still be considered to be the most sensible option.~\cite{pcc,bbr,perc}
 

### 2017-12-12 Iroko/plot_qlen.py
Plot qlen for selected traffic patterns

@@ -17,16 +17,16 @@ args = parser.parse_args()
     unix_timestamp;iface_name;bytes_out;bytes_in;bytes_total;packets_out;packets_in;packets_total;errors_out;errors_in
     '''
 
-traffics = ['stag_prob_0_2_3_data', 'stag_prob_1_2_3_data', 'stag_prob_2_2_3_data',
-            'stag_prob_0_5_3_data', 'stag_prob_1_5_3_data', 'stag_prob_2_5_3_data', 'stride1_data',
-            'stride2_data', 'stride4_data', 'stride8_data', 'random0_data', 'random1_data', 'random2_data',
-            'random0_bij_data', 'random1_bij_data', 'random2_bij_data', 'random_2_flows_data',
-            'random_3_flows_data', 'random_4_flows_data', 'hotspot_one_to_one_data']
+traffics = ['stag_prob_2_2_3_data',
+            'stag_prob_2_5_3_data', 'stride1_data',
+            'stride2_data', 'random0_data',
+            'random_2_flows_data',
+            'random_3_flows_data', 'random_4_flows_data']
 
-labels = ['stag0(0.2,0.3)', 'stag1(0.2,0.3)', 'stag2(0.2,0.3)', 'stag0(0.5,0.3)',
-          'stag1(0.5,0.3)', 'stag2(0.5,0.3)', 'stride1', 'stride2',
-          'stride4', 'stride8', 'rand0', 'rand1', 'rand2', 'randbij0',
-          'randbij1', 'randbij2', 'randx2', 'randx3', 'randx4', 'hotspot']
+labels = ['stag2(0.2,0.3)',
+          'stag2(0.5,0.3)', 'stride1', 'stride2',
+          'rand0',
+          'randx2', 'randx3', 'randx4']
 
 '''
 traffics=['stag_prob_0_2_3_data']

### 2017-12-12 Iroko/plot_qlen.py
Adjust plot_qlen legend

@@ -93,7 +93,7 @@ def plot_results(args):
         plt.xlabel(labels[i])
         plot_queue([dctcp_input, ecmp_input, iroko_input, hedera_input, nb_input],
                    ["dctcp", "ecmp", "iroko", "hedera", "nonblocking"], t)
-    plt.legend(bbox_to_anchor=(1., 1.22, 1., 1.))
+    plt.legend(bbox_to_anchor=(1.5, 1.22), loc='upper right', fontsize='x-large')
     plt.savefig("qlen.png")
     plt.show()
 

### 2017-12-12 Iroko/plot_qlen.py
Resize plot_qlen

@@ -85,12 +85,12 @@ def plot_results(args):
         hedera_input = args.files + '/fattree-hedera/%s/qlen.txt' % t
 
         fig = plt.figure(1)
-        fig.set_size_inches(36, 12)
+        fig.set_size_inches(40, 12)
         ax = fig.add_subplot(2, len(traffics) / 2, i + 1)
         ax.yaxis.grid()
         plt.ylim((0.8, 1.0))
-        plt.ylabel("Fraction")
-        plt.xlabel(labels[i])
+        plt.ylabel("Fraction", fontsize=16)
+        plt.xlabel(labels[i], fontsize=18)
         plot_queue([dctcp_input, ecmp_input, iroko_input, hedera_input, nb_input],
                    ["dctcp", "ecmp", "iroko", "hedera", "nonblocking"], t)
     plt.legend(bbox_to_anchor=(1.5, 1.22), loc='upper right', fontsize='x-large')

### 2017-12-12 Documents/Report/eval.tex
Modify eval.tex to use selected qlen plot

@@
 on all switches for the same set of traffic patterns and plotted a cumulative
-distribution graph (Figure \ref{fig:qlen}). A high queue size is considered to
-be bad. We have fixed the maximum queue size to 50 packets.
+distribution graph. In Figure \ref{fig:qlen} we are plotting some selected
+graph. A high queue size is considered to be bad. We have fixed the maximum
+queue size to 50 packets.
 

### 2017-12-12 Documents/Report/implementation.tex
Some modification to Bandwidth allocation in Implementation

@@
 Iroko rate limits traffic at the applications level, and thus uses UDP instead of TCP to communicate with hosts.
-It reads a traffic matrix shared among all the hosts and sends network traffic accordingly.
-Iroko also manages the current allocated bandwidth per host by sending a packet which updates their current allocated bandwidth.
+Traffic is generated using a custom load generation application we wrote which reads a traffic matrix shared among all the hosts and sends network traffic accordingly.
+Iroko manages the current allocated bandwidth per host by sending a controller packet which updates their current allocated bandwidth.
 

### 2017-12-12 Documents/Report/discussion.tex
Cleaned up a few things, added to discussion and future work

@@
 
-Rather predictably, Iroko struggles on random traffic patterns, however we
-do not see this as particularly concerning. Firstly \textit{Iroko} still outperforms
-both DCTCP and ECMP implementations. Further, as these traffic patterns
+Not unexpectedly, Iroko struggles on a few random traffic patterns, however we
+do not see this as particularly concerning. Firstly Iroko still outperforms
+both DCTCP and ECMP implementations on most random traffic patterns, and in general.
+Further, as these traffic patterns
 were designed by the authors of the Hedera paper \cite{hedera}, it remains
 likely these patterns are somewhat amenable to the Hedera implementation.
-Finally random traffic patterns seem unlikely to be prevalent in a
+Finally as discussed, fully random traffic patterns seem unlikely to be prevalent in a
 datacenter context. While the general predictability of datacenter
@@
 in our test network. A heavier handed approach may yield better
-results, however these may come at the cost of scalability.
+results, however these may come at the cost of scalability. We
+also noted that further tuning of the current implementation is
+likely to lead to stronger results. Due to testing limitations, 
+the controller had a very short number of periods to learn 
+on. In longer running cases, we full expect Iroko to 
+deliver stronger results as there is more time for machine learning 
+algorithms to converge.
+
+For efficient routing, we used EMCP with Iroko, however betting
+routing regimes exist, in particular Hedera. While we compared
+Iroko to Hedera, in fact the two systems are largely orthogonal
+and may benefit, from one another. Iroko does not directly deal
+with routing or so called elephant flows. Thus Iroko could be
+futher improved by employing a routing scheme similar to that of
+Hedera, while still retaining bandwidth prediction capabilities.

### 2017-12-12 Documents/Report/future_work.tex
Cleaned up a few things, added to discussion and future work

@@
 machine learning in their toolkit. However machine learning alone may not
-be sufficient to achieve the best possible results. While we  
-compared our results to Hedera, the work remains largely
-orthogonal to our approach, and scheduling flows would likely 
-complement machine learning based bandwidth control.
+be sufficient to achieve the best possible results. Machine learning
+based central traffic controllers can benefit from dynamic flow
+scheduling as well as further traffic engineering. Thus expanding
+the Iroko framework to include flow rather than interface based 
+predictions, or a combination of both, would be an interesting 
+avenue to pursue. Combined with a robust routing algorithm, we
+believe such a solution show superior performance when
+compared to virtual all currently existing solutions.
 
@@
 network resources at the termination such as MAP-REDUCE
-workloads, host may have very predictable completion times
+workloads, hosts may have very predictable completion times
 for these jobs. In such cases, hosts could report these 
@@
 of epochs, and therefore reduce efficiency. Such a
-system would be complementary to \textit{Iroko} and
-allow for finer grained control.
+system would be complementary to Iroko and
+allow for finer grained control. By taking further
+host level input, Iroko could reason about individual 
+flows, and allocate not only bandwidth, but also 
+specific paths.
 
@@
 central controller. This would allow hosts more ability to manage
-their own bandwidth with intertemporal substitution of flows,
+their own bandwidth with inter-temporal substitution of flows,
 while still enforcing centralized predictive control.

### 2017-12-12 Documents/Report/implementation.tex

@@
 
-We represent a state from the data center as an array combining a bit vector corresponding to the current host and it's allocated bandwidth, free bandwidth, packet loss, packet overlimits, and queue length:
+We represent a state from the data center as an array combining a bit vector corresponding to the current host and it's allocated bandwidth, free bandwidth, packet loss, packet overlimits, and queue length:\\
 
-\[ [0 0 0 ... 1 ... 0 0, bandwidth, free bandwidth, packet loss, overlimit, queue length ] \]
+[$0 0 ... 1 ... 0$, b/w, free b/w, packet loss, overlimit, queue length ]\\
+
 The advantage of this is that Iroko will learn a policy which considers all hosts together in unified representation instead of using separate agents for each network which wouldn't consider bandwidth allocation to each separate host.

### 2017-12-12 Documents/Report/conclusion.tex
Edits

@@
 \label{sec:conclusion}
+
+Iroko represents a first attempt at using machine learning
+to predict data center traffic. The eandeavour relies crucially
+on data center traffic having enough structure to in fact allow
+some level of prediction. Yet even for largely random traffic
+Iroko showed decent results, often besting baseline solutions.
+Iroko is only the opening salvo, but the attempt has merit. 
+More work must be done before claims of superiority to all
+incumbent solutions can be made, but there are reasons to be
+optimistic.
+
+Machine learning is perhaps the king of "black-box" solitions, 
+and it's efficacy is undeniable if sometimes misleading. While
+the standoff between so called "white-box" and "black-box" is 
+likely to continue for the forseeable future, Iroko shows
+that "black-box" solutions do have a role to play. It may turn
+out that neither comes to rule, but rather future SDN's will 
+employ both techinques to build ever more efficient solutions.
+It may well turn out that the future is grey.

### 2017-12-12 Documents/Report/discussion.tex
Edits

@@
 
-For efficient routing, we used EMCP with Iroko, however betting
-routing regimes exist, in particular Hedera. While we compared
+For efficient routing, we used EMCP with Iroko, however better
+routing regimens exist, in particular Hedera. While we compared
 Iroko to Hedera, in fact the two systems are largely orthogonal
-and may benefit, from one another. Iroko does not directly deal
+and may in fact benefit from one another. Iroko does not directly deal
 with routing or so called elephant flows. Thus Iroko could be
-futher improved by employing a routing scheme similar to that of
+further improved by employing a routing scheme similar to that of
 Hedera, while still retaining bandwidth prediction capabilities.

### 2017-12-12 Documents/Report/conclusion.tex
Speeling

--- 2017-12-10 Documents/Report/conclusion.tex
@@
 \label{sec:conclusion}
+
+Iroko represents a first attempt at using machine learning
+to predict data center traffic. The endeavour relies crucially
+on data center traffic having enough structure to in fact allow
+some level of prediction. Yet even for largely random traffic
+Iroko showed decent results, often besting baseline solutions.
+Iroko is only the opening salvo, but the attempt has merit. 
+More work must be done before claims of superiority to all
+incumbent solutions can be made, but there are reasons to be
+optimistic.
+
+Machine learning is perhaps the king of "black-box" solutions, 
+and it's efficacy is undeniable if sometimes misleading. While
+the standoff between so called "white-box" and "black-box" is 
+likely to continue for the foreseeable future, Iroko shows
+that "black-box" solutions do have a role to play. It may turn
+out that neither comes to rule, but rather future SDN's will 
+employ both techniques to build ever more efficient solutions.
+It may well turn out that the future is grey.

### 2017-12-12 Documents/Report/conclusion.tex
Edits

@@
 Iroko showed decent results, often besting baseline solutions.
-Iroko is only the opening salvo, but the attempt has merit. 
+Iroko is only the opening salvo, but the attempt has merit.Iroko
+struggles with reducing drop rate, and some random traffic
+patterns. With more fine-tuned
+prediction, and better rate-limiting algorithms, Iroko can likely
+overcome these challenges. 
 More work must be done before claims of superiority to all

### 2017-12-12 Documents/Report/conclusion.tex
Edits

--- 2017-12-12 Documents/Report/conclusion.tex
@@
 Iroko represents a first attempt at using machine learning
-to predict data center traffic. The endeavour relies crucially
-on data center traffic having enough structure to in fact allow
-some level of prediction. Yet even for largely random traffic
+to predict datacenter traffic. The endeavour relies crucially
+on datacenter traffic having enough structure to in fact allow
+some level of prediction. Yet even for largely random traffic,
 Iroko showed decent results, often besting baseline solutions.
-Iroko is only the opening salvo, but the attempt has merit. 
+For more predictable traffic, Iroko can compete with "best-in-class"
+solutions, and at times shows the strongest performance of all the
+tested solutions.
+Iroko is only the opening salvo, but the attempt has merit. Iroko
+struggles with reducing drop rate, and some random traffic
+patterns. With more fine-tuned
+prediction, and better rate-limiting algorithms, Iroko can likely
+overcome these challenges. 
 More work must be done before claims of superiority to all
@@
 Machine learning is perhaps the king of "black-box" solutions, 
-and it's efficacy is undeniable if sometimes misleading. While
-the standoff between so called "white-box" and "black-box" is 
+and its efficacy is undeniable, if sometimes misleading. While
+the standoff between so called "white-box" and "black-box" solutions is 
 likely to continue for the foreseeable future, Iroko shows
-that "black-box" solutions do have a role to play. It may turn
-out that neither comes to rule, but rather future SDN's will 
+that "black-box" solutions do have a role to play. It might be the case
+that neither comes to rule, but rather future SDN's will 
 employ both techniques to build ever more efficient solutions.

### 2017-12-12 Documents/Report/design.tex
Edits

@@
 global network view.~\footnote{For now, we do not expect our design to scale up 
-to thousands of switches. Iroko is intended for small to mid-tier size data 
-centers, which may benefit from a simplistic, centralized scheduling model.}
+to thousands of switches. Iroko is intended for small to mid-tier size datacenters,
+which may benefit from a simplistic, centralized scheduling model.}
 By polling each switch individually for port and utilization statistics we are 
@@
 allocated to a given host even if it not utilizing network resources. Thus, the 
-network, by design, will never reach 100\% utilization.
-
-Conversely however one should rarely see dropped packets due to network 
+network, by design, will never reach 100\% utilization. Conversely, one should rarely see
+dropped packets due to network 
 congestion. Furthermore, the controller must be able to react quickly to 
-changing bandwidth needs for hosts. Ideally, using a predictive method.
+changing bandwidth needs for hosts, ideally, using a predictive method.
 
@@
 switch and the current end-host to rate limit.
-This decision is based on the fact that these data are easy to collect and provides a current snapshot of the performance of the data center.
-Using this representation, we take full advantage of all the information provided in our data center.
+This decision is based on the fact that these data are easy to collect and provides a current snapshot of the performance of the datacenter.
+Using this representation, we take full advantage of all the information provided in our datacenter.
 
@@
 One limitation of the current reward signal is it's inability to distinguish cases when loss is 0 because no bandwidth was used in the epoch.
-This means, a more intricate reward functions can be devised for  a future work.
+This means, more intricate reward functions can be devised for  a future work.
 
-Given the data center assumption, the number of end-hosts is known in the 
+Given the datacenter assumption, the number of end-hosts is known in the 
 topology. We define our arbiter's actions as a single continuous value which will either increase 
@@
 actions, we will store the state-action value, also referred to as the Q(s,a) function,  in an artificial neural network. The Q(s,a) function 
-stores the reward signal information that is used to make future action choices.
-We chose an artificial neural network to approximate this function, because of they are theoretically proven function approximators ~\cite{Cybenko1989} which have made intractable environmental representations manageable~\cite{Sutton:1998:IRL:551283}.
+stores the reward signal information that is used to make future action choices. We chose an artificial neural network to approximate this function. Neural networks are theoretically proven function
+approximators ~\cite{Cybenko1989} which have made intractable environmental representations manageable~\cite{Sutton:1998:IRL:551283}.
 
@@
 
-When a node opens a flow, it will look up the current bandwidth restriction for 
-the destination IP in its token database and calculate the maximum congestion 
+When a node opens a flow, it looks up the current bandwidth restriction for 
+the destination IP in its token database and calculates the maximum congestion 
 window possible for the particular TCP connection. In our simple design, flows 
@@
 \centering
-\includegraphics[width=1\linewidth]{Topology3}
+\includegraphics[width=1\linewidth]{Iroko}
 \caption{Simple sketch of the Iroko architecture. A central scheduler maintains 

### 2017-12-12 Documents/Report/discussion.tex
Edits

--- 2017-12-10 Documents/Report/discussion.tex
@@
 
-Rather predictably, Iroko struggles on random traffic patterns, however we
-do not see this as particularly concerning. Firstly \textit{Iroko} still outperforms
-both DCTCP and ECMP implementations. Further, as these traffic patterns
+Not unexpectedly, Iroko struggles on a few random traffic patterns, however we
+do not see this as particularly concerning. Firstly, Iroko still outperforms
+both DCTCP and ECMP implementations on most random traffic patterns, and in general.
+Further, as these traffic patterns
 were designed by the authors of the Hedera paper \cite{hedera}, it remains
 likely these patterns are somewhat amenable to the Hedera implementation.
-Finally random traffic patterns seem unlikely to be prevalent in a
+Finally, as discussed, fully random traffic patterns seem unlikely to be prevalent in a
 datacenter context. While the general predictability of datacenter
@@
 Our machine learning approach shows promise, but machine learning remains
-largely a black-box discipline. This nature makes it difficult for us
+largely a "black-box" discipline. This nature makes it difficult for us
 to determine precisely which parameters to use, and what drives
@@
 in our test network. A heavier handed approach may yield better
-results, however these may come at the cost of scalability.
+results, however these may come at the cost of scalability. We
+also note that further tuning of the current implementation is
+likely to lead to stronger results. Due to testing limitations, 
+the controller had a very short number of periods to learn 
+on. In longer running cases, we fully expect Iroko to 
+deliver stronger results as there is more time for machine learning 
+algorithms to converge.
+
+For efficient routing, we used EMCP with Iroko, however better
+routing regimens exist, in particular Hedera. While we compared
+Iroko to Hedera, the two systems are largely orthogonal
+and may in fact benefit from one another. Iroko does not directly deal
+with routing or so called elephant flows. Thus, Iroko could be
+further improved by employing a routing scheme similar to that of
+Hedera, while still retaining bandwidth prediction capabilities.

### 2017-12-12 Documents/Report/eval.tex
Edits

--- 2017-12-11 Documents/Report/eval.tex
@@
 
-This section describes the evaluation of Iroko using our testbed built on top of
+This section describes the evaluation of Iroko using our testbed built on the
 Mininet simulator. The main goal is to compare the performance against the
 existing centralized and decentralized techniques like Hedera~\cite{hedera} and
-DCTCP~\cite{dctcp} respectively.
+DCTCP~\cite{dctcp}.
 
@@
 
-For evaluation purposes we are targeting a small scale data center of 16 hosts
-and 20 switches with FatTree topology. We took measurements for Hedera, DCTCP,
-ECMP and with Non-blocking switch on the same test setup. ECMP measurement is
-considered as the base line. Non-blocking is a hypothetical single switch fully
-connected setup and is considered as the theoretical maximum we are aiming to
-achieve.
+For evaluation purposes, we are targeting a small scale data center of 16 hosts
+and 20 switches with a FatTree topology. We take measurements for Hedera, DCTCP,
+ECMP and a non-blocking switch on the same test setup. We consider the ECMP measurement
+to be the base line. The non-blocking result represents a hypothetical single switch with full
+backplane bandwidth, and is considered the theoretically achievable maximum.
 
-For traffic simulation, we are using the same traffic matrices used in
-Hedera~\cite{hedera}. We took measurements for the variations of the following 3
+For traffic simulation, we use the same traffic matrices used in
+Hedera~\cite{hedera}. We measured the following 3
 traffic patterns:
@@
 the host with index $(x + i)mod(num hosts)$.  \item Staggered Prob $\left(EdgeP,
-PodP\right)$: A host sends to another host in the same edge switch with
+PodP\right)$: A host sends to another host on the same edge switch with
 probability $EdgeP$, and to its same pod with probability $PodP$, and to the
@@
 while keeping utilization at maximum, we measured bisection bandwidth and
-    average queue length on switches. The experiments last for 60 seconds for
+    average queue length on the switches. The experiments lasted for 60 seconds for
     each benchmark and we consider only the middle 40 seconds.
@@
 consistently provides better utilization compared to ECMP and DCTCP in almost
-all experiments. As expected non-blocking is providing the best performance out
+all experiments. As expected, non-blocking is providing the best performance out
 of all four techniques. Utilization of non-blocking is lower for traffic
-patterns with high incast. Hedera perform extremely well in all experiments
-except for some where Iroko outperforms Hedera. Interestingly enough, Hedera
-even outperforms non-blocking limit in stride8 experiment. ECMP does reasonably
-well when there is high probability of local communication for example smaller
-strides, but start to deteriorate as the stride length increases because of hash
-collisions.
+patterns with high incast. Hedera performs extremely well in all experiments
+except for a few like stride1, stag2$(0.5,0.3)$ where Iroko outperforms Hedera.
+Interestingly enough, Hedera even outperforms non-blocking in the stride8
+experiment. ECMP does reasonably well when there is high probability of local
+communication; for example smaller strides, but starts to deteriorate as the
+stride length increases because of hash collisions.
 
 Another interesting observation is that DCTCP gives abysmal utilization for some
-experiments. Further investigation into this revealed that this is because of
-DCTCP's aggressive congestion window size reduction strategy on incast. The
-proposed solution for incast problem is to use dynamic buffer allocation in
-switches which is not implemented in our test setup. General performance of
-DCTCP is not satisfactory either. For some experiments it even under performs
-ECMP.  According to DCTCP paper it is optimized for some specific data center
-traffic patterns like web search, soft real-time applications, recommendation
-systems etc which we don't include in our experiments.
+experiments, for example, randx2 and randx4. Further investigation into this
+revealed that this is because of DCTCP's aggressive congestion window size
+reduction strategy on incast. The proposed solution for incast problems is to use
+dynamic buffer allocation in switches, which is not implemented in our test
+setup. General performance of DCTCP is not satisfactory either. For some
+experiments it even under-performs ECMP.  According to the DCTCP paper, it is
+optimized for some specific data center traffic patterns like web search, soft
+real-time applications, recommendation systems etc., which we don't include in our
+experiments.
+
+\subsection{Drop rate}
+
+\begin{figure*}
+	\centering
+	\includegraphics[width=1\linewidth]{qlen.png}
+	\caption{CDF of queue lengths}
+	\label{fig:qlen}
+\end{figure*}
+
+To analyze the drop rate we measured the number of packets built up in the queue
+on all switches for the same set of traffic patterns and plotted a cumulative
+distribution graph. In Figure \ref{fig:qlen} shows some selected
+graphs. Recall that higher queue utilization is considered less optimal. For our experiments, we fixed the maximum
+queue size at 50 packets.
+
+Iroko struggles with queue management; consistently keeping $90\%$ of the queues
+empty while the remaining $10\%$ are full. It is worth noting however, that Iroko performs
+very well for traffic patterns like stride1, stag2$(0.5,0.3)$ etc., where it
+shows high utilization. DCTCP maintains the queues at a minimum across
+all experiments. This is not surprising considering the low utilization of
+DCTCP. ECMP and Hedera behave almost the same. Non-blocking has the highest
+queue size out of all techniques. We surmise that the CPU is the bottleneck here;
+non-blocking uses only a single switch to connect the entire network with full
+backplane bandwidth.
+
+Iroko could definitely benefit from more fine tuning for the prediction algorithm. We further
+need to investigate why the CDF of queue lengths exhibits a sudden jump from $0$ to $50$. Initial
+analysis shows that a small fixed set of switches are the bottleneck in all
+experiments. Our guess is that this happens because of the unfair bandwidth
+allocation across traffic flows. Analyzing fairness might provide better insight
+into this problem.
 
@@
 
-Apart from utilization and loss rate following are the other benchmarks that we
-are interested. We have not measured these yet due to time concerns.
+Apart from utilization and loss rate, the following benchmarks remain of interest.
+Due to time constraints, we were not able to fully investigate these.
 
@@
 
-\item \textbf{99th percentile latency:} Since we are aiming for a low latency
-which means that 99th percentile latency in the network across all flows should
+\item \textbf{99th percentile latency:} We are aiming for a low latency solution,
+which means that the 99th percentile latency in the network across all flows should
 be as low as possible.
@@
 to see if all the hosts gets fair share of the total bandwidth. Currently we are
-assuming all flows should be at equal priority. Differential priority is out of
+assuming that all flows should be given equal priority. Differential priority is out of
 scope.

### 2017-12-12 Documents/Report/future_work.tex
Edits

@@
 future widely deployed datacenter traffic control mechanisms will include
-machine learning in their toolkit. However machine learning alone may not
+machine learning in their toolkit. However, machine learning alone may not
 be sufficient to achieve the best possible results. Machine learning
@@
 avenue to pursue. Combined with a robust routing algorithm, we
-believe such a solution show superior performance when
-compared to virtual all currently existing solutions.
+believe such a solution could show superior performance when
+compared to virtually all currently existing solutions.
 
@@
 allowing the controller to account for them pre-emptively
-rather than trying to infer them which may take a couple
+rather than trying to infer them, which may take a couple
 of epochs, and therefore reduce efficiency. Such a

### 2017-12-12 Documents/Report/implementation.tex
Edits

--- 2017-12-11 Documents/Report/implementation.tex
@@
 \subsection{Topology Simulation}
-We chose to use  Mininet, a python data center emulation API, to prototype Iroko.
-Mininet's implementation leverages the Linux kernel to provide a lightweight virtualization of a data center in a single machine ~\cite{mininet}.
-This allowed us to create a prototype of Iroko we can test quickly using limited amounts of hardware.
-Other methods of testing Iroko would require more resources, or specialized code specific to the testbed which cannot be deployed in a real system ~\cite{mininet}.
-In contrast, Mininet has allowed us to develop Iroko such that we could deploy it in a real system without re-implementation independent of testbed specific code.
+We chose to use  Mininet, a python datacenter emulation API, to prototype Iroko.
+Mininet's implementation leverages the Linux kernel to provide a lightweight virtualization
+of a datacenter in a single machine ~\cite{mininet}. This allowed us to create a prototype of
+Iroko we could test quickly using limited amounts of hardware. Other methods of testing Iroko 
+would require more resources, or specialized code specific to the testbed which cannot be 
+deployed in a real system ~\cite{mininet}. In contrast, Mininet has allowed us to develop
+Iroko such that we could deploy it in a real system without re-implementation independent of testbed specific code.
 
-In the Mininet environment, we created a fat-tree topology consisting of 20 switches and 16 hosts .
-The 20 switches are configured into 5 pods as described in ~\cite{fattree}.
-This design choice was to allow a direct comparison to Hedera  ~\cite{hedera}, a global arbiter which uses flow statistics to re-route elephant flows.
-In ~\cite{hedera}, Hedera was evaluated  using a fat-tree topology as we have described, and this allows us to directly compare Iroko against Hedera's performance .
+In the Mininet environment, we created a FatTree topology consisting of 20 switches and 16 hosts.
+The 20 switches were configured into 5 pods as described in ~\cite{fattree}. This design choice was
+to allow a direct comparison to Hedera  ~\cite{hedera}, a global arbiter which uses flow statistics
+to re-route elephant flows. In ~\cite{hedera}, Hedera was evaluated using a FatTree topology as
+we have described, and this allows us to directly compare Iroko against Hedera's performance .
 
-Mininet’s lightweight virtualization also allows the API to scale such as to support bigger topologies which otherwise couldn’t be prototyped easily ~\cite{mininet}.
-This feature will be useful for future research, because we can then test Iroko on other topologies beyond the fat-tree topology we looked at in this work.
+Mininet's lightweight virtualization also allows the API to scale; supporting bigger topologies which otherwise couldn't be prototyped easily ~\cite{mininet}.
+This feature will be useful for future research. Using Mininet, we can test Iroko on other topologies beyond the FatTree topology.
 
 \subsection{Data Collection }
-Since we knew prior to testing which port connected to each host, we collected different data from port connections to each end-host.
+Since we knew prior to testing which port connected to each host,
+we collected different data from port connections to each end-host. 
 This was accomplished using the awk scripting language and the Linux tc qdisc command to collect information.
@@
 Awk is a text parsing language available on most Linux distributions by default.
-Our awk script reads the /proc/net/dev file on each interface to collect the currently used  bandwidth on a link.
-We then used this to calculate free bandwidth which is the difference between theoretical maximum bisection bandwidth, and the reported bandwidth on a link.
+Our awk script reads the /proc/net/dev file on each interface to collect the currently
+used  bandwidth on a link. We then used this to calculate free bandwidth which is the difference
+between theoretical maximum bisection bandwidth, and the reported bandwidth on a link.
 
 In addition, we used the following Linux command:\\
+\vspace{1mm}
+
 \texttt{tc -s qdisc show dev <interface>}\\
+
+\vspace{1mm}
 where \texttt{<interface>} is the current host we are polling from. Using the 
 obtained text, we then applied regex expressions to collect the packet loss, 
-packet overlimits, and interface queues. Each of these data provide further 
-information to hint at the current state of the data center. Our goal was to 
-eliminate dropped packets, and used this to help Iroko make rate-limiting 
-decisions. Overlimit packets provides information on when a link goes over it’s 
+packet overlimits, and interface queues. Each of these data points provide further 
+information about the current state of the datacenter. Our goal was to 
+eliminate dropped packets. We used this technique to help Iroko make rate-limiting 
+decisions. Overlimit packets provide information on when a link goes over its 
 allocated bandwidth, and adds a potentially useful feature for our controller 
@@
 length of the host, which measures how many packets are buffered to be 
-processed.  These additional statistics we hoped would provide our controller 
+processed.  We hoped these additional statistics would provide our controller 
 enough data to make meaningful inferences of the appropriate bandwidth 
@@
 \subsection{Bandwidth Allocation}
-Iroko rate limits traffic at the applications level, and thus uses UDP instead of TCP to communicate with hosts.
-It reads a traffic matrix shared among all the hosts and sends network traffic accordingly.
-Iroko also manages the current allocated bandwidth per host by sending a packet which updates their current allocated bandwidth.
+Iroko rate limits traffic at the application level, and thus uses UDP instead
+of TCP to communicate with hosts. Traffic is generated using a custom load generation
+application. We wrote this application to read a traffic matrix shared among all the hosts and send network traffic accordingly.
+Iroko manages the current allocated bandwidth per host by sending a controller packet
+which updates the current allocated bandwidth.
 
@@
 
-We chose to use the deep deterministic policy gradient policy algorithm (DDPG) almost entirely as described in ~\cite{DDPG} .
-DDPG is a form of actor-critic reinforcement learning algorithm ~\cite{Sutton:1998:IRL:551283} which uses neural networks to approximate $Q(s,a)$, and learns the policy by taking the actor network gradients relative to the critic $Q(s,a)$ approximations ~\cite{DDPG}.
-Our agent learns using replay memory, which is a technique that helps deep reinforcement learning algorithms learn by making the data uncorrelated via random batch sampling ~\cite{DQLearning}.
-This decision might be unnecessary for a data center where authors have suggested that data center traffic is inherently unpredictable ~\cite{microte} and this suggests that we could assume data center data is inherently uncorrelated.
-We recognize these findings may also give an upper limit to the expected performance of Iroko.
+We chose to use the deep deterministic policy gradient
+policy algorithm (DDPG) as described in ~\cite{DDPG} .
+DDPG is a form of actor-critic reinforcement learning algorithm ~\cite{Sutton:1998:IRL:551283}
+which uses neural networks to approximate $Q(s,a)$, and learns the policy by taking the actor
+network gradients relative to the critic's $Q(s,a)$ approximations ~\cite{DDPG}.  Our agent learns using replay memory;
+a technique that helps deep reinforcement learning algorithms learn by
+making the data uncorrelated via random batch sampling ~\cite{DQLearning}.
+This decision might be unnecessary for a datacenter as datacenter traffic may be inherently unpredictable ~\cite{microte}.
+Thus we can reasonably assume datacenter traffic is inherently uncorrelated. We recognize these findings
+may also give an upper limit to the expected performance of Iroko.
 
-To find the optimal values, an agent must balance exploiting the current best actions with exploring an environment with a new action which may lead to bigger future rewards ~\cite{Sutton:1998:IRL:551283}.
-Instead of the exploration approach used in ~\cite{DDPG}, we use $\epsilon-greedy\ learning$, where an agent takes the best action with probability $\epsilon$ and a random action with probability $1 - \epsilon$ ~\cite{Sutton:1998:IRL:551283}.
-Our random action is then defined as follows:
+To find the optimal values, an agent must balance exploiting
+the current best actions with exploring an environment with 
+a new action which may lead to bigger future rewards ~\cite{Sutton:1998:IRL:551283}.
+Instead of the exploration approach used in ~\cite{DDPG}, we use $\varepsilon-greedy\ learning$,
+where an agent takes the best action with probability $\varepsilon$ and a random action with probability
+$1 - \varepsilon$ ~\cite{Sutton:1998:IRL:551283}. Our random action is defined as follows:
 
@@
 
-uniform(0,1), &\text{P(x = increase) = .333},\\
-uniform(-1,0), &\text{P(x = decrease) = .3333},\\
-0, & \text{(x = unchanged) = .3333}
+uniform(0,1), &\text{P(x = increase) = 1/3},\\
+uniform(-1,0), &\text{P(x = decrease) = 1/3},\\
+0, & \text{P(x = unchanged) = 1/3}
 \end{cases}
 \]
+\\
 
 where uniform(a,b) refers to taking a random value in the range between [a, b]
-This means we will take the best predicted value $(.5 * 1 + .5 * .33) =  ~.66.6\%$  and will increase $P(x= increase) * P(\epsilon) = 16.
-667\%$ of the time or decrease $P(x= decrease) * P(\epsilon) = 16.
-667\%$ of the time.
-It should be noted that the action value is still clamped between -0.9 and 0.9 regardless of the noise introduced.
-Alternative exploration policies could be used, but this one was chosen due it’s simplicity, and theoretical capability to explore the whole action space if allowed to run indefinitely.
+This means we will take the best predicted value $(.5 * 1 + .5 * .33) =  ~.66.6\%$  
+and will increase $P(x= increase) * P(\varepsilon) = 16. 667\%$ of the time or
+decrease $P(x= decrease) * P(\varepsilon) = 16. 667\%$ of the time.
+It should be noted that the action value is still clamped between
+-0.9 and 0.9 regardless of the noise introduced. Alternative exploration policies 
+could be used, but this one was chosen due it's simplicity, and theoretical capability to explore the whole action space if allowed to run indefinitely.
 
-We represent a state from the data center as an array combining a bit vector corresponding to the current host and it's allocated bandwidth, free bandwidth, packet loss, packet overlimits, and queue length:
+We represent the traffic state of the datacenter as an array combining a bit vector
+corresponding to the current host and its allocated bandwidth, free bandwidth, packet loss, packet overlimits, and queue length. 
+The advantage of this technique is that Iroko will learn a policy which considers all hosts together 
+in a unified representation.
 
-\[ [0 0 0 ... 1 ... 0 0, bandwidth, free bandwidth, packet loss, overlimit, queue length ] \]
-The advantage of this is that Iroko will learn a policy which considers all hosts together in unified representation instead of using separate agents for each network which wouldn't consider bandwidth allocation to each separate host.
-
-The action Iroko takes at each step is the amount to increase or decrease the current hosts allocated bandwidth.
-Iroko will generate an adjustment value clamped between $-.9 - .9$.
-We chose this value range, because we wanted to avoid host starvation.
-Iroko then modifies the allocated bandwidth similar to the approach used in ~\cite{remy} for congestion windows by using the action with allocated bandwidth as follows:
+At each time-step, Iroko increases or decreases the hosts' allocated bandwidth. 
+Iroko generates an adjustment value clamped between $-0.9$ and $0.9$. We chose this value range to avoid host starvation. 
+Iroko modifies the allocated bandwidth similar to the approach used in ~\cite{remy} for congestion windows by using the action with allocated bandwidth as follows:
 
 \[bandwidth_{t+1}   \leftarrow bandwidth_t +  a_t * bandwidth_t\]
-
-where a is the chosen bandwidth adjustment, t is the previous epoch, and t+1 is the new allocation.
-This representation could be  modified in a number of ways, such as forcing a minimum expected bandwidth, or add the capacity to starve a host.
+\\
+where $a$ is the chosen bandwidth adjustment, $t$ is the previous epoch, and $t+1$ is the current epoch. 
+This representation could be  modified in a number of ways, for example: enforcing a minimum expected bandwidth, or adding capacity to a starved host.

### 2017-12-12 Documents/Report/intro.tex
Edits

@@
 Developments in the past decade have changed the general networking 
-environment. Data centers have emerged as an exciting new research frontier, 
+environment. Datacenters have emerged as an exciting new research frontier, 
 posing novel design challenges and opportunities. Driven by minimization of 
@@
 
-Despite the innovation potential of data centers, TCP has dominated as 
+Despite the innovation potential of datacenters, TCP has dominated as 
 the congestion control algorithm of choice.
@@
 under the fairness principle. This design has not fundamentally changed 
-over the last thirty years, and new research advancements raise the question if TCP can still be considered to be the most sensible option.~\cite{pcc,bbr,perc}
+over the last thirty years, and new research advancements raise the question whether TCP
+can still be considered the most sensible option.~\cite{pcc,bbr,perc}
 
@@
 TCP.~\cite{expresspass, fastpass, perc}
-While idea of admission control and service guarantees in networks is not 
+While the idea of admission control and service guarantees in networks is not 
 new~\cite{access_limit, access_limit2}, these designs traditionally 
-aim to assure quality and bandwidth guarantees in a contentious, decentralized, 
-and untrusted environments such as the internet. In a data center, these 
+aim to assure quality and bandwidth guarantees in contentious, decentralized, 
+and untrusted environments such as the internet. In a datacenter, these 
 conditions do not apply. End-hosts are generally considered reliable and 
@@
 and prioritization policies. Admission-based systems can give administrators 
-the advantage of tight control over the data center, mitigating the effects of 
+the advantage of tight control over the datacenter, mitigating the effects of 
 queue-buildup and bursts.
 
-Although it is unclear how predictable data center traffic truly is, various 
+Although it is unclear how predictable datacenter traffic truly is, various 
 works have shown that repeated patterns may 
 exist.\cite{msr_dc,fb_dc,traffic,microte}
-If data center traffic exhibits predictable patterns, it may be 
+If datacenter traffic exhibits predictable patterns, it may be 
 possible to devise a proactive mechanism which forecasts the traffic matrix of 
@@
 the benefits and drawbacks of centralized learning and admission control in the 
-data center context.
+datacenter context.

### 2017-12-12 Documents/Report/related_work.tex
Edits

@@
 The first, Data Center TCP (DCTCP)~\cite{dctcp}, is a congestion control 
-algorithm developed to reduce bufferbloat in data center networks. DCTCP 
-leverages Explicit Congestion Notifications (ECNs) as early warning signal to 
+algorithm developed to reduce bufferbloat in datacenter networks. DCTCP 
+leverages Explicit Congestion Notifications (ECNs) as early warning signals to 
 identify queue buildup and adjusts its sending rate to the amount of ECN-marked 
@@
 
-The second algorithm we compare to is TCP CUBIC~\cite{cubic}, the default 
+The second algorithm we compare is TCP CUBIC~\cite{cubic}, the default 
 congestion control protocol of the Linux kernel. CUBIC uses a cubic function   
@@
 assumption that a subset of short-term traffic is predictable. The controller 
-isolates flows which it determined to be predictable and install a custom 
+isolates flows which it determined to be predictable and installs a custom 
 optimal routing entry. The remaining unpredictable traffic is routed according 
@@
 it is guaranteed to experience zero queuing latency and jitter during 
-transmission. This "zero-loss" network provides application with a stable and 
+transmission. This "zero-loss" network provides applications with a stable and 
 reliable transmission experience. However, FastPass still works reactively and 
-only operates in incoming requests. Queueing is thus pushed to end-hosts and 
+only operates on incoming requests. Queueing is thus pushed to end-hosts and 
 the arbiter itself.
@@
 However, we intend to go further and schedule rate-limits ahead of time in 
-order to reduce sending delays and host starvation of admission-control systems 
-substantially.
+order to substantially reduce sending delays and host starvation of admission-control systems.
 
@@
 trainability of network traffic.~\cite{throwdown, learning_tcp}
-A common assumption is that data center traffic is largely unpredictable and 
+A common assumption is that datacenter traffic is largely unpredictable and 
 untrainable.~\cite{fb_dc}
-However, Winstein and Balakrishnan,~2013\cite{remy} demonstrated that it is 
+However, Winstein and Balakrishnan,~2013 \cite{remy} demonstrated that it is 
 possible to learn near-optimal TCP congestion control parameters for various 
 network topologies. The system, Remy, represents an offline learning algorithm 
-inferring optimal TCP parameters based on pre-supplied topology and 
-traffic-model.
+inferring optimal TCP parameters based on pre-supplied topologies and 
+traffic-models.
 While Remy only performed static offline training, PCC~\cite{pcc,throwdown} 
@@
 As part of the contribution of MicroTE, Benson et 
-al.,~2014~\cite{microte} analyzed data center traffic and 
+al.,~2014~\cite{microte} analyzed datacenter traffic and 
 inferred that around 30\% of short-lived traffic is predictable over a period 
